---
title: "Gilchrist- Bootstrapping Motif Mean and Variance"
author: "Tara Empson"
date: "`r Sys.Date()`"
output: html_document
---

# Goal

Create a bootstrapped dataset from the mean number of songs from male zebra finch, repeatidly tested in 40C.

# Background

Kayci Messerly, a previous post-bac, preformed multiple tests of trying to see how male zebra finch song varies in heat. Being an animal behavoir study there is a lot of variation and a very limited sample size. 

# Libraries

```{r}
library(tidyverse)
library(modelr)
library(infer)
library(mosaic)
library(janitor)
library(boot)
```

# Local Functions

```{r}

# Function won't work if you use function(..., raw.data = raw.data)
# Why? R doesn't know which raw.data you mean (local or global)
# This is why you sometimes see .data as an argument name

bootstrapMeanAndVarOfColumn<- function(male, sample = 1000, col.name = "motif_count", data.raw = raw.data) {

  my_means <- vector(mode="double", length = sample)
  my_vars <- my_means

  data <- data.raw %>%
    filter(male %in% male) %>% 
    select(male, all_of(col.name)) # tidy warning told me to use 'all_of()' around col.name
  for(i in 1:sample){
    bootstrap_sample <- sample(raw.data$motif_count,n, replace = TRUE)
    my_means[i] <- mean(bootstrap_sample)
    my_vars[i] <- var(my_means)   
  }
  results <- tibble(male = male, mean =  my_means, var = my_vars)
  return(results)
}

```
# Bootstraping data

## Load Data & Define Parameters

```{r}
setwd("../../data/collated/")
set.seed(1)
raw.data <- read.csv("HSPi-Repeatability-Song-Count.csv")
raw.data$day<- c(1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3)
#adding in day order

unique.male <- unique(raw.data$male)

n_sample <- 1000
n <- 3
#my_means <- 1:1000
#my_var <- 1:1000
```

## Bootstrap Data to Get Sample of Means and Vars for Each Male

```{r}



## Create tibble of mean and var from bootstrapped data
bs_stats  <- lapply(unique.male, bootstrapMeanAndVarOfColumn) %>%
  ## Combine bootstraped stats across males
  ## Note bind_rows will only work for a list, if you use sapply() you'll have a matrix of matricies
  bind_rows()

dim(bs_stats)

```

## Editing data to be the mean of the bootstrapped means and mean of the bootstrapped variance. I couldn't think of an efficient way to only keep one row for each male.
```{r}
simplified<- bs_stats %>% 
    group_by(male) %>% 
    mutate(mean.mean=mean(mean)) %>% 
    mutate(mean.var= mean(var)) %>% 
    select(male, mean.mean, mean.var)
```



# I was able to make a good graph
- the Linear and Quad curves don't seem very different.
```{r}
simplified %>% 
    ggplot(aes(x=mean.mean, y=mean.var))+
     geom_point()+
    geom_smooth(method = "lm", formula = y ~ x, color="blue")+
    geom_smooth(method = "lm", formula = y ~ x + I(x^2), color="red")+
    labs(x="Mean of Mean Number of Motifs", y="Variance of Mean Number of Motifs", caption = "Red line is quadradic and Blue is linear.", title="Bootstraping 1000 replications for each bird")+
    theme_classic()
```


