---
title: "Gilchrist- Bootstrapping Motif Mean and Variance"
author: "Tara Empson"
date: "`r Sys.Date()`"
output: html_document
---

# Goal

Create a bootstrapped dataset from the mean number of songs from male zebra finch, repeatidly tested in 40C.

# Background

Kayci Messerly, a previous post-bac, preformed multiple tests of trying to see how male zebra finch song varies in heat. Being an animal behavior study there is a lot of variation and a very limited sample size. 

# Libraries

```{r message=FALSE}
library(tidyverse)
library(modelr)
library(infer)
library(mosaic)
library(janitor)
library(boot)
library(brms)
```

# Bootstraping data

## Load Data & Define Parameters
```{r}
setwd("../../data/collated/")
set.seed(1)
raw.data <- read.csv("HSPi-Repeatability-Song-Count.csv")
raw.data$day<- c(1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3)
#adding in day order
unique.male <- unique(raw.data$male)
n <- 3
sample = 1000
col.name = "motif_count"
my_means <- vector(mode="double", length = sample)
my_vars <- my_means
```

### Bootstrap still wasn't pulling from within each individual. Worked on it more for a while and got it working. So now the male who did not sing in any trial has 0 mean and 0 var.
```{r}
boot_male <-function(one_male, boot_set){
  sample<-1000
  data <- raw.data %>%
    filter(male == one_male)
  for(i in 1:sample){
    bootstrap_sample <- sample(data$motif_count,n, replace = TRUE)
    my_means[i] <- mean(bootstrap_sample)
    my_vars[i] <- var(my_means)   
    results <- tibble(male = one_male, a.mean =  my_means, b.var = my_vars)
  }
boot_set<-  rbind(boot_set, results)
return(boot_set)
}

out <- tibble()
full.boot <- bind_rows(lapply(unique.male,boot_male, out))

```


```{r}
simplified<- full.boot %>% 
    group_by(male) %>% 
  mutate(sd.mean=sd(a.mean)) %>% 
  mutate(sd.var=sd(b.var)) %>% 
  mutate(mean.mean=mean(a.mean)) %>% 
  mutate(mean.var=mean(b.var)) %>% 
  select(-a.mean, -b.var)
 
simply <- unique(simplified)

```



# Graph Remake
```{r}
simply %>% 
    ggplot(aes(x=mean.mean, y=mean.var))+
     geom_point()+
  geom_errorbar(aes(ymin=mean.var-sd.var, ymax=mean.var+sd.var))+
  geom_errorbar(aes(xmin=mean.mean-sd.mean, xmax=mean.mean+sd.mean))+
    geom_smooth(method = "lm", formula = y ~ x, color="blue")+
    geom_smooth(method = "lm", formula = y ~ x + I(x^2), color="red")+
    labs(x="Mean of Mean Number of Motifs", y="Variance of Mean Number of Motifs", caption = "Red line is quadradic and Blue is linear.", title="Bootstraping 1000 replications for each bird")+
    theme_classic()
```

##Excluding T230
```{r}
simplified.omit <- simply %>% 
  filter(male!="T230")
simplified.omit %>% 
    ggplot(aes(x=mean.mean, y=mean.var))+
     geom_point()+
  geom_errorbar(aes(ymin=mean.var-sd.var, ymax=mean.var+sd.var))+
  geom_errorbar(aes(xmin=mean.mean-sd.mean, xmax=mean.mean+sd.mean))+
    geom_smooth(method = "lm", formula = y ~ x, color="blue")+
    geom_smooth(method = "lm", formula = y ~ x + I(x^2), color="red")+
    labs(x="Mean of Mean Number of Motifs", y="Variance of Mean Number of Motifs", caption = "Red line is quadradic and Blue is linear.", title="Bootstraping 1000 replications for each bird- T230 Omitted")+
    theme_classic()+
  annotate("text", x = 75, y = 12000, 
             label = "Not much changed",
             family = "Arial",
           color="red")
```

# 1. Filter out birds who didn't sing in a given round.
- I did this because we are interested in how song production changes at different temperatures, so birds who never sang are not relevant.
- All birds sang at least once in Round 2 so all were included.
```{r}
setwd("../../data/collated/")
round2 <- read.csv("HSPi-Round-2-Heat-Trials.csv") %>% 
  select(-mass,-bill_length,-bill_depth,-bill_width,-date_morph_data_collected,-time_spent_panting_30.15_before_female,-time_spent_panting_15.0_before_female,-time_spent_panting_0.15_after_female,-time_spent_panting_15.30_after_female)
r2 <- round2 %>% 
  pivot_wider(names_from = test_order, values_from = c(date, temp_target, motif_count, motif_rate, temp_mean, temp_median, humidity_mean))
```



# 2. Try to use BRM on round2 data. I don't understand this output and am stopping here.
```{r}
fit1 <- brm(motif_count ~ temp_target * male + (1|chamber),
            data = round2, family = poisson())
```
