---
title: "Gilchrist- Deming and Simulating Data"
author: "Tara Empson"
date: "`r Sys.Date()`"
output: html_document
---

# Goal

Create a linear model and run an anova from bootstrapped dataset from the mean number of songs from male zebra finch, repeatidly tested in 40C.

# Background

Kayci Messerly, a previous post-bac, preformed multiple tests of trying to see how male zebra finch song varies in heat. Being an animal behavior study there is a lot of variation and a very limited sample size.

# Libraries

```{r message=FALSE}
library(tidyverse)
library(janitor)
library(deming)
library(patchwork)
```

## Load Data & Define Parameter
-   Not loading in T230 because he did not sing at all during this round and Deming/Simulating didn't like 0s.
```{r}
repo.dir <- system("git rev-parse --show-toplevel", intern = TRUE)
setwd(file.path(repo.dir, "/code/Empson"))

data.dir <- file.path("..", "..", "data", "collated")
set.seed(0)
raw.data <- read.csv(file.path(data.dir, "HSPi-Repeatability-Song-Count.csv"))
filtered.data <- raw.data %>% 
    filter(male!="T230") %>%
    group_by(male) %>%
    mutate(test_order = rank(date)) %>% #adding in day order
    ungroup()

unique.male <- unique(filtered.data$male)
n.males <- length(unique.male)
n <- 3 # number of trials per male
w = 1000 # number of bootstrap and simulated samples

col.name = "motif_count"
my_means <- vector(mode="double", length = w)
my_vars <- my_means
out <- tibble()
simulated <- vector()
```

#Observed Data
```{r}
filtered.data <- filtered.data %>% 
    group_by(male) %>% 
    mutate(observed_sd=sd(motif_count), 
    observed_mean=mean(motif_count), 
    observed_var=var(motif_count))
```

#Graph
```{r}
observed.graph<-filtered.data %>% 
  ggplot(aes(x=observed_mean, y=observed_var))+
  geom_point(color="black")+
  theme_minimal()+
  xlim(0, NA)+
  ylim(0, NA)+
  labs(x="Mean", y="Var", title="Observed Data")
  #annotate("text", label=str_wrap("T231 had 2 trials under 40 motifs, but the other trial was 262 motifs", width = 40), x=170, y=18500)
observed.graph
observed.graph+scale_y_log10()+labs(caption="*Log*")

More.filtered.graph <- filtered.data %>% 
  filter(male!="T231", male!="T229") %>%
  ggplot(aes(x=observed_mean, y=observed_var))+
  geom_point(color="black")+
  theme_minimal()+
  xlim(0, NA)+
  ylim(0, NA)+
  labs(x="Mean", y="Var", title="Filtered out Extremes")
More.filtered.graph+scale_y_log10()+labs(caption="*Log*")
  
```

### Bootstrap Function
```{r}
boot_male <-function(one_male, data.from,boot_set){
  w<-1000
  data <- data.from %>%
    filter(male == one_male)
  for(i in 1:w){ #Check the first code chunk for what sample is
    bootstrap_sample <- sample(x=data$motif_count,size=3, replace = TRUE) #Check the first code chunk for what n is
    my_means[i] <- mean(bootstrap_sample)
    my_vars[i] <- var(bootstrap_sample)
    results <- tibble(male = one_male, a_mean = my_means, b_var = my_vars) #this should be 1000 bootstrapped points from a single male
  }
boot_set<-  rbind(boot_set, results) #binding every males bootstrap together
return(boot_set)
}
```

# Testing Bootstrap Function
- I think this is where the issue is.
- real mean is similar to bootstrapped but variance is VERY different.
```{r}
set.seed(3)
checking.real.T229 <- filtered.data %>% 
  filter(male=="T229")
mean(checking.real.T229$motif_count) #191.3333
var(checking.real.T229$motif_count) #13602.33

testing.bootstrap.T229<- boot_male(one_male="T229", data.from=filtered.data,boot_set=out)
mean(testing.bootstrap.T229$a_mean) #190.7173
mean(testing.bootstrap.T229$b_var) #8977.355



checking.real.T236 <- filtered.data %>% 
  filter(male=="T236")
mean(checking.real.T236$motif_count) #201
var(checking.real.T236$motif_count)  #775

testing.bootstrap.T236<- boot_male(one_male="T236", data.from=filtered.data,boot_set=out)
mean(testing.bootstrap.T236$a_mean) #200.67
mean(testing.bootstrap.T236$b_var) #510.3167



checking.real.T247 <- filtered.data %>% 
  filter(male=="T247")
mean(checking.real.T247$motif_count) #76.33333
var(checking.real.T247$motif_count) #30.33333

testing.bootstrap.T247<- boot_male(one_male="T247", data.from=filtered.data,boot_set=out)
mean(testing.bootstrap.T247$a_mean) #76.34833
mean(testing.bootstrap.T247$b_var) #20.04167
```

# Running Bootstrap Function
```{r}
simply_boot <- bind_rows(lapply(unique.male,boot_male, filtered.data, out))
simply_boot <- simply_boot %>% 
  group_by(male) %>%
  summarise(sd_boot_mean=sd(a_mean),
  sd_boot_var=sd(b_var),
  mean_boot_mean=mean(a_mean),
  mean_boot_var=mean(b_var))
```

# Bootgraph
```{r}
boot.graph<- simply_boot %>% 
  ggplot(aes(x=mean_boot_mean, y=mean_boot_var))+
  geom_point(color="black")+
    geom_errorbar(aes(ymin=mean_boot_var-sd_boot_var, ymax=mean_boot_var+sd_boot_var))+
    geom_errorbar(aes(xmin=mean_boot_mean-sd_boot_mean, xmax=mean_boot_mean+sd_boot_mean))+
  theme_minimal()+
  xlim(0, NA)+
  ylim(0, NA)+
  labs(x="Mean", y="Var", title="Bootstrapped Data")

boot.graph
```


### Which is the best fit model
```{r}
m1 <-lm(simply_boot$mean_boot_var~0+simply_boot$mean_boot_mean)
m2 <- lm(simply_boot$mean_boot_var~1+simply_boot$mean_boot_mean)
m3 <- lm(simply_boot$mean_boot_var~0+simply_boot$mean_boot_mean+ I(simply_boot$mean_boot_mean^2))
m4 <- lm(simply_boot$mean_boot_var~1+simply_boot$mean_boot_mean+ I(simply_boot$mean_boot_mean^2))
```
- lm(mean_boot_var ~ 0 + mean_boot_mean)
    - AIC= `r AIC(m1)`
- lm(mean_boot_var ~ 1 + mean_boot_mean)
    - AIC= `r AIC(m2)`
- **lm(mean_boot_var~ 0 + mean_boot_mean + I (mean_boot_mean^2))**
    - **AIC= `r AIC(m3)`**
- lm(mean_boot_var~ 1 + mean_boot_mean + I (mean_boot_mean\^2))
    - AIC= `r AIC(m4)`


# Deming
```{r}
obs.to.combine <- filtered.data %>% 
  group_by(male) %>% 
  summarise(observed_sd=sd(motif_count), 
    observed_mean=mean(motif_count), 
    observed_var=var(motif_count)) %>% 
  unique()
combined<- left_join(obs.to.combine, simply_boot, by = join_by(male))
x <- combined$observed_mean # Observed Mean for each male
y <- combined$observed_var  # Observed Var for each male
xstd <- combined$sd_boot_mean # Bootstrapped sd(Mean) for each male
ystd <- combined$sd_boot_var # Bootstrapped sd(Var) for each male

deming.original <- deming(y ~ x + 0,  # don't know if 0 + x + I(x^2) would work, Error in deming(y ~ x + I(x^2), xstd = xstd, ystd = ystd) : Deming regression requires a single predictor variable
        xstd=xstd,
        ystd=ystd
        )
a <- as.double(deming.original$coefficients[2])

z <- (0 + x + I(x^2))
deming.better<- deming(y~z ,
       xstd=xstd,
        ystd=ystd
        )
b<- as.double(deming.better$coefficients[2])
```
-   Deming doesn't seem to work with the quadratic slope. (Warning: NAs produced) All were NA, so switched to the other one.

# Simulate Function
```{r}
simulating.for.boot<- function(a.male,full.results){
    a_male_data <- combined %>%
        filter(male == a.male)
    for(i in 1:3){
          simulated[i] <- rnbinom(n=1, size = a_male_data$observed_mean/(a-1),prob=1/a)
          results <- tibble(male = a.male, simulated_motif= simulated)
    }
full.results<- rbind(results)
return(full.results)
}
```

# Running Simulation
```{r}
full.simulated.results <- bind_rows(lapply(unique.male, simulating.for.boot, out))
simplified_simulated<- full.simulated.results %>%
    group_by(male) %>%
    summarise(sim_sd=sd(simulated_motif),
              simulated_mean = mean(simulated_motif),
              simulated_var = var(simulated_motif))
simplified_simulated <- unique(simplified_simulated)
```

## Check all its models
```{r}
m1 <-lm(simply_simulated$simulated_var~0+simply_simulated$simulated_mean)
m2 <- lm(simply_simulated$simulated_var~1+simply_simulated$simulated_mean)
m3 <- lm(simply_simulated$simulated_var~0+simply_simulated$simulated_mean+ I(simply_simulated$simulated_mean^2))
m4 <- lm(simply_simulated$simulated_var~1+simply_simulated$simulated_mean+ I(simply_simulated$simulated_mean^2))

```

- lm(simulated_var ~ 0 + simulated_mean)
    - AIC= `r AIC(m1)` 
    
- lm(simulated_var ~ 1 + simulated_mean)
    - AIC= `r AIC(m2)` 
    
- **lm(mean_boot_var~ 0 + mean_boot_mean + I (mean_boot_mean^2))**
    - **AIC= `r AIC(m3)`**
    
- lm(simulated_var~ 1 + simulated_mean + I (simulated_mean\^2))
    - AIC= `r AIC(m4)`
    
#Bootstrap Simulated Data

```{r}
full.sim.boot <- boot_male(one_male=unique.male, data.from=full.simulated.results, boot_set=out)

simplified_simulated_boot<- full.sim.boot %>%
    group_by(male) %>%
  mutate(sd_boot_sim_mean=sd(a_mean)) %>%
  mutate(sd_boot_sim_var=sd(b_var)) %>%
  mutate(mean_boot_sim_mean=mean(a_mean)) %>%
  mutate(mean_boot_sim_var=mean(b_var)) %>%
  select(-a_mean, -b_var)
simply_simulated_boot <- unique(simplified_simulated_boot)
```

#Checking the models
```{r}
m1 <-lm(simply_simulated_boot$mean_boot_sim_var~0+simply_simulated_boot$mean_boot_sim_mean)
m2 <- lm(simply_simulated_boot$mean_boot_sim_var~1+simply_simulated_boot$mean_boot_sim_mean)
m3 <- lm(simply_simulated_boot$mean_boot_sim_var~0+simply_simulated_boot$mean_boot_sim_mean+ I(simply_simulated_boot$mean_boot_sim_mean^2))
m4 <- lm(simply_simulated_boot$mean_boot_sim_var~1+simply_simulated_boot$mean_boot_sim_mean+ I(simply_simulated_boot$mean_boot_sim_mean^2))

```
- lm(mean_boot_sim_var ~ 0 + mean_boot_sim_mean)
    - AIC= `r AIC(m1)` 
    
- lm(mean_boot_sim_var ~ 1 + mean_boot_sim_mean)
    - AIC= `r AIC(m2)` 
    
- **lm(mean_boot_sim_var~ 0 + mean_boot_sim_mean + I (mean_boot_sim_mean^2))**
    - **AIC= `r AIC(m3)`**
    
- lm(mean_boot_sim_var~ 1 + mean_boot_sim_mean + I (mean_boot_sim_mean\^2))
    - AIC= `r AIC(m4)`
    
#Stimulate 1000 Trials

```{r}
simulate.entire <- function(a.male,full.results){
    a_male_data <- combined %>%
        filter(male == a.male)
    for(i in 1:1000){
          simulated[i] <- rnbinom(n=1, size = a_male_data$observed_mean/(a-1),p=1/a)
          results <- tibble(male = a.male, simulated_motif= simulated)
    }
full.results<- rbind(results)
return(full.results)
}

simulate.entire.results <- bind_rows(lapply(unique.male,simulate.entire, out))


simplified_simulated.entire<- simulate.entire.results %>%
    group_by(male) %>%
    mutate(sim_sd=sd(simulated_motif)) %>%
    mutate(simulated_mean = mean(simulated_motif)) %>%
    mutate(simulated_var = var(simulated_motif)) %>%
    select(-simulated_motif)
simply_entire.simulated <- unique(simplified_simulated.entire)
```
### Check all its models

```
- m4 is now the best fit.
```

```{r}
m1 <-lm(simply_entire.simulated$simulated_var~0+simply_entire.simulated$simulated_mean)
m2 <- lm(simply_entire.simulated$simulated_var~1+simply_entire.simulated$simulated_mean)
m3 <- lm(simply_entire.simulated$simulated_var~0+simply_entire.simulated$simulated_mean+ I(simply_entire.simulated$simulated_mean^2))
m4 <- lm(simply_entire.simulated$simulated_var~1+simply_entire.simulated$simulated_mean+ I(simply_entire.simulated$simulated_mean^2))

```


- lm(simulated_var ~ 0 + simulated_mean)
    - AIC= `r AIC(m1)` 
    
- lm(simulated_var ~ 1 + simulated_mean)
    - AIC= `r AIC(m2)` 
    
- lm(simulated_var~ 0 + simulated_mean + I (simulated_mean^2))
    - AIC= `r AIC(m3)`
    
- **lm(simulated_var~ 1 + simulated_mean + I (simulated_mean\^2))**
    - **AIC= `r AIC(m4)`**


# Cleaning for graph
```{r}
simply_observed <- filtered.data %>%
    mutate(type="Observed Data", graph_order=1) %>%
    rename(mean=observed_mean, var=observed_var, sd_mean=observed_sd)

simply_boot <- simply_boot %>%
    mutate(type="Bootstrap of Observed Data", graph_order=2) %>%
    rename(mean=mean_boot_mean, var=mean_boot_var, sd_mean=sd_boot_mean, sd_var=sd_boot_var)

simply_simulated <- simply_simulated %>%
    mutate(type="Simulation of 9 Data Points", graph_order=3) %>%
    rename(mean=simulated_mean, var=simulated_var, sd_mean=sim_sd)

simply_simulated_boot <- simply_simulated_boot %>%
    mutate(type="Bootstrap of Simulated Data", graph_order=4) %>%
    rename(mean=mean_boot_sim_mean, var=mean_boot_sim_var, sd_mean=sd_boot_sim_mean, sd_var=sd_boot_sim_var)

simply_entire.simulated <- simply_entire.simulated %>%
    mutate(type="Simulation of 1000 Data Points", graph_order=5) %>%
    rename(mean=simulated_mean, var=simulated_var, sd_mean=sim_sd)

big.combination<- rbind(simply_observed, simply_boot, simply_simulated, simply_simulated_boot, simply_entire.simulated) %>%
    arrange(graph_order)
```


```{r}
observed.graph+ylim(0,21000)+boot.graph
```


# Graphing

```{r}
strip <- ggh4x::strip_themed(background_x = ggh4x::elem_list_rect(fill = c( "skyblue","skyblue", "slateblue","hotpink", "hotpink")))

big.combination%>%
    ggplot(aes(x=mean, y=var))+
    geom_point()+
    ggh4x::facet_wrap2(~type, strip = strip)+
    geom_errorbar(aes(ymin=var-sd_var, ymax=var+sd_var))+
    geom_errorbar(aes(xmin=mean-sd_mean, xmax=mean+sd_mean))+
    geom_smooth(method = "lm", formula = y ~ x, color="blue")+
    geom_smooth(method = "lm", formula = y ~ x + I(x^2), color="red")+
    labs(x="Mean Number of Motifs", y=str_wrap("Variance in Number of Motifs", width = 16), caption=str_wrap("Purple=Observed Pink=Simulated Blue=Bootstraped", width = 8))+
    ylim(0,NA)+
    xlim(0,NA)+
    theme_classic()
```
