---
title: "Gilchrist- Deming and Simulating Data"
author: "Tara Empson"
date: "`r Sys.Date()`"
output: html_document
---

# Goal

Create a linear model and run an anova from bootstrapped dataset from the mean number of songs from male zebra finch, repeatidly tested in 40C.

# Background

Kayci Messerly, a previous post-bac, preformed multiple tests of trying to see how male zebra finch song varies in heat. Being an animal behavior study there is a lot of variation and a very limited sample size. 

# Libraries

```{r message=FALSE}
library(tidyverse)
library(janitor)
library(deming)
library(patchwork)
```

# Bootstraping data

## Load Data & Define Parameter
- Not loading in T230 because he did not sing at all during this round and Deming/Simulating didn't like 0s. 
```{r}
repo.dir <- system("git rev-parse --show-toplevel", intern = TRUE)
setwd(file.path(repo.dir, "/code/Empson"))

data.dir <- file.path("..", "..", "data", "collated")
set.seed(1)
og.raw.data <- read.csv(file.path(data.dir, "HSPi-Repeatability-Song-Count.csv"))
raw.data <- og.raw.data %>% 
    filter(male!="T230") %>%
    group_by(male) %>%
    mutate(test_order = rank(date)) %>%
    ungroup()


unique.male <- unique(raw.data$male)
n.males <- length(unique.male)
n <- 3 ## trials per male
sample = 1000 # bootstrap samples
raw.data$day<- rep(1:n, n.males)
#adding in day order


col.name = "motif_count"
my_means <- vector(mode="double", length = sample)
my_vars <- my_means
out <- tibble()
simulated <- vector()
```


```{r}

select.out <- c("chamber",
                "date",
                "motif_count",
                "counter",
                "temp_target",
                "temp_mean",
                "temp_median",
                "humidity_mean",
                "test_order")

simplified_real <- raw.data %>% 
    group_by(male) %>% 
    mutate(real_sd=sd(motif_count)) %>% 
     mutate(real_mean=mean(motif_count)) %>% 
    mutate(real_var=var(motif_count)) %>% 
    select(!all_of(select.out))

simply_real <- unique(simplified_real)
```

### Bootstrap 
```{r}
boot_male <-function(one_male, boot_set){
  sample<-1000
  data <- raw.data %>%
    filter(male == one_male)
  for(i in 1:sample){
    bootstrap_sample <- sample(data$motif_count,n, replace = TRUE)
    my_means[i] <- mean(bootstrap_sample)
    my_vars[i] <- var(my_means)   
    results <- tibble(male = one_male, a_mean =  my_means, b_var = my_vars)
  }
boot_set<-  rbind(boot_set, results)
return(boot_set)
}

full.boot <- bind_rows(lapply(unique.male,boot_male, out))
```

##Simplify the bootstrap to the mean of the means per male and the sd of the means per male. 
```{r}
simplified_boot<- full.boot %>% 
    group_by(male) %>% 
  mutate(sd_boot_mean=sd(a_mean)) %>% 
  mutate(sd_boot_var=sd(b_var)) %>% 
  mutate(mean_boot_mean=mean(a_mean)) %>% 
  mutate(mean_boot_var=mean(b_var)) %>% 
  select(-a_mean, -b_var)

simply_boot <- unique(simplified_boot)
```

### Quadradic (m3) is the best fit model
```{r}
m1 <-lm(simply_boot$mean_boot_var~0+simply_boot$mean_boot_mean)
m2 <- lm(simply_boot$mean_boot_var~1+simply_boot$mean_boot_mean)
m3 <- lm(simply_boot$mean_boot_var~0+simply_boot$mean_boot_mean+ I(simply_boot$mean_boot_mean^2))
m4 <- lm(simply_boot$mean_boot_var~1+simply_boot$mean_boot_mean+ I(simply_boot$mean_boot_mean^2))
anova(m1)
step(m1) #m1 AIC=133.23

anova(m2)
step(m2) #m2 AIC=124.43

anova(m3)
step(m3) #m3 AIC=120.22

anova(m4)
step(m4) #m4 AIC=121.66

```

## Combine bootstraped data and real data into a dataset.
```{r}
combined<- left_join(simply_real, simply_boot, by = join_by(male))
```

# Deming
```{r}
x <- combined$real_mean # Observed Mean for each male
y <- combined$real_var  # Observed Var for each male
xstd <- combined$sd_boot_mean # Bootstrapped sd(Mean) for each male
ystd <- combined$sd_boot_var # Bootstrapped sd(Var) for each male

deming.original <- deming(y ~ x + 0,  # don't know if 0 + x + I(x^2) would work, Error in deming(y ~ x + I(x^2), xstd = xstd, ystd = ystd) : Deming regression requires a single predictor variable
        xstd=xstd,
        ystd=ystd
        )
a <- as.double(deming.original$coefficients[2])

z <- (0 + x + I(x^2))
deming.better<- deming(y~z , 
       xstd=xstd,
        ystd=ystd
        )
b<- as.double(deming.better$coefficients[2])
```

# Simulate Data
- Simulating doesn't seem to work with the quadradic slope. (Warning: NAs produced) All were NA, so switched to the other one. 
```{r}
simulating.for.boot<- function(a.male,full.results){
    a_male_data <- combined %>% 
        filter(male == a.male)
    for(i in 1:n){
          simulated[i] <- rnbinom(n=1, size = a_male_data$real_mean/(a-1),p=1/a)
          results <- tibble(male = a.male, simulated_motif= simulated)
    }
full.results<- rbind(results)
return(full.results)
}

full.simulated.results <- bind_rows(lapply(unique.male,simulating.for.boot, out))

 
simplified_simulated<- full.simulated.results %>% 
    group_by(male) %>% 
    mutate(sim_sd=sd(simulated_motif)) %>% 
    mutate(simulated_mean = mean(simulated_motif)) %>% 
    mutate(simulated_var = var(simulated_motif)) %>% 
    select(-simulated_motif)
simply_simulated <- unique(simplified_simulated)
```

### Check all its models
```{r}
m1 <-lm(simply_simulated$simulated_var~0+simply_simulated$simulated_mean)
m2 <- lm(simply_simulated$simulated_var~1+simply_simulated$simulated_mean)
m3 <- lm(simply_simulated$simulated_var~0+simply_simulated$simulated_mean+ I(simply_simulated$simulated_mean^2))
m4 <- lm(simply_simulated$simulated_var~1+simply_simulated$simulated_mean+ I(simply_simulated$simulated_mean^2))
anova(m1)
step(m1) #m1 AIC=145.35

anova(m2)
step(m2) #m2 AIC=146.49

anova(m3)
step(m3) #m3 AIC=144.39

anova(m4)
step(m4) #m4 AIC=145.61
```


#Bootstrap Simulated Data
```{r}
sim_boot_male <-function(one_male, boot_set){
  sample<-1000
  data <- full.simulated.results %>%
    filter(male == one_male)
  for(i in 1:sample){
    bootstrap_sample <- sample(data$simulated_motif,n, replace = TRUE)
    my_means[i] <- mean(bootstrap_sample)
    my_vars[i] <- var(my_means)   
    results <- tibble(male = one_male, a_mean =  my_means, b_var = my_vars)
  }
boot_set<-  rbind(boot_set, results)
return(boot_set)
}
full.sim.boot <- bind_rows(lapply(unique.male,sim_boot_male, out))

simplified_simulated_boot<- full.sim.boot %>% 
    group_by(male) %>% 
  mutate(sd_boot_sim_mean=sd(a_mean)) %>% 
  mutate(sd_boot_sim_var=sd(b_var)) %>% 
  mutate(mean_boot_sim_mean=mean(a_mean)) %>% 
  mutate(mean_boot_sim_var=mean(b_var)) %>% 
  select(-a_mean, -b_var)
simply_simulated_boot <- unique(simplified_simulated_boot)
```

#Checking the models
```{r}
m1 <-lm(simply_simulated_boot$mean_boot_sim_var~0+simply_simulated_boot$mean_boot_sim_mean)
m2 <- lm(simply_simulated_boot$mean_boot_sim_var~1+simply_simulated_boot$mean_boot_sim_mean)
m3 <- lm(simply_simulated_boot$mean_boot_sim_var~0+simply_simulated_boot$mean_boot_sim_mean+ I(simply_simulated_boot$mean_boot_sim_mean^2))
m4 <- lm(simply_simulated_boot$mean_boot_sim_var~1+simply_simulated_boot$mean_boot_sim_mean+ I(simply_simulated_boot$mean_boot_sim_mean^2))

anova(m1)
step(m1) #m1 AIC=142.49

anova(m2)
step(m2) #m2 AIC=135.12

anova(m3)
step(m3) #m3 AIC=104.19

anova(m4)
step(m4) #m4 AIC=105.35
```


#Stimulate 1000 Trials
```{r}
simulate.entire <- function(a.male,full.results){
    a_male_data <- combined %>% 
        filter(male == a.male)
    for(i in 1:1000){
          simulated[i] <- rnbinom(n=1, size = a_male_data$real_mean/(a-1),p=1/a)
          results <- tibble(male = a.male, simulated_motif= simulated)
    }
full.results<- rbind(results)
return(full.results)
}

simulate.entire.results <- bind_rows(lapply(unique.male,simulate.entire, out))

 
simplified_simulated.entire<- simulate.entire.results %>% 
    group_by(male) %>% 
    mutate(sim_sd=sd(simulated_motif)) %>% 
    mutate(simulated_mean = mean(simulated_motif)) %>% 
    mutate(simulated_var = var(simulated_motif)) %>% 
    select(-simulated_motif)
simply_entire.simulated <- unique(simplified_simulated.entire)
```


### Check all its models
    - m4 is now the best fit.
```{r}
m1 <-lm(simply_entire.simulated$simulated_var~0+simply_entire.simulated$simulated_mean)
m2 <- lm(simply_entire.simulated$simulated_var~1+simply_entire.simulated$simulated_mean)
m3 <- lm(simply_entire.simulated$simulated_var~0+simply_entire.simulated$simulated_mean+ I(simply_entire.simulated$simulated_mean^2))
m4 <- lm(simply_entire.simulated$simulated_var~1+simply_entire.simulated$simulated_mean+ I(simply_entire.simulated$simulated_mean^2))

anova(m1)
step(m1) #m1 AIC=88.92

anova(m2)
step(m2) #m2 AIC=90.3

anova(m3)
step(m3) #m3 AIC=86.12

anova(m4)
step(m4) #m4 AIC=79.9
```

# Graphing
```{r}
strip <- ggh4x::strip_themed(background_x = ggh4x::elem_list_rect(fill = c( "skyblue","skyblue", "slateblue","hotpink", "hotpink")))
simply_real <- simply_real %>% 
    mutate(type="Real Data", graph_order=1) %>% 
    mutate(mean=real_mean, var=real_var, sd_mean=real_sd) %>% 
    select(-real_sd, -real_mean, -real_var)
simply_boot <- simply_boot %>% 
    mutate(type="Bootstrap of Real Data", graph_order=2) %>% 
    mutate(mean=mean_boot_mean, var=mean_boot_var, sd_mean=sd_boot_mean, sd_var=sd_boot_var) %>% 
    select(-mean_boot_mean, -mean_boot_var, -sd_boot_mean, -sd_boot_var)
simply_simulated <- simply_simulated %>% 
    mutate(type="Simulation of Real Data", graph_order=3) %>% 
    mutate(mean=simulated_mean, var=simulated_var, sd_mean=sim_sd) %>% 
    select(-simulated_mean, -simulated_var, -sim_sd)
simply_simulated_boot <- simply_simulated_boot %>% 
    mutate(type="Bootstrap of Simulated Data", graph_order=4) %>% 
    mutate(mean=mean_boot_sim_mean, var=mean_boot_sim_var, sd_mean=sd_boot_sim_mean, sd_var=sd_boot_sim_var) %>% 
    select(-mean_boot_sim_mean, -mean_boot_sim_var, -sd_boot_sim_mean, -sd_boot_sim_var)
simply_entire.simulated <- simply_entire.simulated %>% 
    mutate(type="Simulating 1000 Trials for Each Bird", graph_order=5) %>% 
    mutate(mean=simulated_mean, var=simulated_var, sd_mean=sim_sd) %>% 
    select(-simulated_mean, -simulated_var, -sim_sd)

big.combination<- rbind(simply_real, simply_boot, simply_simulated, simply_simulated_boot, simply_entire.simulated) %>%
    arrange(graph_order)
big.combination%>% 
    ggplot(aes(x=mean, y=var))+
    geom_point()+
    ggh4x::facet_wrap2(~type, strip = strip)+
    geom_errorbar(aes(ymin=var-sd_var, ymax=var+sd_var))+
    geom_errorbar(aes(xmin=mean-sd_mean, xmax=mean+sd_mean))+
    geom_smooth(method = "lm", formula = y ~ x, color="blue")+
    geom_smooth(method = "lm", formula = y ~ x + I(x^2), color="red")+
    labs(x="Mean Number of Motifs", y=str_wrap("Variance in Number of Motifs", width = 16), caption=str_wrap("Purple=Real Pink=Simulated Blue=Bootstraped", width = 8))+
    ylim(0,NA)+
    xlim(0,NA)+
    theme_classic()
```

```{r}

```

