---
title: "Piecewise Regression QPoisson Error on Real Data using STAN Directly"
author: "Michael Gilchrist"
date: "Created on 2022-12-22; Code run on `r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=2E2, output = TRUE)

```
# Goal

- Fit two piece quasipoisson to data

## Recap

- My quasipoisson is actually not a quasipoisson model, but a constrained/modified negative binomial where the `shape = mu/(theta-1)` which ensures that `var(x| mu, shape) = theta mu`
- Limitation of this work is the lack of a male index.
  This prevents me from using male as a grouping variable for the data when I try and use 2 population level `y_0` or `x_0` values.
  I can, however, use individual parameters.
- I later began trying to add this capability, but then decided it would be a good excuse to try and use `brms`.
  Unfortunately, it's not obvious how to implement this behavior in `brms` and the standard Nbin regression approach does not have the desired relationship between the mean and variance.
- I either need to create my own `family` in `brms` or go back to my stan code.
-  It might be possible to use `brms` to generate the starting stan code.


# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(loo)


## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Load Data

```{r}
load(file.path("input", "data.processing_2022-12-15.Rda"),
     verbose = TRUE)


motif_data
```


# Examine Data

## Create Working Dataset

```{r}

filter_data <- TRUE

if(filter_data) {
  males_filtered_disp <- motif_stats_40C %>%
    filter(dispersion < 50) %>%
    pull(male)

  males_filtered_mean <- motif_stats %>%
    filter(mean > 10) %>%
    pull(male)

  male_vector <- intersect(males_filtered_mean, males_filtered_disp)
} else {
  male_vector <- motif_data %>% select(male) %>% distinct()
}

data_ind <- motif_data %>%
    filter(male %in% male_vector) %>%
    mutate(male = droplevels(male)) %>%
    mutate(id = factor(male)) %>% 
    mutate(male = as.integer(male)) %>%
    arrange(male) %>%
    select(male, id, motif_count, temp, round, trial_round, date, counter) %>% 
  ##    left_join(male_shape, by = "male") %>%
  mutate()

stats_ind <- motif_stats %>%
    filter(male %in% male_vector) 


data_ind <- data_ind %>% filter(temp < 38) %>%
    group_by(male) %>% mutate(y0_simple.est = mean(motif_count), phi_ind = var(motif_count)/y0_simple.est) ## phi is overdispersion parameter

summary(data_ind)

summary_stats <- data_ind %>% ungroup() %>%  summarize(y0_bar = mean(y0_simple.est), y0_sd = sd(y0_simple.est), phi_bar = median(phi_ind))

n_male <- length(unique(data_ind$male))

```


# Analyze Real Data:


## Set Up Data

```{r}

data <- data_ind

motif_count <- data %>% pull(motif_count)
temp <- data %>% pull(temp)
N <- length(temp)
male <- data %>% pull(male)
## parameters to be printed

pars <- c("t0", "y0")
pars_full <- c(pars, "lp__")
```

## Assume QPoisson
 
### Males pooled: `y0`, `x0`, and $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson.stan")

stan_model(file="two.piece_qpoisson.stan")

model <- stan(file = "two.piece_qpoisson.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = 1,
                        M = 1,
                        ll = rep(1,N),
                        mm = rep(1,N),                        
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 100,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)

fit_qpoisson_pooled.x0_pooled_y0_pooled_theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")

title <- "Two piece Pooled Males"
print(title)
quietgg(print(model, pars = pars))
#traceplot(model, pars = pars, inc_warmup = FALSE)

quietgg(plot(model, pars = pars, point_est = "mean")) #, ggtitle(title)) 
quietgg(pairs(model, pars = pars_full))
quietgg(stan_hist(model))

summary(model, probs = c(0.025, 0.5, 0.975))$summary


```
### Separate: `y0`, and `x0`, Pooled  $\theta$

```{r}

iter <- 20000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

     
model <- stan(file = "two.piece_qpoisson.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = n_male,
                        M = 1,
                        ll = data$male, ## y0 and x0
                        mm = rep(1, N), ## Theta
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 10,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter*3/4),
              thin = 2,
              verbose = TRUE)
fit_qpoisson_separate.x0_separate.y0_pooled.theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")


print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title)) 

#quietgg(stan_hist(model, pars = "x0"))

## Plot sampling for each male
tmp_plot <- list()
for(i in list(1:3, 4:6, 7:10, 11)) {
    tmp_pars <-  c(paste0("x0[", i, "]"), paste0("y0[", i, "]"), "theta")
    tmp_plot[[as.character(i)]] <- quietgg(pairs(model, pars = tmp_pars))
}

# Doesn't work.  I believe pairs() might be returning null
#gt <- arrangeGrob(grobs = tmp_plot)
#as_ggplot(gt)

## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- quietgg(stan_plot(model, pars = par))
}

gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

## Doesn't work as expected
## tmp_plots <- stan_plot(model, pars = pars)
## tmp_plots +
##    facet_wrap(vars(grepl("x0.*", params)),
##               scales = "free")

                                        #quietgg(pairs(model, pars = pars_full))
quietgg(stan_hist(model, pars = pars))

               
# summary(fit_qpoisson.separate.y0.x0, probs = c(0.025, 0.5, 0.975))$summary

```
- Histograms of `x0[]` suggest following groupings
  1. `c(2, 4, 6, 8, 11)`
  2. `c(1, 3, 5, 7, 10)`
  3. `9` is ambiguous.

### Separate: `y0`, `x0`, and  $\theta$

- didn't run given noise in previous one with separate `y0` and `x0`


### Separate: `y0`, Shared `x0`, and  $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
 cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")

stan_model(file="two.piece_qpoisson_2.0.stan")

model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = 1,
                        Y = n_male,
                        NB = 1,
                        xx = rep(1,N),
                        yy = data$male,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_pooled.x0_separate.y0_pooled.theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")


print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title)) 

## Plot sampling for each male
tmp_plot <- list()
for(i in 1:n_male) {
    tmp_pars <-  c(paste0("x0[", 1, "]"), paste0("y0[", i, "]"), "lp__")
    tmp_plot[[as.character(i)]] <- quietgg(pairs(model, pars = tmp_pars))
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)


## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```
- `y0[7]` should probably be removed from dataset
- Possible groupings for `y0[i]`
  1. $i = 1, 5, 6, (7), 10, 11$
  2. $i = 2, 3, 4, 9$
  3. $i = 8$ could possibly be in either; 2. seems most likely.
  
### Two groups `y0`,  Shared `x0` and Shared $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
#cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")
#stan_model(file="two.piece_qpoisson_2.0.stan")

y0_group <- list()
y0_group[[1]] <- c(1, 5, 6, 7, 10, 11)
y0_group[[2]] <- c(2, 3, 4, 8, 9)

y0_grouping <- map_int(data$male, ~ if_else(. %in% y0_group[[1]], 1, 2))
                     
model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = 1,
                        Y = 2,
                        NB = 1,
                        xx = rep(1,N),
                        yy = y0_grouping,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_pooled.x0_grouped.y0_pooled.theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")


print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
plot(model, pars = pars_full) #, ggtitle(title)) 

quietgg(pairs(model, pars = pars_full))

## Plot sampling for each male
tmp_plot <- list()
## Plot sampling for each male

for(i in 1:2) {
    tmp_pars <-  c(paste0("x0[", y0_group[[i]], "]"),
                   paste0("y0[", i, "]"),
                   "lp__")
    tmp_plot[[as.character(i)]] <- quietgg(pairs(model, pars = tmp_pars))
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)


## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```


### Separate `y0`,  Grouped `x0` and Shared $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
#cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")
#stan_model(file="two.piece_qpoisson_2.0.stan")

x0_group <- list()
x0_group[[1]] <- c(2, 4, 6, 8, 11, 9)
x0_group[[2]] <- c(1, 3, 5, 7, 10)

x0_grouping <- map_int(data$male, ~ if_else(. %in% x0_group[[1]], 1, 2))
                     
model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = 2,
                        Y = n_male,
                        NB = 1,
                        xx = x0_grouping,
                        yy = data$male,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_grouped.x0_separate.y0_pooled.theta <- model
#model <- fit_qpoisson.grouped.y0.and.pooled.x0.theta

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")

print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title))

## Plot sampling for each male
tmp_plot <- list()
for(i in 1:2) {
    tmp_pars <-  c(paste0("x0[", i, "]"), paste0("y0[", x0_group[[i]], "]"),  "lp__")
    tmp_plot[[as.character(i)]] <- quietgg(pairs(model, pars = tmp_pars))
}

gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)


## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```

- 

### Grouped `y0`,  Grouped `x0` and Shared $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
#cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")
#stan_model(file="two.piece_qpoisson_2.0.stan")

x0_group <- list()
x0_group[[1]] <- c(2, 4, 6, 8, 11, 9)
x0_group[[2]] <- c(1, 3, 5, 7, 10)
x0_grouping <- map_int(data$male, ~ if_else(. %in% x0_group[[1]], 1, 2))

y0_group <- list()
y0_group[[1]] <- c(1, 5, 6, 7, 10, 11)
y0_group[[2]] <- c(2, 3, 4, 8, 9)
y0_grouping <- map_int(data$male, ~ if_else(. %in% y0_group[[1]], 1, 2))

model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = length(x0_group),
                        Y = length(y0_group),
                        NB = 1,
                        xx = x0_grouping,
                        yy = y0_grouping,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_grouped.x0_grouped.y0_pooled.theta <- model
#model <- fit_qpoisson.grouped.y0.and.pooled.x0.theta

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")

print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title))

quietgg(pairs(model, pars = pars_full))

## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```
- Results suggest clustering of `x0[]` is justified 

## Model Comparison via LOO

- TODO!

```{r}

ls("fit_qpoisson*")

``
