---
title: "Visualize Piecewise Regression with Negative Binomial Type I Error on Real Data using `brms` Custom Family"
author: "Michael Gilchrist"
date: "Created: 2023-03-09; Compiled: `r date()`"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)


if(interactive()) {
  default::default(.ess.eval) <- list(max.deparse.length=2E2, output = TRUE)
  output_dir = "output"  
} else { 
  output_dir = "output/render"  
}

```
# Goal

- Visualize two piece negative binomial type 1 formulation to data

## Recap


## Insights


# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(ggExtra)
library(cowplot)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(brms)
library(bayesplot)
library(tidybayes)
library(loo)
library(modelr)

## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(latex2exp)

```

## Source family

```{r}

source("../../../custom-brms-families/families/nbinom_type1.R")

## Define fmax() om case expose_functions()

fmax <- function(x, x0) max(x, x0)

```
## Load custom family


## Load All Files in `input`

```{r}

load(file.path("input", "fit_tbl.Rda"), verbose = TRUE)


```

# Visualize Model Fits


## Set up functions, parameters, and results tibble 

```{r}

xmax <- 46

## Here I create a variable using a string and then access it using a string
for(name in names(fit_tbl)[1:5]) {
  # create
  var <- paste0(name, "_vec")
  assign(var, pull(fit_tbl, name) %>% unique(.))
  # access via get()
  print(paste0(var, ": ", paste0(get(var), collapse = ", ")))
}

flags_y0 <- y0_flag_vec
flags_x0 <- x0_flag_vec 
flags_s0 <- flags_x0


```

### Print fits

```{r, message = FALSE }

sampling = "nbinom_type1"

flags_x0_used <- c("individual")
flags_y0_used <- c("individual")
disp_prior_used <- disp_prior_vec[6 ]
models_used <- c("x0_piecewise") #, "x0_piecewise")

print_mcmc_pairs <- FALSE 
print_mcmc_scatter <- TRUE
print_stan_hist <- FALSE

fit_index <- 0;

for(model in models_used) {  

  for(x0_flag in flags_x0_used) {
    s0_flag <- x0_flag ## define synonym
    
    for(y0_flag in flags_y0_used) {

      print_get_prior <- TRUE ## reset value
      print_prior_summary <- TRUE
      
      for(disp_prior in disp_prior_used) {
        
        desc_short <- paste0("x0 " , x0_flag, "; y0 ", y0_flag, "; disp prior ", disp_prior) 
        desc <- paste0(sampling, "; ", model, "; ", desc_short)
        filename_desc <- gsub("_", "-", desc) %>%
          gsub("; ", "_", .) %>%
          gsub(" ", "-", .)
        
          
        curr_row <- which(fit_tbl$model == model &
                            fit_tbl$sampling_dist == sampling &
                            fit_tbl$x0_flag == x0_flag &
                            fit_tbl$y0_flag == y0_flag &
                            fit_tbl$disp_prior == disp_prior)
        print(desc)

        print("Working with Pre-existing Fits")

        
        ## Try to assign from local memory.
        fit_brms <- fit_tbl[[curr_row, "fit"]][[1]]

        if(is.na(list(fit_brms))) {
          warning(paste0("model fit_brms ", desc, "does not exist.\n Skipping."))
        } else {
          ## Print and plot results, regardless of which fits one uses           
          print(desc)
          if(print_prior_summary) {
            print("Fit Prior Information")
            print(prior_summary(fit_brms)) # %>% filter(nlpar!="y0"))
            print_prior_summary <- TRUE
          }
          print("Fit Information")
          print(summary(fit_brms)) #, pars = "x0*"))          %>% filter(nlpar!="y0"))
          fit_stan <- fit_brms$fit
          #clean up variable names
          fit_stan_rename <-
            fit_stan %>%
            setNames(gsub("b_", "", names(.)) %>%
                       gsub("(x0|s0|y0)_male(T[0-9]{3})", "\\2_\\1", .) %>%
                       gsub("__", "_", .) %>%
                       gsub("r_male_(x0|s0|y0)\\[(T[0-9]{3}),Intercept\\]", "\\2_\\1_r", .) %>%
                       gsub("\\.", " ", .))

          ## 

          vars_clean <- names(fit_stan_rename) %>% na.omit(.)
          male_vec <- unique(fit_stan$data$male) %>% as.character(.)
          ## get male specific vars (start with "T")
          vars_T <- grep("^T[0-9]{3}", vars_clean, value = TRUE)
          vars_Intercept <- grep("Intercept", vars_clean, value = TRUE)
          vars_non_T <- vars_clean[!(vars_clean %in% c(vars_T, vars_Intercept))]

          
          ## Examine priors and exclude any constants from vars_non_T
          priors_tmp <- prior_summary(fit_brms)
          constants_non_T <- ggdist::parse_dist(priors_tmp) %>% filter(grepl("constant", .dist)) %>% pull(class)
          vars_non_T <- vars_non_T[!(vars_non_T %in% constants_non_T)]
          
          
          ## Count occurence of each male in model fit_brms
          male_instance <- sapply(male_vec, function(x) {sum(str_detect(x, string=vars_clean))})

          if(all(unique(male_instance) == 2)) {
            
            if(print_mcmc_pairs) {
              ## Use mcmc_pairs
              pairs_list <- list()            
              ## Update font size for grid
              ggplot2::theme_set(theme_default(base_size = 10))
              ## update panel.boarder
              ## Note brms uses 'variable' while stanfit uses 'pars'
              for(male in male_vec) {
                print(male)
                vars_male <- grep(male, vars_T, value = TRUE)

                ## Generate pairs() plot using mcmc_pairs()
                ## Don't use pairs.stanfit which does not return an object
                pairs_tmp <-  mcmc_pairs(
                  fit_stan_rename,
                  pars = vars_male,
                  ## set point size and transparency
                  off_diag_args = list(size = 1, alpha = 0.35)
                )
                # list of 4 ggplot objects, diagonal plots are 2 and 3
                plots_tmp <- pairs_tmp$bayesplots 
                
                # add density curves
                plots_tmp[[2]] <- plots_tmp[[2]] + stat_density_2d(color = "lightgray", size = .5)
                # What is plot 3 anyway?
                plots_tmp[[3]] <- plots_tmp[[3]] + stat_density_2d(color = "lightgray", size = .5)
                # recreate the grid
                pairs_tmp <-bayesplot_grid(plots = plots_tmp)
                pairs_list[[male]] <- pairs_tmp

              }
              if(length(vars_Intercept) > 0) {
                pairs_list[["Intercept"]] <- mcmc_pairs(
                  fit_stan_rename, pars = vars_Intercept)
              }
              pairs_list[["non-T"]] <- stan_hist(fit_stan_rename, pars = vars_non_T)

              ## Restore theme settings
              ggplot2::theme_set(theme_default())


              cowplot::plot_grid(plotlist = pairs_list,
                                 ncol = 3)
            }

            if(print_mcmc_scatter) {

              ggplot2::theme_set(theme_default(base_size = 10))
              scatter_list <- list()
              for(male in male_vec) {
                print(male)
                vars_male <- grep(male, vars_T, value = TRUE)
                
                scatter_tmp <-  mcmc_scatter(
                  fit_stan_rename,
                  pars = vars_male,
                  size = 1, alpha = 0.35
                )  +
                  stat_density_2d(color = "lightgray", size = .5, bins = 5)

                scatter_list[[male]] <- ggExtra::ggMarginal(scatter_tmp, type = "histogram")
                
              }
              scatter_list[["non-T"]] <- stan_hist(fit_stan_rename, pars = vars_non_T)

              cowplot::plot_grid(plotlist = scatter_list,
                                 ncol = 3)
              ## Restore theme settings
              ggplot2::theme_set(theme_default())
            }
          }
          
          if(print_stan_hist) {
            ncol <- 4 
            hist <-  stan_hist(fit_stan_rename,
                               pars = vars_fit,
                               bins = 25,
                               ncol = ncol) +
              ggtitle(desc_short)
            print(hist)
            filename <- paste0("histogram_", filename_desc, ".pdf")
            ggsave(filename = filename, path = file.path(output_dir, "figures"), dpi=300)
          }
        }
      }
    }
  }
}

```

# Plot Data and Predictions

Data suggests that disp should vary between males

```{r}

curr_row <- 181 
curr_desc <- fit_tbl[[curr_row, "desc"]]
fit_brms <- fit_tbl[[curr_row, "fit"]][[1]]

expose_functions(fit_brms, vectorize = TRUE, show_compiler_warnings=FALSE)
data_fit_brms <- fit_brms$data


## Create grid of x values for epred/predictions

data_grid <- data_fit_brms %>%
  group_by(male) %>%
  data_grid(x = seq_range(c(20, 40), n = 51))


## add expected values
data_epred <- data_grid %>%
  add_epred_draws(object = fit_brms, ndraws = 1000)

## Running into issues with predictions
## Only returns 1 draw per value in data_grid
## I believe this may have to do with the generation of NaN values
## HOwever, avoiding NaN by limiting x range doesn't seem to work.

## add predicted values
tmp_pp <- posterior_predict(fit_brms, newdata = data_grid)

## Using tidybayes
## DOesn't work as expected

## this seems to remove the .xy
data_pred <- add_predicted_draws(object = fit_brms, newdata = data_grid) # %>% filter(male == "T257"))

data_y <- data_fit_brms$y

## Plot all results
## Gets a bit crowded and takes a while; best to skip

if(FALSE) {
  ggplot(data = data_epred, aes(x = x, y = data_y, color = factor(male))) +
## Combine Scatter Plots and Model vs Data Plots
    stat_lineribbon(aes(y = .epred)) +
    geom_point(data = data_fit_brms) +
    scale_fill_brewer(palette = "Greys") +
    scale_color_brewer(palette = "Set2") +
    ylim(0, max(tmp_y)*1.1)
}
```

## Create Model vs. Data Plots

```{r}

plot_model <- list()

list_plot_mean <- list()
  
for(male in male_vec) {
    
  tmp_data <- data_fit_brms %>% filter(male == !!male)
  tmp_y <- tmp_data$y
  tmp_mean <- data_epred %>% filter(male == !!male)

  tmp_plot_mean <- ggplot(data = tmp_mean,
                            aes(x = x, y = tmp_y)) +
    ylim(0, max(tmp_y)*1.1) +
    stat_lineribbon(aes(y = .epred)) +
    scale_fill_brewer(palette = "Greys") +
    scale_color_brewer(palette = "Set2") +
    geom_point(data = tmp_data, aes(x = x, y = y), color = "Red") +
    labs(title=male) +
    theme(legend.position="none")
  #print(tmp)
  list_plot_mean[[male]] <- tmp_plot_mean
  
}

tmp_plot <- cowplot::plot_grid(
  plotlist = list_plot_mean,
  ncol = 4,
  title = "mean"
)

data_grid <- data_fit_brms %>%
  group_by(male) %>%
  data_grid(x = seq_range(c(20, 40), n = 51))

plot_model[["mean"]] <- tmp_plot

tmp_plot

```

```{r}

list_plot_data <- list()

for(male in male_vec) {


  tmp_data <- data_fit_brms %>% filter(male == !!male)
  tmp_y <- tmp_data$y
  tmp_pred  <- data_pred %>% filter(male ==!!male)
  tmp_plot_data <- ggplot(data = tmp_pred,
                          aes(x = x, y = tmp_y)) +
    ylim(0, max(tmp_y)*1.1) +
    stat_lineribbon(aes(y = .prediction),
                    .width = c(.95, .80, .50),
                    alpha = 1/4) +
    geom_point(data = tmp_data, aes(x = x, y = y), color = "Red") +
    scale_fill_brewer(palette = "Greys") +
    #scale_color_brewer(palette = "Set2") +
    labs(title=male) +
    theme(legend.position="none")
  print(tmp_plot_data)
  list_plot_data[[male]] <- tmp_plot_data
}

  
  #gridExtra::grid.arrange(grobs = plot_model_list)
  #
  #plot_tmp <- cowplot::plot_grid(plotlist = plot_model_list,
  #                  ncol = 3)


cowplot::plot_grid(
  plotlist = list_plot_data,
  ncol = 4,
  title = "data"
)


plot_model[["data"]] <- tmp_plot_data




```

## Combine Scatter Plots and Model vs Data Plots

```{r}

plot_combo <- list()

for(prediction in c("data")) {

    list_tmp <- list()
    for(male in male_vec) {
      list_tmp[[male]] <-
        cowplot::plot_grid(
          plotlist = list(
            scatter_list[[male]],
            plot_model[[prediction]][[male]] +
              labs(title = NULL) +
              xlab(NULL) +
              ylab(NULL)
          )
        )  + labs(title = male)
    }

  plot_tmp <- cowplot::plot_grid(
      plotlist = list_tmp,
      ncol = 2,
      title = prediction
    )
    print(plot_tmp)
    plot_combo[[prediction]] <- plot_tmp

}

```

## Exit rendering
```{r}

knitr::knit_exit()

```
