---
title: "Piecewise Regression on Simulated and Observed Data using STAN Directly"
author: "Michael Gilchrist"
date: "date: 2022-12-12"
output: pdf_document
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=1E3, output = TRUE)

```
# Goal

- Fit two piece Negative Binomial 

## Recap

# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(loo)


## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Load Data

```{r}
load(file.path("input", "data.processing_2022-11-09.Rda"),
     verbose = TRUE)

```


# Examine Data

## Create Working Dataset

```{r}

filter_data <- TRUE

if(filter_data) {
  males_filtered_disp <- song_stats_40C %>%
    filter(dispersion < 50) %>%
    pull(male)

  males_filtered_mean <- song_stats %>%
    filter(mean > 10) %>%
    pull(male)

  male_vector <- intersect(males_filtered_mean, males_filtered_disp)
} else {
  male_vector <- song_data %>% select(male) %>% distinct()
}


##males_selected <-

data_ind <- song_data %>%
  filter(male %in% male_vector) %>%
    arrange(male) %>%
    select(male, song_count, temp, round, trial_round, date, counter) %>% 
  ##    left_join(male_shape, by = "male") %>%
  mutate()

stats_ind <- song_stats %>%
  filter(male %in% male_vector)

data_ind <- data_ind %>% filter(temp < 38) %>%
    group_by(male) %>% mutate(y0_obs = mean(song_count), phi_ind = var(song_count)/y0_obs)

summary(data_ind)

summary_stats <- data_ind %>% ungroup() %>%  summarize(y0_bar = mean(y0_obs), y0_sd = sd(y0_obs), phi_bar = median(phi_ind))
```


# Simulate Data

From `rnbinom` help
  >  An alternative parametrization (often used in ecology) is by the
  >   _mean_ ‘mu’ (see above), and ‘size’, the _dispersion parameter_,
  >   where ‘prob’ = ‘size/(size+mu)’.  The variance is ‘mu + mu^2/size’
  >   in this parametrization.

```{r}
sim_counts <- function(temp, y0, tau0, tmax, phi) {

    ## Calculate expected value given parametes
    mu <- ifelse(temp < tau0, y0, y0 * (1-(temp-tau0)/(tmax-tau0))) 
    
    ## calculate size parameter based on mu and phi
    size = mu/(phi - 1)
    #print(paste("size: ", size, "count: ", count))
    rnbinom(1, size = size, mu = mu)
    }


phi <- summary_stats$phi_bar ## overdispersion parameter
tmax <- 46

data_tmp <- data_ind[, c("male", "temp", "y0_obs")] %>%
    mutate(male = droplevels(male)) %>%
    mutate(male = as.integer(male))


male_tmp <- data_tmp$male %>% unique()
n_male <- male_tmp  %>% length()
n_male


## Generate 'true' parameters
## If TRUE, replace estimated y_0 with simulated values,
## else use estimates from observed data 
if(FALSE) {
    set.seed(2022)
    y0_sim <- rnorm(n_male, summary_stats$y0_bar, summary_stats$y0_sd) %>% round(0)
    data_tmp <- data_tmp %>% mutate(y0 = y0_sim[.data$male])
} else {
    data_tmp <- data_tmp %>% mutate(y0 = y0_obs)
}

## made these tau0 values up
set.seed(2021)
tau0 <- rnorm(n_male, 39, 2) %>% round(1)
## Match y0 and tau0 to males
data_tmp <- data_tmp %>% mutate(tau0 = tau0[.data$male])
    
data_sim <- data_tmp %>% rowwise() %>% mutate(song_count = sim_counts(temp, y0, tau0, tmax = tmax, phi = phi))


## t_target <- c(30, 35, 38, 40, 42, 44)


```

# Analyze Data:


# Stan


## Set Up Data

```{r}

data <- data_sim

song_count <- data %>% pull(song_count)
temp <- data %>% pull(temp)
N <- length(temp)
male <- data %>% pull(male)
## parameters to be printed

pars <- c("t0", "y0")
pars_full <- c(pars, "b0", "lp__")
```
## Two Piece Model


### All Males Pooled

#### Original Code: Can't be used to pool males

```{r}

iter <- 5000
tmax <- 46
t0max <- tmax - 0.75;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

model <- stan(file = "two.piece_poisson.stan",
              model_name = "Two piece poisson",
              data=list(t = temp,
                        y = song_count,
                        N = N,
                        tmax = tmax,
                        t0min = t0min,
                        t0max = t0max,
                        ## max threshold value.
                        ## having it too close to tmax *sometimes* leads to sampling
                        ## near tmax, but with lower lp and very high E13) b0 values
                        y_tmax = 0,
                        tp = tp),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2))


## Examine output
title <- "Two piece"
print(title)
print(model, pars = pars)
traceplot(model, pars = pars, inc_warmup = FALSE, title = title)
plot(model, pars = pars) #, ggtitle(title)) 
pairs(model, pars = pars_full, title = title)

fit_poisson <- model

#ggplot()
```



#### Updated Code: Males can be pooled but aren't.

```{r}

iter <- 5000
tmax <- 46
t0max <- tmax - 0.75;
t0min <- 25;
## values to use for model predictions
tp = seq(25, xmax, length.out = 100)

male <-

    
model <- stan(file = "two.piece_poisson.grouped.by.male.stan",
              model_name = "Two piece poisson pooled",
              data=list(x = temp,
                        y = song_count,
                        N = N,
                        L = 1,
                        ll = rep(1,N),
                        xmax = tmax,
                        x0min = t0min,
                        x0max = t0max,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)


```

### Males Not Pooled
```{r}

iter <- 5000
tmax <- 46
t0max <- tmax - 0.75;
t0min <- 25;
## values to use for model predictions
tp = seq(25, xmax, length.out = 100)

male <-

    
model <- stan(file = "two.piece_poisson.grouped.by.male.stan",
              model_name = "Two piece poisson pooled",
              data=list(x = temp,
                        y = song_count,
                        N = N,
                        L = n_male,
                        ll = male,
                        xmax = tmax,
                        x0min = t0min,
                        x0max = t0max,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)


```
### Leave One Out (LOO) analysis


```{r}
## Code from:
# Extract pointwise log-likelihood
# using merge_chains=FALSE returns an array, which is easier to 
# use with relative_eff()
for(fit in c("fit_1", "fit_p")) {
  print(paste0("Model Fit: ", fit))  
  log_lik_1 <- extract_log_lik(eval(parse(text = fit)), merge_chains = FALSE)
# as of loo v2.0.0 we can optionally provide relative effective sample sizes
# when calling loo, which allows for better estimates of the PSIS effective
# sample sizes and Monte Carlo error
r_eff <- relative_eff(exp(log_lik_1), cores = 2) 

# preferably use more than 2 cores (as many cores as possible)
# will use value of 'mc.cores' option if cores is not specified
loo_1 <- loo(log_lik_1, r_eff = r_eff, cores = 2)
  print(loo_1)
}

```
- Don't really know how to interpret these results.


