---
title: "Piecewise Regression on Simulated qpoisson data using STAN Directly"
author: "Michael Gilchrist"
date: "date: 2022-12-12"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=1E3, output = TRUE)

```
# Goal

- Fit two piece poisson, quasipoisson, and negative binomial 

## Recap

# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(loo)


## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Load Data

```{r}
load(file.path("input", "data.processing_2022-12-15.Rda"),
     verbose = TRUE)

```


# Examine Data

## Create Working Dataset

```{r}

filter_data <- TRUE

if(filter_data) {
  males_filtered_disp <- motif_stats_40C %>%
    filter(dispersion < 50) %>%
    pull(male)

  males_filtered_mean <- motif_stats %>%
    filter(mean > 10) %>%
    pull(male)

  male_vector <- intersect(males_filtered_mean, males_filtered_disp)
} else {
  male_vector <- motif_data %>% select(male) %>% distinct()
}


##males_selected <-

data_ind <- motif_data %>%
  filter(male %in% male_vector) %>%
    arrange(male) %>%
    select(male, motif_count, temp, round, trial_round, date, counter) %>% 
  ##    left_join(male_shape, by = "male") %>%
  mutate()

stats_ind <- motif_stats %>%
  filter(male %in% male_vector)

data_ind <- data_ind %>% filter(temp < 38) %>%
    group_by(male) %>% mutate(y0_simple.est = mean(motif_count), phi_ind = var(motif_count)/y0_simple.est) ## phi is overdispersion parameter

summary(data_ind)

summary_stats <- data_ind %>% ungroup() %>%  summarize(y0_bar = mean(y0_simple.est), y0_sd = sd(y0_simple.est), phi_bar = median(phi_ind))
```


# Simulate Data

From `rnbinom` help
  >  An alternative parametrization (often used in ecology) is by the
  >   _mean_ ‘mu’ (see above), and ‘size’, the _dispersion parameter_,
  >   where ‘prob’ = ‘size/(size+mu)’.  The variance is ‘mu + mu^2/size’
  >   in this parametrization.

```{r}
sim_counts <- function(temp, y0, x0, tmax, phi) {

    ## Calculate expected value given parametes
    mu <- ifelse(temp < x0, y0, y0 * (1-(temp-x0)/(tmax-x0))) 
    
    ## calculate size parameter based on mu and phi
    size = mu/(phi - 1)
    #print(paste("size: ", size, "count: ", count))
    rnbinom(1, size = size, mu = mu)
    }


phi <- summary_stats$phi_bar ## overdispersion parameter
tmax <- 46

data_tmp <- data_ind[, c("male", "temp", "y0_simple.est")] %>%
    mutate(male = droplevels(male)) %>%
    mutate(male = as.integer(male))


male_tmp <- data_tmp$male %>% unique()
n_male <- male_tmp  %>% length()
n_male


## Generate 'true' parameters
## If TRUE, replace estimated y_0 with simulated values,
## else use estimates from observed data 
if(FALSE) {
    set.seed(2022)
    y0_sim <- rnorm(n_male, summary_stats$y0_bar, summary_stats$y0_sd) %>% round(0)
    data_tmp <- data_tmp %>% mutate(y0 = y0_sim[.data$male])
} else {
    data_tmp <- data_tmp %>% mutate(y0 = y0_simple.est)
}

## made these x0 values up
set.seed(2021)
x0 <- rnorm(n_male, 39, 2) %>% round(1)
## Match y0 and x0 to males
data_tmp <- data_tmp %>% mutate(x0 = x0[.data$male])
    
data_sim <- data_tmp %>% rowwise() %>% mutate(motif_count = sim_counts(temp, y0, x0, tmax = tmax, phi = phi))


## t_target <- c(30, 35, 38, 40, 42, 44)


```

# Analyze Simulated Data:


## Set Up Data

```{r}

data <- data_sim

motif_count <- data %>% pull(motif_count)
temp <- data %>% pull(temp)
N <- length(temp)
male <- data %>% pull(male)
## parameters to be printed

pars <- c("t0", "y0")
pars_full <- c(pars, "lp__")
```

## Assume (incorrectly) Poisson Error


### All Males Pooled

#### Original Stan Code: Can't be used to pool males

- `t0` = threshold parameter (later `x0`)
- `y0` = pre-threshold response rate


```{r}

iter <- 5000
tmax <- 46
t0max <- tmax - 0.75;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

model <- stan(file = "two.piece_poisson.stan",
              model_name = "Two piece poisson",
              data=list(t = temp,
                        y = motif_count,
                        N = N,
                        tmax = tmax,
                        t0min = t0min,
                        t0max = t0max,
                        ## max threshold value.
                        ## having it too close to tmax *sometimes* leads to sampling
                        ## near tmax, but with lower lp and very high E13) b0 values
                        y_tmax = 0,
                        tp = tp),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2))
fit_poisson <- model


## Examine output
title <- "Two piece, pooled males"
print(title)
print(model, pars = pars_full)
traceplot(model, pars = pars, inc_warmup = FALSE)
plot(model, pars = pars) #, ggtitle(title)) 
pairs(model, pars = pars_full)
summary(fit_poisson)$summary
#ggplot()
```



#### Updated Stan Code: Can be used to assign males to separate groups


```{r}

iter <- 5000
tmax <- 46
t0max <- tmax - 0.75;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

model <- stan(file = "two.piece_poisson.grouped.by.male.stan",
              model_name = "Two piece poisson pooled",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = 1,
                        ll = rep(1,N),
                        xmax = tmax,
                        x0min = t0min,
                        x0max = t0max,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0,
                        sd_y_prior = 100),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_poisson.pooled <- model


## Examine output
pars <- c("x0[1]", "y0[1]")
pars_full <- c(pars, "lp__")

title <- "Two piece Pooled Males"
print(title)
print(model, pars = pars_full)
traceplot(model, pars = pars, inc_warmup = FALSE)
plot(model, pars = pars) #, ggtitle(title)) 
pairs(model, pars = pars_full)
summary(fit_poisson)$summary
summary(fit_poisson.pooled)$summary

```

#### Males Separate

```{r}

iter <- 5000
tmax <- 46
t0max <- tmax - 0.75;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

model <- stan(file = "two.piece_poisson.grouped.by.male.stan",
              model_name = "Two piece poisson pooled",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = n_male,
                        ll = data_sim$male,
                        xmax = tmax,
                        x0min = t0min,
                        x0max = t0max,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0,
                        sd_y_prior = 100),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_poisson.separate <- model

pars <- c("y0", "x0")
par(mfrow=c(1,2))

map(pars, ~ plot(model, pars = .))

stan_hist(model, pars = pars)
stan_hist(model, pars = pars)


summary(fit_poisson.separate)$summary

```

##### Plot Results


######  Wide data

```{r}

pars <- c("y0", "x0")

pars_tmp <- summary(model, pars = pars, probs = NULL)$summary %>%
                                                    as_tibble(rownames = "pars")
pars_tmp <- pars_tmp %>% separate_wider_regex(cols = pars, c(par = "^[a-z]+[0-9]+", "\\[", male_index = "[0-9]+", "\\]"), too_few = "error") %>%
    mutate(male_index = as.integer(male_index))

## The code below also works somewhat
## pars_tmp <- pars_tmp %>%
##    mutate(par = str_extract(pars, "^[A-z]+[0-9]+"), male_index = str_replace(pars, ".*\\[([0-9]+)\\]$", "\\1")) %>%

pars_est <- pars_tmp %>% pivot_wider(names_from = par, values_from = c(mean, sd), id_cols = male_index, id_expand = TRUE)

pars_sim <- data_sim %>% select(male_index = male, y0, x0 = x0) %>% distinct()

pars_combined <- full_join(pars_est, pars_sim, by = "male_index")

ggplot(data = pars_combined) +
    geom_point(aes(x = x0, y = mean_x0)) +
    geom_errorbar(aes(x = x0, ymin = mean_x0- 2 * sd_x0, ymax = mean_x0 + 2 * sd_x0)) +
    geom_abline(aes(intercept = 0, slope = 1))
               

ggplot(data = pars_combined) +
    geom_point(aes(x = y0, y = mean_y0)) +
    geom_errorbar(aes(x = y0, ymin = mean_y0- 2 * sd_y0, ymax = mean_y0 + 2 * sd_y0)) +
    geom_abline(aes(intercept = 0, slope = 1))


```
######  Tidy data

```{r}

pars_sim <- data_sim %>%
    select(male_index = male, y0, x0 = x0) %>%
    distinct() %>%
#    mutate(type = "sim") %>%
    pivot_longer(cols = c(x0, y0), values_to = "sim", names_to = "par")


pars <- c("y0", "x0")

## Get basic numbers
pars_tmp <- summary(model, pars = pars, probs = c(0.025, 0.975))
pars_tmp <- pars_tmp$summary %>%
    as_tibble(rownames = "pars") %>%
    select(pars, mean, "2.5%", "97.5%", sd)
## Split y0[ID] and x[ID] into y0 and ID 
pars_est <- pars_tmp %>% separate_wider_regex(cols = pars, c(par = "^[a-z]+[0-9]+", "\\[", male_index = "[0-9]+", "\\]"), too_few = "error") %>%
    mutate(male_index = as.integer(male_index)) %>%
    rename(est = mean)

pars_combined <- full_join(x = pars_est, y = pars_sim, by = c("male_index", "par"))

ggplot(data = pars_combined) +
    geom_point(aes(x = sim, y = est)) +
    geom_errorbar(aes(x = sim, ymin = `2.5%`, ymax = `97.5%`)) +
    geom_abline(aes(intercept = 0, slope = 1)) +
    facet_wrap(~ par, scales = "free",
               labeller = as_labeller(c(`y0` = "Normal Motif Rate", `x0` = "Threshold Temp"))) +
    labs(x = "Simulated", y = "Estimated", title = "Fitting 2 Piece Model to Simulated Data", subtitle = "error ~ Poisson")

```

- Note CI are under estimates due to use of `poisson` model rather than `qpoisson`.

## Assume QPoisson
 
### Males pooled: `y0`, `x0`, and $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

stanc(file = "two.piece_qpoisson.stan",
      verbose = TRUE,
      model_name = "two.piece_qpoisson")

## For debugging
## cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson.stan")

     
model <- stan(file = "two.piece_qpoisson.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = 1,
                        M = 1,
                        ll = rep(1,N),
                        mm = rep(1,N),                        
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 100,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0,
                        sd_y_prior = 100),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson.pooled <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")

title <- "Two piece Pooled Males"
print(title)
print(model, pars = pars)
traceplot(model, pars = pars, inc_warmup = FALSE)
plot(model, pars = pars) #, ggtitle(title)) 
pairs(model, pars = pars_full)




summary(fit_qpoisson.pooled)$summary


```
### Males Separate


#### Separate: `y0`, and `x0`, Pooled  $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

stanc(file = "two.piece_qpoisson.stan",
      verbose = TRUE,
      model_name = "two.piece_qpoisson")

cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson.stan")

     
model <- stan(file = "two.piece_qpoisson.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = n_male,
                        M = 1,
                        ll = data_sim$male, ## y0 and x0
                        mm = rep(1, N), ## Theta
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 100,
                        alpha_theta_prior = 10,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0,
                        sd_y_prior = 100),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson.separate.y0.x0 <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")


print(model, pars = pars)
traceplot(model, pars = pars, inc_warmup = FALSE)
plot(model, pars = pars) #, ggtitle(title)) 
#pairs(model, pars = pars_full)
stan_hist(model, pars = pars)
summary(fit_qpoisson.separate.y0.x0)$summary


```

#### Plot Results

```{r}

pars_sim <- data_sim %>%
    select(male_index = male, y0, x0 = x0) %>%
    distinct() %>%
#    mutate(type = "sim") %>%
    pivot_longer(cols = c(x0, y0), values_to = "sim", names_to = "par")


pars <- c("y0", "x0")

## Get basic numbers
pars_tmp <- summary(model, pars = pars, probs = c(0.025, 0.975))
pars_tmp <- pars_tmp$summary %>%
    as_tibble(rownames = "pars") %>%
    select(pars, mean, "2.5%", "97.5%", sd)
## Split y0[ID] and x[ID] into y0 and ID 
pars_est <- pars_tmp %>% separate_wider_regex(cols = pars, c(par = "^[a-z]+[0-9]+", "\\[", male_index = "[0-9]+", "\\]"), too_few = "error") %>%
    mutate(male_index = as.integer(male_index)) %>%
    rename(est = mean)

pars_combined <- full_join(x = pars_est, y = pars_sim, by = c("male_index", "par"))

ggplot(data = pars_combined) +
    geom_point(aes(x = sim, y = est)) +
    geom_errorbar(aes(x = sim, ymin = `2.5%`, ymax = `97.5%`)) +
    geom_abline(aes(intercept = 0, slope = 1)) +
    facet_wrap(~ par, scales = "free",
               labeller = as_labeller(c(`y0` = "Normal Motif Rate", `x0` = "Threshold Temp"))) +
    labs(x = "Simulated", y = "Estimated", title = "Fitting 2 Piece Model to Simulated Data", subtitle = "error ~ QPoisson")
              


```





#### Separate: `y0`, `x0`, and  $\theta$

- didn't run given noise in previous one with separate `y0` and `x0`
