---
title: "Fit Thermal Models"
author: "Michael Gilchrist"
date: date: "`r Sys.time()`"
output: pdf_format
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Goal

- Fit series of thermal models found in `rTPC` to data collected in the Derryberry lab.
- Based on initial `.R` file created by Kayci Messerly `<messerlykayci1@gmail.com>`
- Intial analysis should be with GLM and mixture models.



# Set up

## Install libraries

```{r}
## install packages user might not have by replacing FALSE with TRUE
if(TRUE) {
    BiocManager::install("mixOmics") ## needed by RVAideMemoire
    install.packages(c("RSQLite", "nls.multstart", "lme4", "RVAideMemoire"))
    ##  Install the thermal curve package from git_hub, not cran
    remotes::install_github("padpadpadpad/rTPC")

}

## load libraries
library(stats)
library(RSQLite)  #not sure why we need this
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)
require(ggplot2)
require(GGally)
require(reshape2)
require(lme4)
require(RVAideMemoire) # provides overdisp.glmer()
require(MASS) # provides negative binomial fitting:  glm.nb
```

## Load Data

```{r}

## Read in ZEFI Data sets
## Treat 'repeatability' as round = 0
## Add round info

## Repeatability was done between round 1 and 2, female was present, but only one temp. so treating as `round = 2` and redefining `round = 2` as `round = 3`

data_raw = list()

data_raw[[1]] <- read.csv(file.path("..", "data", "raw_data", "HSPi-Repeatability-Song-Count.csv")) %>% mutate(round = 2)
data_raw[[2]] <- read.csv(file.path("..", "data", "raw_data", "HSPi-Round-1-Heat-Trials.csv")) %>% mutate(round = 1) ## Note T237 and T230 are missing numbers in the song_count column
data_raw[[3]] <-read.csv(file.path("..", "data", "raw_data", "HSPi-Round-2-Heat-Trials.csv")) %>% mutate(round = 3)

## Join data and discard empty columns
data_full <- full_join(data_raw[[1]], data_raw[[2]]) %>%
    full_join(data_raw[[3]]) %>%
    discard(~all(is.na(.) | . =="")) %>% ## get rid of columns of only NA
    mutate(trial_completed = !(is.na(song_count)) ) %>%
    mutate(song_count = ifelse(is.na(song_count), 0, song_count)) %>%
    mutate(song_count = song_count*1.0) %>% ## convert to a double so it's not treated as an integer
    rename(trial_order = test_order) %>%
    mutate(chamber = as.factor(chamber), male = as.factor(male)) %>%
    ## Using ordered = TRUE only changes the format of the output.
    ## The contrasts change and it's confusing, so don't use it.
    mutate(trial_order = as.factor(trial_order)) %>%
    mutate(song_count_plus_1 = (song_count + 1)) %>%
    mutate(log_song_count_plus_1 = log(song_count + 1)) %>%
    mutate() ## Dummy function so we can comment out lines above it w/o any issues


## Create a database
## Don't see where myConn this is used anywhere else

if(FALSE){
    db_data <- c(data_raw[[1]], data_raw[[2]], data_raw[[3]])
    ZEFI_DB <-"portalR.db"
    myConn <-dbConnect(drv = SQLite(), dbname=ZEFI_DB)

    dbListTables(myConn)

    if(!file.exists(ZEFI_DB))
    {
    dbWriteTable(myConn, "repeatability", data_raw[0])
    dbWriteTable(myConn, "data_raw1", data_raw1)
    dbWriteTable(myConn, "data_raw2", data_raw2)
    dbListTables(myConn)
    }
}

```

# Analysis


## Examine Data

```{r}

data <- data_full #%>% filter(round ==3)

data %>% group_by(temp_target) %>%
    filter(song_count < 400) %>%
    summarize(ave = mean(song_count), sd = sd(song_count), n = length(song_count), n_completed = sum(trial_completed))

```

## Analyze `round = 3` only

- In `round = 3`, the only `trial_completed = FALSE` is in highest temp.
- Only one of 6 champters was used for a given male, thus chamber and male are conflated.
  - Actually makes life easier!

```{r}

data_r3 <- data_full %>% filter(round ==3)

```
## LM Fits on `log(song_counts_plus_one)`

- None of these models fit very well
- Learning how to use `lmList()` function


### Fit Models

```{r}


lm = list()



formula <- as.formula("log_song_count_plus_1 ~ 1 + temp_target + I(temp_target^2)")
## Note syntax below model `formula` doesn't appear in the model summary().
## lm$basic <- lm(formula = formula , data = data_r3)

## Here string of formula is shown in summary()
## The .() forces the evaluation in bquote()
## See: https://win-vector.com/2018/09/01/r-tip-how-to-pass-a-formula-to-lm/
## for detailed explanation of how this works
lm$basic <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp)

## Individual fits to each male
## This isn't that useful, but is my first use of lmList()
formula <- as.formula("log_song_count_plus_1 ~ 1 + temp_target + I(temp_target^2)| male")
lm$male_separate <- tmp <- eval(bquote(lmList(formula = .(formula) , data = data_r3)))
summary(tmp)

## Add trial_order effect
formula <- as.formula("log_song_count_plus_1 ~ 1 + trial_order+ temp_target + I(temp_target^2)")
lm$trial_order <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3) ))
summary(tmp)

## Add male effect 
formula <- as.formula("log_song_count_plus_1 ~ 1 + male + temp_target + I(temp_target^2)")
lm$male <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp) 

## Add chamber effect (fewer chambers than males)
formula <- as.formula("log_song_count_plus_1 ~ 1+ chamber + temp_target + I(temp_target^2)")
lm$chamber <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp)

## Add trial_order effect
formula <- as.formula("log_song_count_plus_1 ~ 1 + trial_order + male + temp_target + I(temp_target^2)")
lm$trial_order_and_male <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3) ))
summary(tmp)


## Add male x trial order
## This doesn't work since we have as many coefficients as data points.
## Oops
formula <- as.formula("log_song_count_plus_1 ~ 1 + male*trial_order + temp_target + I(temp_target^2)")
lm$trial_order_by_male <- tmp <- lm(formula = eval(formula) , data = data_r3)
summary(tmp)

```

### Compare Models

```{r}


## Results: Male is much better predictor than trial_order
anova(lm$male, lm$trial_order)

## Results: Male is better predictor than chamber.
## This indicates males vary more than chambers
anova(lm$male, lm$chamber)

## Results: Adding trial_order to male provides no real gain in fit
anova(lm$male, lm$trial_order_and_male)
```

### Notes

- We do see evidence of male effects.
  The important males have lower intercepts than the rest
- There is little evidence of consistent trial_order effects (an ordered factor).


## GLM and GLMM Poisson, Quasipoisson and Negative-Binomial Fits



## GLM on Song Rate

- See [webpage](https://stats.oarc.ucla.edu/r/faq/random-coefficient-poisson-models/)
  for mixed-effects poisson models
  
```{r}

glm <- list()

## chamber and male are conflated
## Data is very overdispersed given poisson assumptions
glm$poisson <- glm(song_count ~ 1 + male + trial_order + temp_target + I(temp_target^2), data = data_r3, family = poisson(link = "log"))

## This allows for over dispersion which greatly elevates the SE
glm$qpoisson <- glm(model$glm_poisson, family = quasipoisson(link = "log"))

## Gives similar results as with the quasipoisson
glm$nb <- glm.nb(model$glm_poisson)

anova.list(glm)
```

## GLMM (glmer) Model fits

```{r}

model$glmer_poisson <- glmer(
## song_count ~ 1 + chamber + (1|male) + (temp_target|male) + (I(temp_target^2)|male),
    song_count ~ 1 + trial_order + ((1 + temp_target + I(temp_target^2))|male),
    data = data_r3,
    family = poisson(link = "log"),
    control = glmerControl(optCtrl = list(maxiter = 300)),
    verbose = TRUE)

## Still a lot of overdispersion
model$glmer_od <- overdisp.glmer(model$glmer_poisson)

## I think this is a correct fit.  If I don't include the 1 + temp... w/o the |male, I don't get any effects.
model$glmer_nb_1 <- glmer.nb(
    song_count ~ 1 + temp_target + I(temp_target^2) + trial_order + ((1 + temp_target + I(temp_target^2))|male),
    data = data_r3,
    family = poisson(link = "log"),
    control = glmerControl(optCtrl = list(maxiter = 300)),
    verbose = TRUE)


model$glmer_nb_2 <- update(model$glmer_nb_1, formula =  song_count ~ 1 + temp_target + I(temp_target^2) + ((1 + temp_target + I(temp_target^2))|male) )

mode
```





## For Later: Lactin2 

```{r}

data <- data_full %>% filter(round ==3)

data %>% group_by(temp_target) %>%
    filter(song_count < 400) %>%
    summarize(ave = mean(song_count, na.rm = TRUE), sd = sd(song_count, na.rm = TRUE)) 


## show the data
## windows()
ggplot(data, aes(temp_target, song_count)) + 
    geom_point() +
    geom_jitter() +
  theme_bw(base_size = 12) +
  labs(x='Temperature (C)', 
       y='Song Count', 
       title = 'Song Count Across Temperature')

## choose the model
mod = 'lactin2_1995'

## get starting values
start_vals <- get_start_vals(data$temp_target, data$song_count, model_name = 'lactin2_1995')
  
## Get limits
low_lims <- get_lower_lims(data$temp_target,data$song_count,model_name = 'lactin2_1995')
upper_lims <- get_upper_lims(data$temp_target, data$song_count, model_name = 'lactin2_1995')

start_vals
low_lims
upper_lims
  
## fit model
fit <- nls_multstart(song_count~lactin2_1995(temp = temp_target, a,b,tmax,delta_t),
                                  data = data,
                                  iter = 600,
                                  start_lower = start_vals-10,
                                  start_upper = start_vals+10,
                                  lower = low_lims,
                                  upper = upper_lims,
                                  supp_errors = 'Y',
                                  convergence_count = FALSE)
  
## look at model fit
summary(fit)

## Predict new data 
preds <- data.frame(temp = seq(min(data$temp_target), max(data$temp_target), length.out = 6))
preds <- broom::augment(fit, newdata = preds)

## plot data and model fit
## windows()
ggplot(preds)+
  geom_point(aes(temp_target,song_count), data)+
  geom_line(aes(temp, .fitted), preds, col = 'blue')+
  theme_bw()+
  labs(x='Temperature (C)',
       y='Song Count',
       title = 'Song Count Across Temperatures')


```


## Unweighted model

- Fit 4 Chosen model formulations in rTPC


```{r}

d <-filter(heat_roundt2, Male =="T236")
d_fits <- nest(d, data = c(temp_target,song_count)) %>%
  mutate(lactin = map(data,~nls_multstart(song_count~lactin2_1995(temp = temp_target, a, b, tmax, delta_t),
                      data = .x,
                      iter = c(3,3,3,3),
                      start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'lactin2_1995')-10,
                      start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'lactin2_1995')+10,
                      lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'lactin2_1995'),
                      upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'lactin2_1995'),
                      supp_errors = 'Y',
                      convergence_count = FALSE)),
weibull = map(data,~nls_multstart(song_count~weibull_1995(temp = temp_target, a, topt, b,c),
                                  data = .x,
                                  iter = c(4,4,4,4),
                                  start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'weibull_1995')-10,
                                  start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'weibull_1995')+10,
                                  lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'weibull_1995'),
                                  upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'weibull_1995'),
                                  supp_errors = 'Y',
                                  convergence_count = FALSE)),
modifiedgaussian = map(data, ~nls_multstart(song_count~modifiedgaussian_2006(temp = temp_target, rmax,topt, a, b),
                                            data = .x,
                                            iter = c(3,3,3,3),
                                            start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006')-10,
                                            start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006')+10,
                                            lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006'),
                                            upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006'),
                                            supp_errors = 'Y',
                                            convergence_count = FALSE)),
briere = map(data,nls_multstart(song_count~briere2_1999(temp = temp_target, tmin, tmax, a,b),
                                data = .x,
                                iter = c(4,4,4,4),
                                start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'briere2_1999')-10,
                                start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'briere2_1999')+10,
                                lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'briere2_1999'),
                                upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'briere2_1999'),
                                supp_errors = 'Y',
                                convergence_count = FALSE)))
