---
title: "Fit Thermal Models"
author: "Michael Gilchrist"
date: date: "`r Sys.time()`"
output: pdf_format
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Goal

- Fit series of thermal models found in `rTPC` to data collected in the Derryberry lab.
- Based on initial `.R` file created by Kayci Messerly `<messerlykayci1@gmail.com>`
- Intial analysis should be with GLM and mixture models.



# Set up

## Install libraries

```{r}
## install packages user might not have by replacing FALSE with TRUE
if(FALSE) {
    BiocManager::install("mixOmics") ## needed by RVAideMemoire
    install.packages(c("RSQLite", "nls.multstart", "lme4", "RVAideMemoire"))
    ##  Install the thermal curve package from git_hub, not cran
    remotes::install_github("padpadpadpad/rTPC")

}

## load libraries
library(stats)
require(MASS) # provides negative binomial fitting:  glm.nb
library(RSQLite)  # Don't think we need this.
library(rTPC)  ## 
library(nls.multstart)
library(broom)
library(tidyverse)
require(ggplot2)
require(GGally)
require(reshape2)
require(lme4)
require(RVAideMemoire) # provides overdisp.glmer()

```

## Load Data

```{r}

## Read in ZEFI Data sets
## Treat 'repeatability' as round = 0
## Add round info

## Repeatability was done between round 1 and 2, female was present, but only one temp. so treating as `round = 2` and redefining `round = 2` as `round = 3`

data_raw = list()

data_raw[[1]] <- read.csv(file.path("..", "data", "raw_data", "HSPi-Round-1-Heat-Trials.csv")) %>% mutate(round = 1) %>%
    ## Note T237 and T230 are missing numbers in the song_count column
    ## so we are filtering these observations out until they are found
    filter(!is.na(song_count))

data_raw[[2]] <- read.csv(file.path("..", "data", "raw_data", "HSPi-Repeatability-Song-Count.csv")) %>%
    mutate(round = 2) %>%
    group_by(male) %>%
    mutate(test_order = rank(date)) %>%
    ungroup()

data_raw[[3]] <-read.csv(file.path("..", "data", "raw_data", "HSPi-Round-2-Heat-Trials.csv")) %>% mutate(round = 3)

## Join data and discard empty columns
data_full <- full_join(data_raw[[1]], data_raw[[2]]) %>%
    full_join(data_raw[[3]]) %>%
    discard(~all(is.na(.) | . =="")) %>% ## get rid of columns of only NA
    mutate(trial_completed = !(is.na(song_count)) ) %>%
    mutate(song_count = ifelse(is.na(song_count), 0, song_count)) %>%
    mutate(song_count = song_count*1.0) %>% ## convert to a double so it's not treated as an integer
    mutate(chamber = as.factor(chamber), male = as.factor(male)) %>%
    ## create a global variable trial_order based on individual rounds
    mutate(trial_index = as.integer(round*10+test_order)) %>%
    mutate(song_count_plus_1 = (song_count + 1)) %>%
    mutate(log_song_count_plus_1 = log(song_count + 1)) %>%
    mutate(temp_target = as.numeric(temp_target)) %>%
    mutate() ## Dummy function so we can comment out lines above it w/o any issues

```

# Analysis


## Examine Data

```{r}

data <- data_full #%>% filter(round ==3)

data %>% group_by(temp_target) %>%
    filter(song_count < 400) %>%
    summarize(ave = mean(song_count), sd = sd(song_count), n = length(song_count), n_completed = sum(trial_completed))

## Look at occurrance of males across rounds
tmp1 <- data_full %>%
    filter(round==1) %>%
    pull(male) %>%
    unique(); tmp1

tmp2 <- data_full %>%
    filter(round %in% c(2,3)) %>%
    pull(male) %>%
    unique() ; tmp2

lapply(list(tmp1, tmp2, intersect(tmp1, tmp2)), length)

```

## Plot Raw Data

```{r}


xlab <- "Temperature"
ylab <- "song_count"

plot_data_only <-
    ggplot(data_full) + 
    aes(x = temp_target,
        y = song_count,
        color = as.factor(round)) +
    facet_wrap("male", scales = "free_y") +
    geom_jitter() +
    labs( title = paste( substitute(x), " vs ", substitute(y)))

pdf(file = "song_count_vs_temp_data_only.pdf")

plot_data_only

dev.off()
```
## Earliest Analysis
### Analyze `round 2 & 3`

- In `round = 3`, 
  - The only `trial_completed = FALSE` is in highest temp.
  - Only one of 6 chambers was used for a given male, thus chamber and male are conflated.
  - Actually makes life easier!

```{r}

data_r3 <- data_full %>% filter(round ==3)

```
### LM Fits on `log(song_counts_plus_one)`

- None of these models fit very well



#### Fit each individual separately

- Learning how to use `lmList()` function

```{r}

## Individual fits to each male
## This isn't that useful, but is my firstuse of lmList()

lmList <- list()

lmList$by_male <- tmp <-
    lme4::lmList(log_song_count_plus_1 ~ 1 + temp_target + I(temp_target^2) | male,
           data = data %>% filter(round %in% c(2,3)),
           na.action = na.omit)
           

summary(tmp)

plot.lmList4(tmp, log_song_count_plus1 ~ fitted(.)| male)


```
#### Fit multiple variations of grouping

```{r}


lm = list()

formula <- as.formula("log_song_count_plus_1 ~ 1 + temp_target + I(temp_target^2)")
## Note syntax below model `formula` doesn't appear in the model summary().
## lm$basic <- lm(formula = formula , data = data_r3)

## Here string of formula is shown in summary()
## The .() forces the evaluation in bquote()
## See: https://win-vector.com/2018/09/01/r-tip-how-to-pass-a-formula-to-lm/
## for detailed explanation of how this works
lm$basic <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp)

## Add trial_index effect
formula <- as.formula("log_song_count_plus_1 ~ 1 + trial_index+ temp_target + I(temp_target^2)")
lm$trial_index <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3) ))
summary(tmp)

## Add male effect 
formula <- as.formula("log_song_count_plus_1 ~ 1 + male + temp_target + I(temp_target^2)")
lm$male <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp) 

## Add chamber effect (fewer chambers than males)
formula <- as.formula("log_song_count_plus_1 ~ 1+ chamber + temp_target + I(temp_target^2)")
lm$chamber <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp)

## Add trial_index and male effects
formula <- as.formula("log_song_count_plus_1 ~ 1 + trial_index + male + temp_target + I(temp_target^2)")
lm$trial_index_and_male <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3) ))
summary(tmp)


## Add male x trial order
## This doesn't work since we have as many coefficients as data points.
## Oops
formula <- as.formula("log_song_count_plus_1 ~ 1 + male*trial_index + temp_target + I(temp_target^2)")
lm$trial_index_by_male <- tmp <- lm(formula = eval(formula) , data = data_r3)
summary(tmp)


## add slope x male interaction term

lm$male_by_temp <- tmp <- lm(formula = log_song_count_plus_1 ~ 1 + temp_target*male + I(temp_target^2), data = data_r3)
summary(tmp)


## slope x male - male intercept
## This suggests that the slopes are similar between males (though the SE are large)
lm$male_by_temp <- tmp <- lm(formula = log_song_count_plus_1 ~ 1 + temp_target:male + I(temp_target^2), data = data_r3)
summary(tmp)

## curvature x male - male intercept
## This suggests that the slopes are similar between males (though the SE are large)
lm$male_by_temp_sq <- tmp <- lm(formula = log_song_count_plus_1 ~ 1 + temp_target + I(temp_target^2):male, data = data_r3)
summary(tmp)
```

## GLMM (glmer) Model fits

```{r}

glmer = list()

glmer$poisson <- tmp <-
    glmer(
        song_count ~
            1+ 1|male + temp_target + I(temp_target^2),
        data = data_r3,
        family = poisson(link = "log"),
        control = glmerControl(optCtrl = list(maxiter = 300)),
        verbose = TRUE)

## Still a lot of overdispersion
glmer$od <- overdisp.glmer(glmer$poisson)

## I think this is a correct fit.  If I don't include the 1 + temp... w/o the |male, I don't get any effects.
glmer$nb_1 <- glmer.nb(
    song_count ~ 1 + temp_target + I(temp_target^2) + trial_index + ((1 + temp_target + I(temp_target^2))|male),
    data = data_r3,
    family = poisson(link = "log"),
    control = glmerControl(optCtrl = list(maxiter = 300)),
    verbose = TRUE)


glmer$nb_2 <- update(glmer$nb_1, formula =  song_count ~ 1 + temp_target + I(temp_target^2) + ((1 + temp_target + I(temp_target^2))|male) )

mode
```


### Compare Models

```{r}


## Results: Male is much better predictor than trial_index
anova(lm$male, lm$trial_index)

## Results: Male is better predictor than chamber.
## This indicates males vary more than chambers
anova(lm$male, lm$chamber)

## Results: Adding trial_index to male provides no real gain in fit
anova(lm$male, lm$trial_index_and_male)
```

### Notes

- We do see evidence of male effects.
  The important males have lower intercepts than the rest
- There is little evidence of consistent trial_index effects.
- Adjusted $R^2$ don't seem to vary much.



## First Analysis Shared with Liz

- Slightly more organized than previous

### lm analysis of `log(song_count_plus_1)`

```{r}


## Plotting based on: https://www.tjmahr.com/plotting-partial-pooling-in-mixed-effects-models/

xlab <- "Temperature"
ylab <- "log(song_count + 1)"

pdf(file = "log-plus-1_vs_temp_all.pdf")

plot_log_plus_1 <- ggplot(data_full  ) + 
    aes(x = temp_target,
        y = log_song_count_plus_1,
        color = (round) +
    stat_smooth(method = "lm",
                se = TRUE,
                formula = y ~ x + I(x^2), size = 1,
                ) +
  # Put the points on top of lines
  #  geom_point() +
    geom_jitter() +
  facet_wrap("male", scales = "free_y") +
    labs( title = paste( substitute(x), " vs ", substitute(y), ": Data Only"), 
         subtitle = "lm:y ~ 1 + x + x^2")

dev.off()



```

### GLM analysis of `song_count` and `temp_target` 

```{r}

## Code below used for posting to stackexchange
data_full %>% filter(male %in% c("T258", "T259"))%>% dplyr::select(male, temp_target, song_count) %>% rename(ID = male, x = temp_target, y = song_count)  %>% dput()

## Choose some males to use for starting values based on log fit
reasonable.males = c("T229", "T237", "T258")

glm_result <- glm_qpoisson_target <- glm(song_count ~ 1 + temp_target + I(temp_target^2),
                    family = quasipoisson(),
    data = data_full %>% filter(male %in% reasonable.males))

xlab <- "Temperature"
ylab <- "song_count"



plot_glm <- plot_data_only +
    aes(x = temp_target,
        y = song_count,
        color = (round == 1)) + ## Combine round 2 and 3
    stat_smooth(method = "glm",
                method.args = list(
                    family = quasipoisson(link = "log"),
                    start = coefficients(glm_result),
                    maxit = 100),
                se = FALSE,
                formula = y ~ 1 + x + I(x^2), size = 1,
                ) +
    labs( title = paste( substitute(x), " vs ", substitute(y), ": Data Only"), 
         subtitle = "glm:y ~quasipoisson(exp[1 + x + x^2])")

pdf(file = "song_count_vs_temp_multiple.pdf")

plot_data_only
plot_log_plus_1
plot_glm

dev.off()


```


    





## Second Analysis for Liz

### GLM analysis of `song_count` and `temp_mean` 

```{r}

data_tmp <- data_full %>% filter(round == 2) %>% dplyr::select(male, song_count) %>% group_by(male)

tmp_summarized <- summarize(data_tmp, var(song_count), mean(song_count))

## Methods of Moments: if x ~ NB(p, p/(p + r)), var(x) = p + p^2/r, mean(x) = p
## Alternative parameterization
## p = 1 - \mu/\sigma^2;
## r = \mu^2/(\sigma^2 - \mu)
## Where,
##   \mu = mean(x) = p r/(1-p)
##   \sigma^2 = var(x) = (1-p)/(1+p)^2


var_tmp = var(data_tmp)


tmp <- glm(song_count ~ male,
           data = data_tmp,
    family = quasipoisson()
    )


data <- data_full %>% filter(round == 3)

xlab <- "Temperature"
ylab <- "song_count"

plot_data_temp <-
    ggplot(data = data) + 
    aes(x = temp_target,
        y = song_count,
        color = as.factor(round)) +
    facet_wrap("male", scales = "free_y") +
    geom_jitter() +
    labs( title = paste( substitute(x), " vs ", substitute(y), ": Data Only") )
plot_data_temp

xlab= "Humidity"

plot_data_humidity <-
    ggplot(data = data) + 
    aes(x = humidity_mean,
        y = song_count,
        color = as.factor(round)) +
    facet_wrap("male", scales = "free_y") +
    geom_jitter() +
    labs( title = paste( substitute(x), " vs ", substitute(y), ": Data Only") ) 
plot_data_humidity


## As one might expect, Humidity is negatively correlated with temperature
ylab = "temp_mean"
plot_data_humidity_vs_temp <-
    ggplot(data = data) + 
    aes(x = temp_mean,
        y = humidity_mean,
        color = as.factor(round)) +
    facet_wrap("male", scales = "free_y") +
    geom_jitter() +
    labs( title = paste( substitute(x), " vs ", substitute(y), ": Data Only") )
plot_data_humidity_vs_temp


## Choose some males to use for starting values based on log fit
reasonable.males = c("T229", "T237", "T258")

glm_result <- glm_qpoisson_mean <- glm(song_count ~ 1 + temp_mean + I(temp_mean^2),
                    family = quasipoisson(),
    data = data_full %>% filter(male %in% reasonable.males))

xlab <- "Temperature"
ylab <- "song_count"



plot_glm <- plot_data_only +
    aes(x = temp_mean,
        y = song_count,
        color = (round == 1)) + ## Combine round 2 and 3
    stat_smooth(method = "glm",
                method.args = list(
                    family = quasipoisson(link = "log"),
                    start = coefficients(glm_result),
                    maxit = 100),
                se = FALSE,
                formula = y ~ 1 + x + I(x^2), size = 1,
                ) +
    labs(x = xlab, y = ylab,
         subtitle = "glm:y ~quasipoisson(exp[1 + x + x^2])",
         title = "Song Count vs Temp: Quasi-Poisson Regression") 


pdf(file = "song_count_vs_temp_mean_glm.pdf")

plot_data_only
plot_log_plus_1
plot_glm

dev.off()


```


    



## Make a model fit tibble - Not complete

```{r}
round_tbl <- tibble(round_1 = FALSE, round_2 = c(TRUE, FALSE),
       round_3 = TRUE
       ) %>% rep(2)



model_cat <- c("lm", "glm", "glmer")
round_1 <- FALSE
round_2 <- c(FALSE, TRUE)
round_3 <- TRUE
male <- c(FALSE, TRUE)
chamber <- c(FALSE, TRUE)
trial_index <- c(FALSE, TRUE)

fit_tbl_names <- c("model_cat",
      "round_1",
      "round_2",
      "round_3",
      "male",
      "chamber",
      "trial_index",
      "temp_slope",
      "temp_curve",
      "male_x_temp_slope",
      "male_x_temp_curve",
      "residual_df".
      "residual_ss"
      "R2",
      "AIC",
      "fit_obj"
      )

fit_tbl <- data.frame("a" = 1, "b"= 1:2, "c" = 1:3)

                 ,
                  names = 
      )
)
comment(fit_tbl) <- "temp_slope = temp_mean, temp_curve = I(temp_mean^2)"

filter_data_lm() <- list(
formulas_lm() <-
    


```

## For Later: Lactin2 

```{r}

data <- data_full %>% filter(round ==3)

data %>% group_by(temp_target) %>%
    filter(song_count < 400) %>%
    summarize(ave = mean(song_count, na.rm = TRUE), sd = sd(song_count, na.rm = TRUE)) 


## show the data
## windows()
ggplot(data, aes(temp_target, song_count)) + 
    geom_point() +
    geom_jitter() +
  theme_bw(base_size = 12) +
  labs(x='Temperature (C)', 
       y='Song Count', 
       title = 'Song Count Across Temperature')

## choose the model
mod = 'lactin2_1995'

## get starting values
start_vals <- get_start_vals(data$temp_target, data$song_count, model_name = 'lactin2_1995')
  
## Get limits
low_lims <- get_lower_lims(data$temp_target,data$song_count,model_name = 'lactin2_1995')
upper_lims <- get_upper_lims(data$temp_target, data$song_count, model_name = l'actin2_1995')

start_vals
low_lims
upper_lims
  
## fit model
fit <- nls_multstart(song_count~lactin2_1995(temp = temp_target, a,b,tmax,delta_t),
                                  data = data,
                                  iter = 600,
                                  start_lower = start_vals-10,
                                  start_upper = start_vals+10,
                                  lower = low_lims,
                                  upper = upper_lims,
                                  supp_errors = 'Y',
                                  convergence_count = FALSE)
  
## look at model fit
summary(fit)

## Predict new data 
preds <- data.frame(temp = seq(min(data$temp_target), max(data$temp_target), length.out = 6))
preds <- broom::augment(fit, newdata = preds)

## plot data and model fit
## windows()
ggplot(preds)+
  geom_point(aes(temp_target,song_count), data)+
  geom_line(aes(temp, .fitted), preds, col = 'blue')+
  theme_bw()+
  labs(x='Temperature (C)',
       y='Song Count',
       title = 'Song Count Across Temperatures')


```


## Unweighted model

- Fit 4 Chosen model formulations in rTPC


```{r}

d <-filter(heat_roundt2, Male =="T236")
d_fits <- nest(d, data = c(temp_target,song_count)) %>%
  mutate(lactin = map(data,~nls_multstart(song_count~lactin2_1995(temp = temp_target, a, b, tmax, delta_t),
                      data = .x,
                      iter = c(3,3,3,3),
                      start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'lactin2_1995')-10,
                      start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'lactin2_1995')+10,
                      lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'lactin2_1995'),
                      upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'lactin2_1995'),
                      supp_errors = 'Y',
                      convergence_count = FALSE)),
weibull = map(data,~nls_multstart(song_count~weibull_1995(temp = temp_target, a, topt, b,c),
                                  data = .x,
                                  iter = c(4,4,4,4),
                                  start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'weibull_1995')-10,
                                  start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'weibull_1995')+10,
                                  lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'weibull_1995'),
                                  upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'weibull_1995'),
                                  supp_errors = 'Y',
                                  convergence_count = FALSE)),
modifiedgaussian = map(data, ~nls_multstart(song_count~modifiedgaussian_2006(temp = temp_target, rmax,topt, a, b),
                                            data = .x,
                                            iter = c(3,3,3,3),
                                            start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006')-10,
                                            start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006')+10,
                                            lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006'),
                                            upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'modifiedgaussian_2006'),
                                            supp_errors = 'Y',
                                            convergence_count = FALSE)),
briere = map(data,nls_multstart(song_count~briere2_1999(temp = temp_target, tmin, tmax, a,b),
                                data = .x,
                                iter = c(4,4,4,4),
                                start_lower = get_start_vals(.x$temp_target, .x$song_count, model_name = 'briere2_1999')-10,
                                start_upper = get_start_vals(.x$temp_target, .x$song_count, model_name = 'briere2_1999')+10,
                                lower = get_lower_lims(.x$temp_target, .x$song_count, model_name = 'briere2_1999'),
                                upper= get_upper_lims(.x$temp_target, .x$song_count, model_name = 'briere2_1999'),
                                supp_errors = 'Y',
                                convergence_count = FALSE)))
