---
title: "Evaluate, Visualize, and Compare Piecewise Regression using BRMS"
author: "Michael Gilchrist"
date: "Created: 2023-02-07; Code run on `r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE, }

## Global options
options(digits = 3)

### Knitr specific
knitr::opts_chunk$set(
                      warning = TRUE, # show warnings
                      message = FALSE, # show messages
                      error = TRUE, # do not interrupt generation in case of errors,
                      echo = TRUE,  # show R code
                      fig.width = 8,
                      fig.path = "Figures/"
                  )

options("warn" = 1) ## print warnings when they occur

if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=5E1, output = TRUE)

```
# Goal

- Visualize the `brms` fits of two piece linear regression model.
- Fit code:  `../2023-02-07_brms_two.piece_focus.on.x0.fittings`


## Recap

# Set up

## Install libraries

```{r, message = FALSE}

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(viridisLite)
#options(ggplot2.continuous.colour="viridis",
#        ggplot2.discrete.colour="viridis",
#        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
#        ggplot2.scale_fill_continuous = scale_fill_viridis_c)
## Loading required package: viridisLite
colourcodes <- c("#d4a665", "#d27fff", "#7fd9ff")
colourpal <- c(NLS="#d4a665", NLME="#d27fff", MCMC="#7fd9ff")

library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(tidybayes)
library(cmdstanr)
library(rstan)
options(mc.cores = 4) #(parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(brms)
library(loo)
library(shinystan)
library(reshape2)
library(lme4)
library(latex2exp)

```
# Load Fits

```{r}

load(file.path("input", "fit_tbl.Rdata"), verbose = TRUE)
load(file.path("input", "data_stan.Rda"), verbose = TRUE)


```

# Evaluate Fits

## Examine structure of `fit_tbl`

```{r}

head(fit_tbl)

```

## Tidybayes vignette

```{r}



fit_brms <- fit_tbl[[9, 7]][[1]]

class(fit_tbl$fit)

fits_brms <- fit_tbl[c(1,4,7), "fit"]

    
loo(fit_tbl[1, "fit"][[1]][[1]], fit_tbl[4, "fit"][[1]][[1]])

fit_stan <- fit_brms$fit

get_variables(fit_brms)  ## same results as for fit_stan

plot(conditional_effects(fit_brms, effects = "x"), points = TRUE) 
plot(conditional_effects(fit_brms, effects = "male"), points = TRUE)

```
## R-Bloggers: Evaluate single fit

Code based on: [R-Bloggers: Nonlinear modeling using nls, nlme and brms](https://www.r-bloggers.com/2020/03/nonlinear-modelling-using-nls-nlme-and-brms/)


```{r}

sampling_dist <- "nb"
x0_flag <- "uniform_1"
y0_flag <- "individual"

curr_row <- which(fit_tbl$sampling_dist == sampling_dist &
                  fit_tbl$x0_flag == x0_flag &
                  fit_tbl$y0_flag == y0_flag)

fit_row <- fit_tbl[curr_row, ]

fit <- fit_row$fit[[1]]

male <- factor(levels(data_stan$male))

pars_male <- paste0("b_y0_male", male)

variable_plot <- c("b_x0_Intercept", "shape", pars_male[1:3])

prior_summary(fit)
summary(fit)

## Look at samples
plot(fit, variable = variable_plot)
pairs(fit, variable = variable_plot)

```


## Generate model predictions

This doesn't currently work

```{r}

## Define Stan function locally
## Doesn't seem to work

fmax <- function(x, y) max(x, y)

# I don't think this is necessary since I don't have any 'user defined functions'
expose_functions(fit, vectorize = TRUE)

pred_x <- seq(20, 46, length.out = 20)

newdata <- crossing(x = pred_x, male = droplevels(male[1:3]))

                                        #newdata <- tibble(x = pred_x)
fit_pred <- predict(fit, newdata) #, newdata = newdata) %>% tibble()

fit_fitted <- fitted(fit, newdata = newdata) %>% tibble()


fit_ribbons <- tibble(
  x = pred_x,
  parentFraction=fit_fitted$Estimate,
  Estimate = fit_fitted$Estimate,
  pred_lower = fit_pred$Q2.5,
  pred_upper = fit_pred$Q97.5,
  fitted_lower = fit_fitted$Q2.5,
  fitted_upper = fit_fitted$Q97.5)

ggplot(data_stan, aes(x=x, y=y)) +
  geom_point(size=3) +
  geom_ribbon(data=fit_ribbons, aes(ymin=pred_lower, ymax=pred_upper), 
              alpha=0.2, fill=colourcodes[3]) +
  geom_ribbon(data=fit_ribbons, aes(ymin=fitted_lower, ymax=fitted_upper), 
              alpha=0.5, fill=colourcodes[3]) +
  geom_line(data=fit_ribbons, aes(y=Estimate), colour=colourcodes[3], 
            size=1)
