---
title: "Piecewise Regression QPoisson Error on Real Data using STAN Directly"
author: "Michael Gilchrist"
date: "date: 2022-12-22"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=2E2, output = TRUE)

```
# Goal

- Fit two piece quasipoisson to data

## Recap

# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(loo)


## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Load Data

```{r}
load(file.path("input", "data.processing_2022-12-15.Rda"),
     verbose = TRUE)


motif_data
```


# Examine Data

## Create Working Dataset

```{r}

filter_data <- TRUE

if(filter_data) {
  males_filtered_disp <- motif_stats_40C %>%
    filter(dispersion < 50) %>%
    pull(male)

  males_filtered_mean <- motif_stats %>%
    filter(mean > 10) %>% # changing from 10 to 40 removes previous male 7 (T258)
    pull(male)

  male_vector <- intersect(males_filtered_mean, males_filtered_disp)
} else {
  male_vector <- motif_data %>% select(male) %>% distinct()
}

data_ind <- motif_data %>%
    filter(male %in% male_vector) %>%
    mutate(male = droplevels(male)) %>%
    mutate(index = as.integer(male)) %>%
    mutate(male = as.character(male)) %>% 
    arrange(index) %>%
    select(male, index, motif_count, temp, round, trial_round, date, counter) %>% 
  ##    left_join(index_shape, by = "index") %>%
  mutate()

stats_ind <- motif_stats %>%
    filter(male %in% male_vector) 


data_ind <- data_ind %>% filter(temp < 38) %>%
    group_by(male) %>% mutate(y0_simple.est = mean(motif_count), phi_ind = var(motif_count)/y0_simple.est) %>% ## phi is overdispersion parameter
    ungroup()

summary(data_ind)

summary_stats <- data_ind %>% ungroup() %>%  summarize(y0_bar = mean(y0_simple.est), y0_sd = sd(y0_simple.est), phi_bar = median(phi_ind))

n_male <- length(unique(data_ind$male))

```


# Analyze Real Data:


## Set Up Data

```{r}

data <- data_ind
motif_count <- data %>% pull(motif_count)
temp <- data %>% pull(temp)
N <- length(temp)
index <- data %>% pull(index)
## parameters to be printed

pars <- c("t0", "y0")
pars_full <- c(pars, "lp__")
```

## Fit Models
 
```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)
n_cores <- 4
n_chains <- n_cores

##y0_grouping <- map_int(data$male, ~ if_else(. %in% y0_group[[1]], 1, 2))

## For debugging
## cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")

stan_model(file="two.piece_qpoisson_2.0.stan",
           verbose = TRUE)


## Define groups

flags <- c("separate", "grouping_1", "pooled")
flags_x <- flags
flags_y <- flags
fit_tbl <- crossing(x0 = flags_x, y0 = flags_y,
                    desc = "NA_character",
                    y0_group_list = list(NA),
                    x0_group_list = list(NA),
                    fit = list(NA),
                    llik = list(NA),
                    r_eff = list(NA),
                    loo = list(NA)
                    )



for(x_flag in flags_x) {
    for(y_flag in flags_y) {

        desc <- paste0("x0 = ", x_flag, ", y0 = ", y_flag)
        curr_row <- which(fit_tbl$x0 == x_flag & fit_tbl$y0 == y_flag)

        fit_tbl[ curr_row, ]$desc <- desc
        print(desc)
        
        x0_group_list <- list()
        y0_group_list <- list()
        
        switch(x_flag,
               separate = {
                   x0_group_list <- data$male %>% unique() %>% as.list()
               },
               grouping_1 = {
                   ## set up groupings based on 2022-12-20 analysis
                   ## Using male ID's instead index to make code more robust
                   x0_group_list[[1]] <- c("T235", "T237", "T244", "T247", "T257", "T260")
                   x0_group_list[[2]] <- c("T234", "T236", "T243", "T246", "T258")
               },
               pooled = {
                   x0_group_list[[1]] <- data$male
               }
               )


        switch(y_flag,
               separate = {
                   y0_group_list <- data$male %>% unique() %>% as.list()
               },
               grouping_1 = {
                   ## set up groupings based on 2022-12-20 analysis
                   ## Using male ID's instead index to make code more robust
                   y0_group_list[[1]] <- c("T234", "T243", "T244", "T246", "T258", "T260") 
                   y0_group_list[[2]] <- c("T235", "T236", "T237", "T247", "T257")
               },
               pooled = {
                   y0_group_list[[1]] <- data$male
               }
               )

        fit_tbl[ curr_row, ]$x0_group_list[[1]] <- x0_group_list

        fit_tbl[ curr_row, ]$y0_group_list[[1]] <- y0_group_list    

        ## Convert lists to a vector of concatenated strings
        ## This will simplify mapping male to an x0/y0 index 
        x0_group <-lapply(x0_group_list, paste, collapse = " ") %>% unlist()
        y0_group <-lapply(y0_group_list, paste, collapse = " ") %>% unlist()

        x0_index <- sapply(as.character(data$male), function(x) str_which(x0_group, x))
        y0_index <- sapply(as.character(data$male), function(x) str_which(y0_group, x))

        fit <- stan(file = "two.piece_qpoisson_2.0.stan",
                      model_name = desc,
                      data=list(x = temp,
                                y = motif_count,
                                N = N,
                                X = length(x0_group),
                                Y = length(y0_group),
                                NB = 1,
                                xx = x0_index,
                                yy = y0_index,
                                nbb = rep(1,N),
                                xmax = tmax,
                                x0_min = t0min,
                                x0_max = t0max,
                                y_xmax = 0,
                                y0_min = 10,
                                sd_y0_prior = 1000,
                                alpha_theta_prior = 40,
                                ##tp = tp,
                                ## max threshold value.
                                ## having it too close to xmax *sometimes* leads to sampling
                                ## near xmax, but with lower lp and very high E13) b0 values
                                y_xmax = 0),
                    cores = n_cores,
                    chains = n_chains,                    
                    iter = iter,
                    warmup = floor(iter/2),
                    verbose = TRUE)
        
        fit_tbl[ curr_row, ]$fit <- list(fit)
        
    }
}


## save(file = "fit_tbl.Rda", fit_tbl)

```

#### LOO Analysis

```{r}

for(curr_row in 1:length(fit_tbl$fit)) {

  desc <- fit_tbl[[curr_row, "desc"]]
  fit <- fit_tbl[[curr_row, "fit"]][[1]]
  print(desc)

  # loo analysis based on: http://mc-stan.org/loo/articles/loo2-with-rstan.html
  #
  # Extract pointwise log-likelihood
  # using merge_chains=FALSE returns an array, which is easier to 
  # use with relative_eff()
  llik <- extract_log_lik(fit, merge_chains = FALSE)
  fit_tbl[[curr_row, "llik"]] <- list(llik)

  # as of loo v2.0.0 we can optionally provide relative effective sample sizes
  # when calling loo, which allows for better estimates of the PSIS effective
  # sample sizes and Monte Carlo error
  r_eff <- relative_eff(exp(llik), cores = n_cores)
  fit_tbl[[curr_row, "r_eff"]] <- list(r_eff)
  
  # preferably use more than 2 cores (as many cores as possible)
  # will use value of 'mc.cores' option if cores is not specified
  loo <- loo(llik, r_eff = r_eff,
             cores = n_cores,
             save_psis = TRUE,
             moment_match = TRUE) 
  fit_tbl[[curr_row, "loo"]] <- list(loo)
  print(loo)

}


comp <- loo_compare(fit_tbl$loo) #, simplify = TRUE)

comp %>% rownames() %>% sub(pattern = "model", x= ., "") %>% as.integer() 

print(comp)
print(fit_tbl$desc)

```

#### Plot Results

```{r}

for(curr_row in 1:length(fit_tbl$fit)) {

    desc <- fit_tbl[[curr_row, "desc"]]
    fit <- fit_tbl[[curr_row, "fit"]]

    x0_group_list <- fit_tbl[[curr_row, "x0_group_list"]][[1]]
    y0_group_list <- fit_tbl[[curr_row, "y0_group_list"]][[1]]
}

    
    pars <- c("x0", "y0", "theta")
    pars_full <- c(pars, "lp__")

    print(desc)
    print(fit, pars = pars)
                                        #traceplot(model, pars = pars, inc_warmup = FALSE)
                                        #plot(model, pars = pars) #, ggtitle(title))

        pairs(model, pars = pars_full)

        ## Plot parameter estimate summaries
        tmp_plot <- list()
        for(par in pars) {
            tmp_plot[[par]] <- stan_plot(model, pars = par)
        }
        gt <- arrangeGrob(grobs = tmp_plot)
        as_ggplot(gt)

    }
}
```
- Results suggest clustering of `x0[]` is justified 



### Males pooled: `y0`, `x0`, and $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson.stan")

stan_model(file="two.piece_qpoisson.stan")

model <- stan(file = "two.piece_qpoisson.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = 1,
                        M = 1,
                        ll = rep(1,N),
                        mm = rep(1,N),                        
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 100,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_pooled.x0_pooled_y0_pooled_theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")

title <- "Two piece Pooled Males"
print(title)
print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)

plot(model, pars = pars, point_est = "mean") #, ggtitle(title)) 
pairs(model, pars = pars_full)
stan_hist(model)

summary(model, probs = c(0.025, 0.5, 0.975))$summary


```
### Separate: `y0`, and `x0`, Pooled  $\theta$

```{r}

iter <- 30000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)

     
model <- stan(file = "two.piece_qpoisson.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        L = n_male,
                        M = 1,
                        ll = data$index, ## y0 and x0
                        mm = rep(1, N), ## Theta
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 10,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              thin = 2,
              verbose = TRUE)
fit_qpoisson_separate.x0_separate.y0_pooled.theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")


print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title)) 

stan_hist(model, pars = "x0")

## Plot sampling for each male
tmp_plot <- list()
for(i in 1:n_male) {
    tmp_pars <-  c(paste0("x0[", i, "]"), paste0("y0[", i, "]"), "lp__")
    tmp_plot[[as.character(i)]] <- pairs(model, pars = tmp_pars)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

#gt <- arrangeGrob(grobs = tmp_plot)
#as_ggplot(gt)

## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

## Doesn't work as expected
## tmp_plots <- stan_plot(model, pars = pars)
## tmp_plots +
##    facet_wrap(vars(grepl("x0.*", params)),
##               scales = "free")

                                        #pairs(model, pars = pars_full)
stan_hist(model, pars = pars)

               
# summary(fit_qpoisson.separate.y0.x0, probs = c(0.025, 0.5, 0.975))$summary

```
- Histograms of `x0[]` suggest following groupings
  1. `c(2, 4, 6, 8, 11)`
  2. `c(1, 3, 5, 7, 10)`
  3. `9` is ambiguous.

### Separate: `y0`, `x0`, and  $\theta$

- didn't run given noise in previous one with separate `y0` and `x0`


### Separate: `y0`, Shared `x0`, and  $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
 cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")

stan_model(file="two.piece_qpoisson_2.0.stan")

model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = 1,
                        Y = n_male,
                        NB = 1,
                        xx = rep(1,N),
                        yy = data$index,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_pooled.x0_separate.y0_pooled.theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")


print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title)) 

## Plot sampling for each male
tmp_plot <- list()
for(i in 1:n_male) {
    tmp_pars <-  c(paste0("x0[", 1, "]"), paste0("y0[", i, "]"), "lp__")
    tmp_plot[[as.character(i)]] <- pairs(model, pars = tmp_pars)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)


## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```
- `y0[7]` should probably be removed from dataset
- Possible groupings for `y0[i]`
  1. $i = 1, 5, 6, (7), 10, 11$
  2. $i = 2, 3, 4, 9$
  3. $i = 8$ could possibly be in either; 2. seems most likely.
  
### Two groups `y0`,  Shared `x0` and Shared $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
#cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")
#stan_model(file="two.piece_qpoisson_2.0.stan")

y0_group <- list()
y0_group[[1]] <- c(1, 5, 6, 7, 10, 11)
y0_group[[2]] <- c(2, 3, 4, 8, 9)

y0_grouping <- map_int(data$index, ~ if_else(. %in% y0_group[[1]], 1, 2))
                     
model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = 1,
                        Y = 2,
                        NB = 1,
                        xx = rep(1,N),
                        yy = y0_grouping,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_pooled.x0_grouped.y0_pooled.theta <- model

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")


print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
plot(model, pars = pars_full) #, ggtitle(title)) 

pairs(model, pars = pars_full)

## Plot sampling for each male
tmp_plot <- list()
## Plot sampling for each male

for(i in 1:2) {
    tmp_pars <-  c(paste0("x0[", y0_group[[i]], "]"),
                   paste0("y0[", i, "]"),
                   "lp__")
    tmp_plot[[as.character(i)]] <- pairs(model, pars = tmp_pars)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)


## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```


### Separate `y0`,  Grouped `x0` and Shared $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
#cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")
#stan_model(file="two.piece_qpoisson_2.0.stan")

x0_group <- list()
x0_group[[1]] <- c(2, 4, 6, 8, 11, 9)
x0_group[[2]] <- c(1, 3, 5, 7, 10)

x0_grouping <- map_int(data$index, ~ if_else(. %in% x0_group[[1]], 1, 2))
                     
model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = 2,
                        Y = n_male,
                        NB = 1,
                        xx = x0_grouping,
                        yy = data$index,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_grouped.x0_separate.y0_pooled.theta <- model
#model <- fit_qpoisson.grouped.y0.and.pooled.x0.theta

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")

print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title))

## Plot sampling for each male
tmp_plot <- list()
for(i in 1:2) {
    tmp_pars <-  c(paste0("x0[", i, "]"), paste0("y0[", x0_group[[i]], "]"),  "lp__")
    tmp_plot[[as.character(i)]] <- pairs(model, pars = tmp_pars)
}

gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)


## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```

- 

### Grouped `y0`,  Grouped `x0` and Shared $\theta$

```{r}

iter <- 15000
tmax <- 46
t0max <- tmax - 1;
t0min <- 25;
## values to use for model predictions
tp = seq(25, tmax, length.out = 100)


## For debugging
#cmodel <- cmdstan_model(stan_file = "two.piece_qpoisson_2.0.stan")
#stan_model(file="two.piece_qpoisson_2.0.stan")

x0_group <- list()
x0_group[[1]] <- c(2, 4, 6, 8, 11, 9)
x0_group[[2]] <- c(1, 3, 5, 7, 10)
x0_grouping <- map_int(data$index, ~ if_else(. %in% x0_group[[1]], 1, 2))

y0_group <- list()
y0_group[[1]] <- c(1, 5, 6, 7, 10, 11)
y0_group[[2]] <- c(2, 3, 4, 8, 9)
y0_grouping <- map_int(data$index, ~ if_else(. %in% y0_group[[1]], 1, 2))

model <- stan(file = "two.piece_qpoisson_2.0.stan",
              model_name = "Two piece qpoisson",
              data=list(x = temp,
                        y = motif_count,
                        N = N,
                        X = length(x0_group),
                        Y = length(y0_group),
                        NB = 1,
                        xx = x0_grouping,
                        yy = y0_grouping,
                        nbb = rep(1,N),
                        xmax = tmax,
                        x0_min = t0min,
                        x0_max = t0max,
                        y_xmax = 0,
                        y0_min = 10,
                        sd_y0_prior = 1000,
                        alpha_theta_prior = 40,
                        ##tp = tp,
                        ## max threshold value.
                        ## having it too close to xmax *sometimes* leads to sampling
                        ## near xmax, but with lower lp and very high E13) b0 values
                        y_xmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = floor(iter/2),
              verbose = TRUE)
fit_qpoisson_grouped.x0_grouped.y0_pooled.theta <- model
#model <- fit_qpoisson.grouped.y0.and.pooled.x0.theta

## Examine output
pars <- c("x0", "y0", "theta")
pars_full <- c(pars, "lp__")

print(model, pars = pars)
#traceplot(model, pars = pars, inc_warmup = FALSE)
#plot(model, pars = pars) #, ggtitle(title))

pairs(model, pars = pars_full)

## Plot parameter estimate summaries
tmp_plot <- list()
for(par in pars) {
    tmp_plot[[par]] <- stan_plot(model, pars = par)
}
gt <- arrangeGrob(grobs = tmp_plot)
as_ggplot(gt)

```
- Results suggest clustering of `x0[]` is justified 

## Model Comparison via LOO

- TODO!

```{r}

ls("fit_qpoisson*")

``
