---
title: "First Fitting of Thermal Models using BRMS"
author: "Michael Gilchrist"
date: "date: 2022-10-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
  )


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=1E5, output = TRUE)

```
# Goal

- Fit series of thermal models in `rTPC` using `brms`.


## Recap

- Previously spent time using glm and glmer models.
- These models had issues with over dispersion when using a poisson family or  converging when using a negative binomial family.

# Current Approach

## Analyze directly with rTPC?

No


According to Padfield et al. (2021)

> However, [the rTPC] pipeline does not accommodate non-independent (related) replicates, and clustered or stratified sampling (possibly with missing values). In such situations, nonlinear mixed effects model fitting (e.g. using the nlme r package; Oddi et al., 2019) or Bayesian approaches (e.g. using the brms r package; Bürkner, 2017) would be more appropriate. Nevertheless, for fitting massive TPC datasets to multiple mathematical models, rTPC offers a simple, reliable and reproducible computational pipeline with robust methods for calculation of model uncertainty, requiring minimal statistical and computational expertise, and suitable for a wide range of applications.

Thus, we can't incorporate individual or random effects.
We could try and use all of the observations in `motif_prop` to deal with the `male` term.




## Analyze with `brms`

Yes.
    
- Can't use `rTPC` functions directly (e.g. `lactin2_1995`) in `brm()` calls since that uses `stan` rather than `R` functions.
- Having issues fitting models, should consider using ouptut from glm fits.


## Data and Explanatory Variables

- Data: Use motif_count from round 3

## Future Plans

### Data and Explanatory Variables

- Use additional rounds when we get `temp_mean` and `humidity_mean` data.
- Include a tensor spline to describe trial order effects.

### Including beak and mass data

- Include beak size (surface area) as additional explanatory variable.
  We do have bird mass as well
- Notes from Liz
  - Do birds with larger beaks maintain singing at higher rates at higher temps? 
  - We calculated bill surface area approximately as the surface area of a cone: length * pi * (width + depth) / 4.
  - Because larger animals produce more heat (Kleiber 1932), we scaled bill size relative to heat production by dividing bill surface area by expected daily energy consumption (mass0.658; Speakman and Kr´ol 2010, Hudson et al. 2013).
  - So, looks like we need to calculate bill surface area and then scale relative to heat production (bill surface area/mass0.658)



### Additional Possibilities

- Follow up with Ray/Juan about ground versus surface temp for operative temperature.
- What is the 0 for operative temperature?
- Ponder utility of other ZF data on panting.

# Set up

## Load libraries

```{r}

## load libraries

library(rstan)
#options(mc.cores = (parallel::detectCores())))
rstan_options(auto_write = TRUE)
if(FALSE) example(stan_model, package = "rstan", run.dontrun = TRUE)


library(brms)

if(FALSE) {
    ## This often causes core dumps
    prior1 <- prior(normal(0,10), class = b) +
    prior(cauchy(0,2), class = sd)
fit1 <- brm(count ~ zAge + zBase * Trt + (1|patient),
            data = epilepsy, family = poisson(),
            prior = prior1, chains = 1, cores = 1)

fit2 <- brm(count ~ zAge + zBase * Trt + (1|patient),
            data = epilepsy, family = poisson(),
            prior = prior1, chains = 2, cores = 1)
    
fit3 <- brm(count ~ zAge + zBase * Trt + (1|patient),
            data = epilepsy, family = poisson(),
            prior = prior1, cores = 2, chains = 2)

}


library(rTPC)
library(nls.multstart)
library(broom)

library(stats)
#library(brms)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggsci) ## provides scale_fill_npg()
library(viridisLite)
library(GGally)
library(reshape2)
library(humidity) ## provides VPD
library(weathermetrics)
library(latex2exp)

```
## Local Functions

```{r}

## Taken from: https://stackoverflow.com/a/51330864/5322644
## Use to get model equations for models in rTPC

help_text <- function(...) {
  file <- help(...)
  path <- dirname(file)
  dirpath <- dirname(path)
  pkgname <- basename(dirpath)
  RdDB <- file.path(path, pkgname)
  rd <- tools:::fetchRdDB(RdDB, basename(file))
  capture.output(tools::Rd2txt(rd, out="", options=list(underline_titles=FALSE)))
}

get_model_eq <- function(model) {
    txt <- help_text(model)
    eqn_line <- grep("^ +rate = .*$", txt, value = TRUE)
    eqn <- gsub("(^ +rate = | *$)", "", eqn_line) %>% gsub("\\.([^0-9])", " * \\1", .)
    df <- tibble(model = model, eq = eqn)
    return(df)
}


plot_brms_fit <- function(brmsfit) {
    plist = list()

    plist[[1]] <- plot(brmsfit, title = paste("brm() summary: ", prior_index), ask = FALSE)
    return(plist)
}

plot_stan_fit <- function(stanfit) {
    plist = list();
    i <- 1
    
    plist[[i]] <-  plot(stanfit, title = paste("stan() summary: ", prior_index))


    ## Plotting routines from: www.weirdfishes.blog/blog/fitting-bayesian-models-with...
    ##
    mack_diagnostics <- rstan::get_sampler_params(stanfit) %>% 
        set_names(1:4) %>% 
        map_df(as_tibble,.id = 'chain') %>% 
        group_by(chain) %>% 
        mutate(iteration = 1:length(chain)) %>% 
        mutate(warmup = iteration <= warmup) %>%
        mutate()
    i <- i+ 1
    
    plist[[i]] <- mack_diagnostics %>% 
        group_by(warmup, chain) %>% 
        summarise(percent_divergent = mean(divergent__ >0)) %>% 
        ggplot() +
        geom_col(aes(chain, percent_divergent, fill = warmup), position = 'dodge', color = 'black') + 
        scale_y_continuous(labels = scales::percent, name = "% Divergent Runs")  + 
        scale_fill_npg()

    i<- i+1
    plist[[i]] <- mack_diagnostics %>% 
        ggplot(aes(iteration, treedepth__, color = chain)) + 
        geom_line() + 
        geom_hline(aes(yintercept = max_treedepth), color = 'red') + 
        scale_color_locuszoom()

    return(plist)

}

```


## Create Model Tibble

```{r}
model_def_tbl <- lapply(get_model_names(), get_model_eq) %>% bind_rows(, .id = NULL)
print(model_def_tbl, n = 200)

```

## Set Output

```{r}

output_dir = "output"
if (!dir.exists(output_dir)) {dir.create(output_dir)}

```


## Load Data

```{r}

## Read in ZEFI Data sets
## Treat 'repeatability' as round = 0
## Add round info

## Repeatability was done between round 1 and 2, female was present, but only one temp. so treating as `round = 2` and redefining `round = 2` as `round = 3`

git_root <- system("git rev-parse --show-toplevel", intern = TRUE)

data_raw = list()

data_raw[[1]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Round-1-Heat-Trials.csv")) %>% mutate(round = 1) %>%
    ## Note T237 and T230 are missing numbers in the motif_count column
    ## so we are filtering these observations out until they are found
    filter(!is.na(motif_count)) %>%
    ungroup()

data_raw[[2]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Repeatability-Song-Count.csv")) %>%
    mutate(round = 2) %>%
    group_by(male) %>%
    mutate(test_order = rank(date)) %>%
    ungroup()

data_raw[[3]] <-read.csv(file.path(git_root, "data", "collated", "HSPi-Round-2-Heat-Trials.csv")) %>%
    mutate(round = 3) %>%
    ## Deal with missing temp_mean and humidity_mean values
    ## in round == 3
    ## 2022/10/19 - code no longer needed
    ## group_by(temp_target) %>% 
    ##mutate(temp = if_else((round == 3 & is.na(temp_mean)),
    ##                      mean(temp_mean, na.rm = TRUE),
    ##                      temp_mean)) %>%
    ##mutate(humidity = if_else((round == 3 & is.na(humidity_mean)),
    ##                          mean(humidity_mean, na.rm = TRUE),
    ##                          humidity_mean)) %>%
    ungroup() 


## Join data and discard empty columns
data_full <- full_join(data_raw[[1]], data_raw[[2]]) %>%
    full_join(data_raw[[3]]) %>%
    discard(~all(is.na(.) | . =="")) %>% ## get rid of columns of only NA
    mutate(trial_completed = !(is.na(motif_count)) ) %>%
    mutate(motif_count = ifelse(is.na(motif_count), 0, motif_count)) %>%
    mutate(motif_count = motif_count*1.0) %>% ## convert to a double so it's not treated as an integer
    mutate(chamber = as.factor(chamber), male = as.factor(male)) %>%
    ## create a global variable trial_order based on individual rounds
    mutate(trial_index = as.integer(round*10+test_order)) %>%
    mutate(motif_count_plus_1 = (motif_count + 1)) %>%
    mutate(log_motif_count_plus_1 = log(motif_count + 1)) %>%
    mutate(temp_target = as.numeric(temp_target)) %>%
    ## Create generic 'temp' column which is either
    ## temp_mean, if it exists, or temp_target, if it doesn't
    mutate(temp = if_else(is.na(temp_mean),
                          temp_target,
                          temp_mean)) %>%
    ## Add column with total motif_count for a given round
    group_by(male, round ) %>%
    mutate(count_total_round = sum(motif_count) ) %>%
    ungroup() %>%
    mutate(motif_prop = motif_count/count_total_round) %>%
    ## assuming poisson error
    ## From glm man page
    ## > Non-‘NULL’ ‘weights’ can be used to indicate that different
    ## >  observations have different dispersions (with the values in
    ## >  ‘weights’ being inversely proportional to the dispersions);
    ## add +1 to deal with single 0
    mutate(count_wt = 1/(motif_count + 1)) %>%
    ## need to rescale wts for motif_prop data
    mutate(prop_wt = count_wt * count_total_round^2) %>% 
    ## Add vpd 
    mutate(svp = SVP(t = temp_mean + 273.15, isK = TRUE), vpd = svp*(1-humidity_mean/100) ) %>%
    group_by(round) %>%
    mutate(vpd_offset = vpd - mean(vpd)) %>%
    ungroup() %>%
    relocate(motif_count, motif_prop, vpd, temp_mean, humidity_mean, .after = male) %>% 
    mutate() ## Dummy function so we can comment out lines above it w/o any issues
```


## Examine Data

```{r}

data_count_total <- data_full %>% group_by(round) %>%
    select(male, round, count_total_round)  %>%
    distinct()

t <- ggplot(data_count_total, aes(count_total_round, fill = male)) +
    geom_histogram(bins = 10) +
    scale_x_log10()

hist_count_total <- t +
    facet_grid(cols =vars(round), scales = "free_x")
hist_count_total



```
## Compare `count_total_round` between round 1 and 3

- See third.fitting.Rmd


### Result

- As before, we see strong consistancy between `round` 1 and 3.
- Consistency with round 2 is weaker, but sample sizes are smaller: 3 trials/male in round 2 vs 6 trials/male in round 3.

  

## Create Filtered Data

```{r}

data_ind <- data_full %>%
    filter(
    (round == 3 & count_total_round >=150) |
    (round == 2 & count_total_round > 30)
    )

dim(data_ind)
## copy data frame and assign `male =  "combined")
data_comb <- data_ind %>% mutate(male = "combined")

data <- bind_rows(data_ind, data_comb)


```


## Plot Filtered Data
```{r}

xlab <- "Temperature"
ylab <- "motif_count"

plot_temp_data <-
    ggplot(data) + 
    aes(x = temp,
        y = motif_count) +
    facet_wrap("male", scales = "free_y") +
    geom_point() +
labs( title = paste( ylab, " vs ", xlab))
last_plot()

```



# First Analysis

## Formal Model Fits to `motif_count`


### Data Set Up

- Filter data

```{r}

## Filter the data a bit more to make things easier
males_in_round_3 <- data_ind %>% select(male, round) %>% unique() %>% filter(round == 3) %>% pull(male)

data <- data_ind %>%
    ##    filter( (male %in% c("T229", "T234", "T244", "T243","T247", "T258"))) %>%
    filter(male %in% males_in_round_3) %>% ## This is apparently unnecessary
    mutate()

dim(data)
```

## Fit `modifiedgaussian`

- rTPC model: `modifiedgaussian`
- `male` as a factor

```{r}

fit.file <- file.path(output_dir, "mgauss_nb.Rda")


refit.model <- FALSE

nlform <-
    bf(motif_count ~ rmax*exp(-0.5*(fabs(temp - topt)/a)^b),
       rmax ~ male -1,
       topt ~ 1,
       a ~ 1,
       b ~ 1,
       nl = TRUE)


iter = 10000
warmup = floor(iter/3)
max_treedepth = 15

if(!file.exists(fit.file) | refit.model) {

    fit_list = list()


    for(prior_index in 1:2) {

        print(paste("fitting prior_index: ", prior_index))
        
        switch(prior_index,
        {
            ## bounded normals
            nlprior <- c(prior(normal(400, 400), nlpar = "rmax", lb = 0),
                         prior(normal(35, 20), nlpar = "topt", lb = 0, ub = 45),
                         prior(normal(0, 3), nlpar = "a", lb = 0),
                         prior(normal(0, 3), nlpar = "b", lb = 0))},
        {
            ## wider bounded normals
            nlprior <- c(prior(normal(400, 800), nlpar = "rmax", lb = 0),
                         prior(normal(35, 20), nlpar = "topt", lb = 0, ub = 45),
                         prior(normal(0, 10), nlpar = "a", lb = 0),
                         prior(normal(0, 6), nlpar = "b", lb = 0))},
        {
            ## gamma(alpha = a , beta = b); E(x) = a/b; var(x) = a/b^2 
            nlprior <- c(prior(normal(400, 800), nlpar = "rmax", lb = 0),
                         prior(normal(35, 20.), nlpar = "topt", lb = 0),
                         prior(gamma(1, 0.01), nlpar = "a", lb = 0),
                         prior(gamma(1, 0.01), nlpar = "b", lb = 0))},
        )



        fit_list[[prior_index]] <- brm(formula = nlform,
                                        data = data,
                                        family = negbinomial(link = "identity",
                                                             link_shape = "identity"),
                                        prior = nlprior,
                                        chains = 4, cores = 4,
                                        iter = iter,
                                        warmup = warmup,
                                        save_model = "model_mgauss_nb_1.stan",
                                        control = list(adapt_delta = 0.96))


    }

    mgauss_nb <- fit_list
   
    save(mgauss_nb, fit.file)
} else {
    load(fit.file)
}




```

### Plot Fits

```{r, eval=TRUE}


for(prior_index in 1:2) {
    model <- mgauss_nb[[prior_index]]
    plot_brms_fit(model)
    print(plot_stan_fit(model$fit))
}
    

```
# End

```{r, error = FALSE}

knitr::knit_exit()

```

## Fit `lactin1`

- `male` as a factor
- fixed `tmax = 45` 
- Set `b = 0` (so essentially lactin1 which removes an unnecessary scaling term in the logan model).
```{r, eval=FALSE}

refit.model <- FALSE
tmax = 45
b = 0

nlform <-
    bf(motif_count ~  exp(a * temp) - exp(a * tmax - ((tmax - temp)/deltat)) + b,
       deltat ~ male - 1,
       a ~ male -1,
       nl = TRUE)

iter = 20000
warmup = floor(iter/2)
max_treedepth = 14

lactin1_nb = list()

### THESE PRIORS AREN'T CORRECT

for(prior_index in 1:2 ){

    switch(prior_index,
    {
        ## bounded normals
        nlprior <- c(prior(normal(400, 400), nlpar = "a", lb = 0),
                     prior(normal(35, 20), nlpar = "topt", lb = 0, ub = 45),
                     prior(normal(0, 3), nlpar = "a", lb = 0),
                     prior(normal(0, 3), nlpar = "b", lb = 0))},
    {
        ## wider bounded normals
        nlprior <- c(prior(normal(400, 800), nlpar = "rmax", lb = 0),
                     prior(normal(35, 20), nlpar = "topt", lb = 0, ub = 45),
                     prior(normal(0, 10), nlpar = "a", lb = 0),
                     prior(normal(0, 6), nlpar = "b", lb = 0))},
    {
        ## gamma(alpha = a , beta = b); E(x) = a/b; var(x) = a/b^2 
        nlprior <- c(prior(normal(400, 800), nlpar = "rmax", lb = 0),
                     prior(normal(35, 20.), nlpar = "topt", lb = 0),
                     prior(gamma(1, 0.01), nlpar = "a", lb = 0),
                     prior(gamma(1, 0.01), nlpar = "b", lb = 0))},
    )



    lactin1_nb[[prior_index]] <- brm(formula = nlform,
                                    data = data,
                                    family = negbinomial(link = "identity",
                                                         link_shape = "identity"),
                                    prior = nlprior,
                                    chains = 4, cores = 4,
                                    iter = iter,
                                    warmup = warmup,
                                    save_model = "model_mgauss_nb_1.stan",
                                    control = list(adapt_delta = 0.96))


}

```
### Plot Fits

```{r, eval=FALSE}


for(prior_index in 1:2 ){
    model <- lactin1_nb[[prior_index]]
    ## plot_brms_fit(model)
    print(plot_stan_fit(model$fit))
    }
    

```
