---
title: "Visualize Piecewise Regression with Negative Binomial Type I Error on Real Data using `brms` Custom Family"
author: "Michael Gilchrist"
date: "Created: 2023-03-29; Compiled: `r date()`"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE,  # show R code
  fig.width = 7.5,
  fig.height = 10
)


if(interactive()) {
  default::default(.ess.eval) <- list(max.deparse.length=2E2, output = TRUE)
  output_dir <- "output"  
} else { 
  output_dir <- file.path("output", "render")
}

input_dir <- file.path("input")
```
# Goal

- Visualize two piece negative binomial type 1 formulation to data

## Recap

- Update to `2023-03-09_visualize.2023-02-28-fit...` for use with more recent fits in `2023-03-18_fit...` folder.
- Part of challenge is bimodal nature of some male's `x0` and `y0` fits
  - `T235` and `T236` in particular




## Insights


# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(ggExtra)
library(cowplot)
library(GGally)
library(RColorBrewer) ## needed to have more than 8 colors with `palette="Set2"
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(brms)
library(bayesplot)
library(tidybayes)
library(loo)
library(modelr)
library(reshape2) ## provides melt

## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(latex2exp)

ggplot2::theme_set(theme_default(base_size = 10))
## Restore theme settings
## ggplot2::theme_set(theme_default())

n_cores <- 4

color_scheme_set("viridis")

```
## Load Local Functions


```{r}

source("../Local.Functions/local.functions_ZFI.fittings.R")

```

## Source family

```{r}

```

## Load `input`

```{r}

infile <- last(dir(file.path(input_dir, "tibbles"), "fit_tbl.*"))
## Override default fo most recent fits
#infile <- "fit_tbl_adapt-delta-0.90.Rda"
infile_tbl <- file.path(input_dir, "tibbles", infile)
load(infile_tbl, verbose = TRUE)
## save filtered tibble incase it becomes corrupted locally
fit_tbl_orig <- fit_tbl
fit_tbl <- fit_tbl_orig %>% filter(!(is.na(fit))) 

```

### Add columns to `fit_tbl` for future use
```{r}

names_col <- c("obs", "epred", "pred", "plots")
fit_tbl <- add_column_safely(fit_tbl, names_col)

```

## Define and/or Expose Functions used by STAN

```{r}

source("../../../custom-brms-families/families/nbinom_type1.R")

## Define fmax() in case expose_functions()
fmax <- function(x, x0) max(x, x0)

## Create functions defined in stan's function header
## Only need to do this once since all fits use the same user
## defined functions such as "two_piece"
## Only do this if not already defined or forcing
force_expose_functions <- FALSE
if(!exists("two_piece") | force_expose_functions) expose_functions(fit_tbl[[1, "fit"]][[1]], vectorize = TRUE, show_compiler_warnings = FALSE)

```


## Generate Columns of Observed Data, Predicted Data and Expected Values 

- Moved to separate file

```

# Visualize Model Fits


## Set up functions, parameters, and results tibble 

```{r}

save_plots = TRUE
force_update_plots = TRUE

xmax <- 45.9

## Here I create a variable using a string and then access it using a string
for(name in names(fit_tbl)[1:8]) {
  # create
  var <- paste0(name, "_vec")
  assign(var, pull(fit_tbl, name) %>% unique(.))
  # access via get()
  cat(paste0(var, ": ", paste0(get(var), sep = "\n\t")))
}

```

### Print fits

```{r, message = FALSE }

pairs_include_lp = TRUE
n_rows_print_half <- 4 ## 1/2 of rows to print of prior
use_plot_grid <- FALSE ## if true print grid, if false print row
print_full_prior <- FALSE
off_diag_fun <- "scatter" # "scatter" or "hex" for mcmc_pairs
#use_mcmc_pairs <- FALSE ## indicate if to make scatter plots using mcmc_pairs

print_mcmc_pairs <- TRUE # individual level parameters
print_mcmc_pairs_group <- TRUE # lp and group level parameters
print_mcmc_scatter <- TRUE # individual level parameters, histograms printed on margins
print_stan_hist <- FALSE

print_get_prior <- TRUE ## 
print_prior_summary <- TRUE

for(row_index in 3) {#1:nrow(fit_tbl)) {

  cat(paste("Row: ", row_index))

  fit_row <- fit_tbl[row_index, ]
  fit_brms <- fit_row$fit[[1]]
  fit_stan <- fit_brms$fit
  data <- fit_brms[["data"]]
  

  chains_n <- length(dimnames(fit_stan)$chains)

  desc <- fit_row$desc

  desc_filename <- gsub("_", "-", desc) %>%
    gsub("; ", "_", .) %>%
    gsub(":? ", "-", .)
  desc_short <- desc %>% str_replace("nbinom_type1; two_piece; ","")

  ## two ways of specifying a title
  ## Second is more 'automatic'
    title_row <- ggdraw() + draw_label(desc_short, fontface='bold', size = 12)

  ## Print and plot results, regardless of which fits one uses           
  print(desc)
  # print(filename_desc)

  if(print_prior_summary) {
    print("Fit Prior Information")
    if(print_full_prior) {
      print(prior_summary(fit_brms)) # %>% filter(nlpar!="y0"))
    } else {
      ## Condensed version
      tmp_prior <- prior_summary(fit_brms)
      #class(tmp_prior) <- "data.frame"
      #tmp_prior <- tmp_prior %>% select(-source)
      rows_print <- unique(c(1:n_rows_print_half, nrow(tmp_prior) - (n_rows_print_half):0))
      tmp_top <- tmp_prior[1:n_rows_print_half, ] 
      tmp_bottom <- tmp_prior[rows_print[-(1:n_rows_print_half)], ]
      tmp_join <- bind_rows(tmp_top, tmp_bottom)
      tmp_join[ n_rows_print_half + 1, ] <- "..."
      ## remove brms class so we can select
      class(tmp_join) <- "data.frame"
      print(tmp_join %>% select(-source))
    }
#    print_prior_summary <- TRUE
  }

  print("Fit Information")
  print(summary(fit_brms) ) # %>% gsub("disp_value", fit_row[["disp_value"]]), .) #, pars = "x0*"))          %>% filter(nlpar!="y0"))


  #clean up variable names
  fit_stan_rename <-
    fit_stan %>%
    clean_var_names()
  
  ## 

  vars_clean <- names(fit_stan_rename) %>% na.omit(.)


  male_vec <- unique(data$male) %>% as.character(.)
  ## get male specific vars (start with "T")
  vars_T <- grep("^T[0-9]{3}", vars_clean, value = TRUE)
  vars_Intercept <- grep("Intercept", vars_clean, value = TRUE)
  vars_non_T <- vars_clean[!(vars_clean %in% c(vars_T, vars_Intercept))]
  #print(vars_clean)

  ## Examine priors and exclude any constants from vars_non_T
  priors_tmp <- prior_summary(fit_brms)
  constants_non_T <- ggdist::parse_dist(priors_tmp) %>% filter(grepl("constant", .dist)) %>% pull(class)
  vars_non_T <- vars_non_T[!(vars_non_T %in% constants_non_T)]
  


  
  ## Count occurence of each male in model fit_brms
  male_instance <- sapply(male_vec, function(x) {sum(str_detect(x, string=vars_clean))})

  ## Check to make sure plotting will work and/or is desired
  if((all(male_instance > 1) | pairs_include_lp) & print_mcmc_pairs) {
    ## Use mcmc_pairs
    pairs_list <- list()
    ## update panel.boarder
    ## Note brms uses 'variable' while stanfit uses 'pars'
    for(male in male_vec) {
      # print(male)
      vars_male <- grep(male, vars_T, value = TRUE)

      ## Generate pairs() plot using mcmc_pairs()
      ## Don't use pairs.stanfit which does not return an object
      pairs_pars <- vars_male
      if(pairs_include_lp) {
        pairs_pars <- c(pairs_pars, "lp_")
      }


      if(TRUE) { ## use mcmc_pairs not ggplot2 scatterplots
      ## Change text size
      # bayesplot_theme_update(text = element_text(size = 6)) #, family = "sans"))
      pairs_tmp <-  mcmc_pairs(
        fit_stan_rename,
        pars = pairs_pars,
        condition = pairs_condition(
          # plot all chains on lower diagonal
          # Could also split by draws
          chains <- list(top = 1:(chains_n), bottom = NULL)
        ),
        ## I believe div_color means color of divergent samples
        ##np_style = pairs_style_np(div_color = chain)
        ## set point size and transparency
        # grid_args = list(text = element_text(size = 20)),
        # off_diag_args are passed to mcmc_scatter() or mcmc_hex
        # off_diag_args = list(color = chain)        
        ## Try to use hex densities        
        off_diag_fun = if_else(off_diag_fun == "hex", "hex", "scatter") ## "hex" won't likely work if trying to visualize separate chains

      )

      #        bayesplot_theme_set() # same as bayesplot_theme_set(theme_default())

      # add density curves to off diagonal
      plots_tmp <- pairs_tmp$bayesplots
        
        ## get list of off diagonal positions
        n_tmp <- sqrt(length(plots_tmp))
        ## plot_index = lower diagonal 
        plot_index <- sapply(1:(n_tmp-1), function(x) (x+1):n_tmp+(x-1)*n_tmp) %>% unlist()

        xy_pars <- combn(pairs_pars, 2) %>% t() %>% as_tibble()
        names(xy_pars) <- c("y", "x")

        xlab_vec <-  xy_pars %>% transmute(x = gsub("T[0-9]{3}_","", x)) %>% unlist()
        ylab_vec <- xy_pars %>% transmute(y = gsub("T[0-9]{3}_","", y)) %>% unlist()
        ## These don't work
        #xlab_vec <- lapply(pairs_pars, rep, times = n_tmp) %>% unlist() %>% gsub("T[0-9]{3}_","", .) 
        #ylab_vec <- rev(xlab_vec)

      } else { ## use ggplot2 scatterplots, not mcmc_pairs
        ## Ideally this woudl eventually be used so that we can color code chains
        
        plots_tmp <- list()
        ## get df of plots to make based on xy axes
        xy_pars <- combn(pairs_pars, 2) %>% t() %>% as_tibble()
        names(xy_pars) <- c("y", "x")

        xy_labs <-  xy_pars %>% mutate(x = gsub("T[0-9]{3}_","", x), y = gsub("T[0-9]{3}_","", y))
        #xy_labs <-  xy_pars %>% across(x = gsub("T[0-9]{3}_","", x), y = gsub("T[0-9]{3}_","", y))
        ## This 'works', but I can't modify the figure
        plot_tmp <- apply(xy_pars, 1, function(x) {
          mcmc_scatter(fit_stan_rename, pars =  x)}
          )                  

      }
      
      plots_row <- list()

      ## add contour lines
      for(j in 1:length(plot_index)) {
        i = plot_index[[j]]
        # pars = c(vars_male): list of 4 ggplot objects diagonal plots are 2 and 3
        plots_tmp[[i]] <- plots_tmp[[i]] +
             stat_density_2d(color = if_else(off_diag_fun == "hex", "black", "red"),
                          size = .75,
                          bins = 5) +
          labs(x = xlab_vec[[j]], y = ylab_vec[[j]])         

        plots_row[[j]] <- plots_tmp[[i]]
        #ggExtra::ggMarginal(plots_tmp[[i]] +
        #                                        geom_point(col="transparent") +
        #                                        geom_hex() +
        #                                        theme(legend.position = "none"), type = "histogram")
      }

      
      #text_axes <- str_remove(pairs_pars, "T[0-9]+_")
      labels_subfig <- str_extract(pairs_pars, "T[0-9]+")[[1]]
      
#      title_subfig <- ggdraw() + draw_label(text_subfig, fontface='bold')

      
      if(use_plot_grid) {
        pairs_male <-bayesplot_grid(plots = plots_tmp) ## Change font size
      } else {
        ## Print summaries as a row
        pairs_male <- plot_grid(
          plotlist = plots_row,
          #labels = labels_subfig,
          #label_size = 8,
          # hjust = 0,  # don't understand why this is so large
          # vjust = 0, These don't work as expected
          label_fontface = "plain",
          nrow = 1) +
          labs(tag = labels_subfig) +
          theme(plot.tag = element_text(size = rel(1), face = "bold", vjust = 1))
      }
      pairs_list[[male]] <- pairs_male
    } # end for(male in male_vec)

    ## Combine plots based on grid or row
    if(use_plot_grid) {
      ## Set number of males per row
      ifelse(length(pairs_pars) > 2, ncol_row <- 1, ncol_row <- 2)
      ## Set number of rows/page
      ifelse(length(pairs_pars) > 2, nrow_page <- 4, nrow_page <- 5)

      p <- plot_grid(plotlist = pairs_list)

      plot_pairs_list <- plot_grid(p,
                                   labels = desc_short,
                                   ncol = ncol_row,
                                   label_size = 10,
                                   hjust = 0,
                                   vjust = - 0.75) +
        # Add some space around the edges  
        theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))
    } else {  ## use_plot_grid is false
      ## Put all plots on the same row
      
      ## Set number of males per row
      ifelse(length(pairs_pars) > 3, ncol_row <- 1, ncol_row <- 2)
      ## Set number of rows/page
      ifelse(length(pairs_pars) > 3, nrow_page <- 4, nrow_page <- 5)
      
      plot_pairs_list <- marrangeGrob(pairs_list, ncol = 1, nrow = nrow_page,
                                      top = desc_short,
                                      bottom = quote(paste("page", g, "of", npages))
                                      )
      
    }


    print(plot_pairs_list)
    if(save_plots) {
      filename <- paste0("plot-pairs_", desc_filename, ".pdf")
      ggsave(filename = filename, path = file.path(output_dir, "figures"),
             plot = plot_pairs_list,
             width = 8, height = 11, units = "in",
             scale = 1,
             dpi=300,
             bg = "white")
    }  
  } ## end if(all(male... for mcmc_pairs

  if(all(male_instance > 1) & print_mcmc_scatter) {

    scatter_list <- list()
    for(male in male_vec) {
      # print(male)
      vars_male <- grep(male, vars_T, value = TRUE)
     # if(length(vars_male) > 2) vars_pair <-  
      scatter_tmp <-  mcmc_hex( #was mcmc_scatter
        fit_stan_rename,
        ## Can only use two variables
        pars = c(first(vars_male), last(vars_male)) #, vars_Intercept),
      )
      
      ## ggMarginal doesn't work natively with mcmc_hex, so we need to make the
      ##  points transparent and then add a hex layer
      scatter_list[[male]] <- ggExtra::ggMarginal(scatter_tmp +
                                                    geom_point(col="transparent") +
                                                    geom_hex() +
                                                    theme(legend.position = "none"), type = "histogram")
      
    }
     
    p <- plot_grid(plotlist = scatter_list,
                   ncol = 3)
    
    plot_scatter_list <- plot_grid(title_row, p, ncol = 1, rel_heights=c(0.1, 1))

    print(plot_scatter_list)
    
    filename <- paste0("plot-scatter_", desc_filename, ".pdf")
    ggsave(filename = filename, path = file.path(output_dir, "figures"),
           width = 8, height = 11, units = "in",
           scale = 1,
           dpi=300,
           bg = "white")

    
      } # end if(print_mcmc_scatter)

  

  
  if(print_mcmc_pairs_group | length(vars_non_T > 1)) {
    pairs_list2 <- list()
    
    if(length(vars_Intercept) > 0) {
      pairs_list2[["Intercept"]] <- mcmc_pairs(
        fit_stan_rename, pars = c(vars_Intercept, "lp_"),
        off_diag_fun = c("hex"))
    }

    pairs_list2[["non-T"]] <- mcmc_pairs(fit_stan_rename,
                                         pars = vars_non_T,
                                         off_diag_fun = c("hex")
                                         ) 

    
    p <- plot_grid(plotlist = pairs_list2,
                   ncol = 1)
    plot_pairs_list2 <- plot_grid(title_row, p,  ncol = 1, rel_heights=c(0.1, 1))
      
    print(plot_pairs_list2)

        filename <- paste0("plot-pairs2_", desc_filename, ".pdf")
    ggsave(filename = filename, path = file.path(output_dir, "figures"),
           width = 8, height = 11, units = "in",
           scale = 1,
           dpi=300,
           bg = "white")
    
  }



  vars_tmp <- vars_clean %>% str_subset("y0");
  
  stan_plot(fit_stan_rename, pars = vars_tmp) +
    ggtitle("Initial Motif Counts", subtitle = desc)

  vars_tmp <- vars_clean %>% str_subset("T[0-9]+_x0")
  if(length(vars_tmp) > 0) {
    fit_stan <- stan_plot(fit_stan_rename, pars = vars_tmp) +
      ggtitle("Thresholds", subtitle = desc_short)
    
    filename <- paste0("fit-stan_", desc_filename, ".pdf")
    ggsave(filename = filename, path = file.path(output_dir, "figures"), dpi=300)
  }
  
  if(print_stan_hist) {
    ncol <- 4 
    hist <-  stan_hist(fit_stan_rename,
                       pars = vars_fit,
                       bins = 25,
                       ncol = ncol) +
      ggtitle(desc_short)
  }
  if(save_plots) {
    filename <- paste0("histogram_", desc_filename, ".pdf")
    ggsave(filename = filename, path = file.path(output_dir, "figures"), dpi=300)
  }
  
    
  obs <- fit_row[["obs"]][[1]]


  print("Plotting Expectations")

  epred <- fit_row[["epred"]][[1]]
  plot_curr <- fit_tbl[[curr_row, "plots"]][[1]][["epred"]]
  if(length(plot_curr) ==0 | force_update_plots) {
    plot_tmp <- plot_epred(data_epred = epred, male_vec = male_vec, desc = desc, data_obs = obs)
    print(plot_tmp)
    ## Why do I need the [[1]]?
    fit_tbl[[curr_row, "plots"]][[1]][["epred"]] <- plot_tmp
    
    if(save_plots){
      filename <- paste0("data.and.epred-vs-x_", desc_filename, ".pdf")
      ggsave(filename = filename, plot = plot_tmp, path = file.path(output_dir, "figures"),
             width = 8, height = 11, units = "in",
             scale = 1,
             dpi=300,
             bg = "white")
    }
  }

  ## Plot Data and Predicted Values vs. Temperature

  ## Reserve term "simulated" for when uisng the best fit MODEL
  ## Warning: Removed XXX rows containing missing values (`stat_slabinterval()`).
  ## This is from ggplot2 and indicates there's data outside the y range
  print("Plotting predictions")
  pred <- fit_row[["pred"]][[1]]
  
  plot_curr <- fit_tbl[[curr_row, "plots"]][[1]][["pred"]]
  if(length(plot_curr) == 0 | force_update_plots) {
    plot_tmp <- plot_pred(data_pred = pred, male_vec = male_vec, desc = desc_short, data_obs = obs)
    print(plot_tmp)
  
    fit_tbl[[curr_row, "plots"]][[1]][["pred"]] <- plot_tmp
    if(save_plots) {
      filename <- paste0("data.and.pred-vs-x_", desc_filename, ".pdf")
      ggsave(filename = filename, plot = plot_tmp, path = file.path(output_dir, "figures"),
             width = 8, height = 11, units = "in",
             scale = 1,
             dpi=300,
             bg = "white")
    }
  }

  
  #mcmc_trace(posterior, pars = c("lp__"))
  
}
```


## Exit rendering
```{r}

knitr::knit_exit()

```

# CRUFT!!!!!!

# Plot Data and Predictions

## Set up colors -- not currently needed

```{r, eval = FALSE}

## "Set2" only has 8 colors by default.
## This code expands that ability, based on: https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
n_colors <- 11 # length(male_vec)
colors_male <- colorRampPalette(brewer.pal(8, "Set2"))(n_colors)
```

Data suggests that disp should vary between males?




## Plot Predicted and Expected Values vs. Temp

```{r}

curr_row_max <- nrow(fit_tbl)
curr_row_max <- 1
save_plots = TRUE
force_update_plots = TRUE

for(curr_row in 1:curr_row_max) {

  fit_row <- fit_tbl[curr_row, ]
  

  desc <- fit_row$desc

  desc_filename <- gsub("_", "-", desc) %>%
    gsub("; ", "_", .) %>%
    gsub(":? ", "-", .)
  desc_short <- desc %>% str_replace("nbinom_type1; two_piece; ","")

  print(paste0("Row: ", curr_row, ", ", desc))
  
  obs <- fit_row[["obs"]][[1]]
  if(class(obs$male) == "character") {
    males <- unique(obs$male)
  } else {
    males <- levels(obs$male) %>% unique()
  }

  print("Plotting Expectations")

  epred <- fit_row[["epred"]][[1]]
  plot_curr <- fit_tbl[[curr_row, "plots"]][[1]][["epred"]]
  if(length(plot_curr) ==0 | force_update_plots) {
    plot_tmp <- plot_epred(data_epred = epred, male_vec = males, desc = desc_short, data_obs = obs)
    print(plot_tmp)
    ## Why do I need the [[1]]?
    fit_tbl[[curr_row, "plots"]][[1]][["epred"]] <- plot_tmp
    
    if(save_plots){
      filename <- paste0("data.and.epred-vs-x_", desc_filename, ".pdf")
      ggsave(filename = filename, plot = plot_tmp, path = file.path(output_dir, "figures"),
             width = 8, height = 11, units = "in",
             scale = 1,
             dpi=300,
             bg = "white")
    }
  }

## Plot Data and Predicted Values vs. Temperature

## Reserve term "simulated" for when uisng the best fit MODEL
  ## Warning: Removed XXX rows containing missing values (`stat_slabinterval()`).
  ## This is from ggplot2 and indicates there's data outside the y range
  print("Plotting predictions")
  pred <- fit_row[["pred"]][[1]]
  
  plot_curr <- fit_tbl[[curr_row, "plots"]][[1]][["pred"]]
  if(length(plot_curr) == 0 | force_update_plots) {
    plot_tmp <- plot_pred(data_pred = pred, male_vec = males, desc = desc, data_obs = obs)
    print(plot_tmp)
  
    fit_tbl[[curr_row, "plots"]][[1]][["pred"]] <- plot_tmp
    if(save_plots) {
      filename <- paste0("data.and.pred-vs-x_", desc_filename, ".pdf")
      ggsave(filename = filename, plot = plot_tmp, path = file.path(output_dir, "figures"),
             width = 8, height = 11, units = "in",
             scale = 1,
             dpi=300,
             bg = "white")
    }
  }
}

```







## OLD Plot Data and Simulated Values vs. Temperature

```{r}


for(curr_row in 1:curr_row_max) {
  
  curr_desc <- fit_tbl[[curr_row, "desc"]]
  curr_desc_short <- curr_desc %>% str_replace("nbinom_type1; two_piece; ","")
  fit_brms <- fit_tbl[[curr_row, "fit"]][[1]]

  if(FALSE) {
    object.size(fit_brms)
    ## Unclear where function information is stored
    ## Fit object does not seem to change in size
    expose_functions(fit_brms, vectorize = TRUE, show_compiler_warnings=FALSE)
    object.size(fit_brms)
  }
  
  data_obs <- fit_brms$data


  ## Create grid of x values for epred/predictions

  #dataframe_tmp <- crossing(x = seq_range(c(20, 30), n = 51), draw = 1:3, male = male_vec)
  
  data_grid <- data_obs %>%
    group_by(male) %>%
    data_grid(x = seq_range(c(20, 45.9), n = 51)) %>%  #, .model = fit_brms) %>% 
    ungroup() 

  
  

  ## add simulated values
  ## Only getting 1 draw/(male temp)
  data_pred <- data_grid %>%
    add_predicted_draws(object = fit_brms)

  
  y_max <- max(data_pred$.prediction, na_rm = TRUE)*1.1

  plot_tmp <- ggplot(data = data_pred, aes(x = x, y = .prediction, color = male)) +
    ## Combine Scatter Plots and Model vs Data Plots
    stat_lineribbon(aes(y=.prediction), .width = c(.95), color = "#08519C") +
    scale_fill_brewer(palette = "Greys") +
    scale_color_manual(values = colors_male) +
    #scale_color_brewer(palette = "Set2") +
    geom_point(data = data_obs,
               aes(x = x, y = y), color = "red") + 
    ylim(0, y_max)

  plot_data_vs_pred <- plot_tmp + facet_wrap(vars(male)) +
    labs(title = "Data vs. Simulated Values", subtitle = curr_desc_short) 
  last_plot()

  filename <- paste0("data-simulated-vs-x_", desc_filename, ".pdf")
  ggsave(filename = filename, path = file.path(output_dir, "figures"),
         width = 8, height = 11, units = "in",
         scale = 1,
         dpi=300,
         bg = "white")
}
```


## Exit rendering
```{r}

knitr::knit_exit()

```




## Create Model vs. Data Plots

```{r}

plot_model <- list()

list_plot_mean <- list()
  
for(male in male_vec) {
    
  tmp_data <- data_obs %>% filter(male == !!male)
  motif_count <- tmp_data$y
  tmp_mean <- data_epred %>% filter(male == !!male)

  tmp_plot_mean <- ggplot(data = tmp_mean,
                            aes(x = x, y = motif_count)) +
    ylim(0, max(motif_count)*1.1) +
    stat_lineribbon(aes(y = .epred)) +
    scale_fill_brewer(palette = "Greys") +
    scale_color_brewer(palette = "Set2") +
    geom_point(data = tmp_data, aes(x = x, y = y), color = "Red") +
    labs(title=male) +
    theme(legend.position="none")
  #print(tmp)
  list_plot_mean[[male]] <- tmp_plot_mean
  
}

tmp_plot <- plot_grid(
  plotlist = list_plot_mean,
  ncol = 4
)

data_grid <- data_obs %>%
  group_by(male) %>%
  data_grid(x = seq_range(c(20, 40), n = 51))

plot_model[["mean"]] <- tmp_plot

tmp_plot

filename <- paste0("data-epred-vs-temp_", desc_filename, ".pdf")
ggsave(filename = filename, path = file.path(output_dir, "figures"),
       width = 8, height = 11, units = "in",
       scale = 1,
       dpi=300,
       bg = "white")
```

```{r}

list_plot_data <- list()

for(male in male_vec) {


  tmp_data <- data_obs %>% filter(male == !!male)
  motif_count <- tmp_data$y
  tmp_pred  <- data_pred %>% filter(male ==!!male)
  tmp_plot_data <- ggplot(data = tmp_pred,
                          aes(x = x, y = motif_count)) +
    ylim(0, max(motif_count)*1.1) +
    stat_lineribbon(aes(y = .prediction),
                    .width = c(.95, .80, .50),
                    alpha = 1/4) +
    geom_point(data = tmp_data, aes(x = x, y = y), color = "Red") +
    scale_fill_brewer(palette = "Greys") +
    #scale_color_brewer(palette = "Set2") +
    labs(title=male) +
    theme(legend.position="none")
  print(tmp_plot_data)
  list_plot_data[[male]] <- tmp_plot_data
}

  
  #gridExtra::grid.arrange(grobs = plot_model_list)
  #
  #plot_tmp <- plot_grid(plotlist = plot_model_list,
  #                  ncol = 3)


plot_grid(
  plotlist = list_plot_data,
  ncol = 4,
  title = "data"
)


plot_model[["data"]] <- tmp_plot_data




```

## Combine Scatter Plots and Model vs Data Plots

```{r}

plot_combo <- list()

for(prediction in c("data")) {

    list_tmp <- list()
    for(male in male_vec) {
      list_tmp[[male]] <-
        plot_grid(
          plotlist = list(
            scatter_list[[male]],
            plot_model[[prediction]][[male]] +
              labs(title = NULL) +
              xlab(NULL) +
              ylab(NULL)
          )
        )  + labs(title = male)
    }

  plot_tmp <- plot_grid(
      plotlist = list_tmp,
      ncol = 2,
      title = prediction
    )
    print(plot_tmp)
    plot_combo[[prediction]] <- plot_tmp

}

```

## Exit rendering
```{r}

knitr::knit_exit()

```

### Identify Best Chains __ Obsolete

```{r, eval = FALSE}

if(FALSE) {
  for(row_index in 1:row_index_max) {
    fit_tibble <- melt(fit_array) %>% as_tibble()
    lp_tbl <- fit_tibble %>% filter(parameters == "lp__")

    lp_stats <- lp_tbl %>%
      group_by(chains) %>%
      summarize(mean = mean(value), sd = sd(value), n = n(), se = sd/sqrt(n) ) %>%
      arrange(mean)

    extract_best_flag <- TRUE
    if(extract_best_flag) {
      
      ## Find chain with greatest mean lp__ value and calculate SE on mean
      mean_max <- lp_stats[[1, "mean"]]
      se <- lp_stats[[1, "se"]]
      ## Use a larger threshold to avoid 1/20 being dropped due to type 1 errors
      threshold <- mean_max - 4 * se

      ## Flag chains with matching means to max
      lp_stats <- lp_stats %>%
        mutate(match_max = mean > threshold) %>%
        arrange(chains)

      # fit_tbl[[row_index, "lp_stats"]] <- lp_stats
      
      chains_best <- filter(lp_stats, match_max) %>%
        pull(chains) %>%
        sub("chain:", "", .) %>%
        as.numeric()
      # fit_tbl[[row_index, "chains_best"]] <- list(chains_best)
      n_chains_best <- length(chains_best)
      n_chains <- levels(fit_tibble$chains) %>% length()

      print(desc_short)
      print(paste("Chains Best:", paste(chains_best, collapse = ", "), " (", n_chains_best, "out of", n_chains, ")" ))

      if(n_chains != n_chains_best) {
        print("Dropping non-best chains")
      }

      if(n_chains_best > n_chains_target) {
        ## just take the first n target Chains
        print("Dropping extra best chains")
        chains_best <-  sample(chains_best, n_chains_target)
      }

      ## Keep only desired chains
      fit_brms_best <- keep_chains(fit_brms, chains_best)
      ## update tbl
      fit_tbl[[row_index, "fit_best"]][[1]] <- fit_brms_best
    }
  }
}

```

### Print Posterior Probability Traces

- Most of this code has been moved to the fitting routines.

```{r}

n_chains_target <- 4

row_index_max <- nrow(fit_tbl)
row_index_max <- 3


save_plots <- FALSE
save_plots_new <- save_plots ## redundant?
render_plots_force <- FALSE
render_plots <- TRUE

for(row_index in 1:row_index_max) {

  cat(paste("fit_tbl Row: ", row_index, "\n\n"))
  
  fit_row <- fit_tbl[row_index, ]
  fit_brms <- fit_row$fit[[1]]
  #data <- fit_brms[["data"]]
  fit_stan <- fit_brms$fit
  chains_n <- length(dimnames(fit_stan)$chains)
  plots_row <- fit_row[["plots"]][[1]]
  
  desc <- fit_row$desc

  desc_filename <- gsub("_", "-", desc) %>%
    gsub("; ", "_", .) %>%
    gsub(":? ", "-", .)
  desc_short <- desc %>% str_replace("nbinom_type1; two_piece; ","")
  print(desc_short)

   ## indicate pXlot name
  plot_name <- "plot_trace"            
  plot_curr <- plots_row[[plot_name]]  
  if( (is.null(plot_curr)  | render_plots_force) & render_plots) { ## Generate and append plot
    ## Plotting code goes below here
    plot_trace <- mcmc_trace(fit, pars = c("lp__")) +
      ggtitle(desc_short)
    ## Finish plotting code
    plots_row[[plot_name]] <- plot_trace
  } else{
plot_trace <- plot_curr
    ## update last_plot() settings
    set_last_plot(plot_trace)
  }
  file_prefix <- sub("plot_", "", plot_name)                
  if(save_plots) last_plot_save(file_prefix = file_prefix)
  
  plot_name <- "plot_violin"            
  plot_curr <- plots_row[[plot_name]]            
  if( (is.null(plot_curr) | render_plots_force) & render_plots) { ## Generate and append plot
    ## Plotting code goes below here              
    ## Plot violin plots of posterior values
    ## Based on: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html#n_chains <-
    fit_array <- as.array(fit_stan)
    if(FALSE) { ## use violin plots
      plot_tmp <- mcmc_violin(fit_array, pars = c("lp__"), probs = c(0.1, 0.5, 0.9))
    } else { # use histograms
      plot_tmp <- mcmc_hist_by_chain(fit_array, pars = c("lp__"))
    }
    
    plot_violin <- plot_tmp +
      yaxis_ticks(on = TRUE) +
      yaxis_text(on = TRUE) +
      ggtitle(desc_short)
    ## Finish plotting code
    plots_row[[plot_name]] <- plot_violin
  } else{
    ## update last_plot() settings
    plot_violin <- plot_curr
    set_last_plot(plot_violin)
  }

  file_prefix <- sub("plot_", "", plot_name)
  if(save_plots) last_plot_save(file_prefix = file_prefix)
}
```

## Plot ??

```{r}


#fit_tbl <- fit_tbl_original

n_chains_max <- 4

for(row_index in 1:nrow(fit_tbl)) {

  cat(paste("Row: ", row_index))
  
  fit_row <- fit_tbl[row_index, ]
  fit_brms <- fit_row$fit[[1]]
  fit_stan <- fit_brms$fit
  chains_n <- length(dimnames(fit_stan)$chains)

  desc <- fit_row$desc

  desc_filename <- gsub("_", "-", desc) %>%
    gsub("; ", "_", .) %>%
    gsub(":? ", "-", .)
  desc_short <- desc %>% str_replace("nbinom_type1; two_piece; ","")
  print(desc_short)
  plot_trace <- mcmc_trace(fit_brms, pars = c("lp__"))
  last_plot()
  ## Plot violin plots of posterior values
  ## Based on: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html

  fit_array <- as.array(fit_stan)
  #n_chains <- 
  plot_violin <- mcmc_violin(fit_array, pars = c("lp__"), probs = c(0.1, 0.5, 0.9))
  last_plot()

  #fit_brms_best <- keep_best_chains(fit_brms, n_chains_max = n_chains_max)
  
  ## update tbl
  #fit_tbl[[row_index, "fit_best"]] <- fit_brms_best
}

## Save a version of tbl 
fit_tbl_filtered <- fit_tbl


```
