---
title: "Piecewise Regression with Negative Binomial Type I Error on Real Data using `brms` Custom Family"
author: "Michael Gilchrist"
date: "Created: 2023-02-28; Compiled: `r date()`"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)


if(interactive()) {
  default::default(.ess.eval) <- list(max.deparse.length=2E2, output = TRUE)
  output_dir = "output"  
} else { 
  output_dir = "output/render"  
}

```
# Goal

- Fit two piece negative binomial type 1 formulation to data
- From 2023-02-28 version of `nbinom_type1.R`

> Negative Binomial distribution parameterized by mean (mu) and overdispersion parameter (theta).
> This parameterization is referred to as NEGBIN type I (Cameron and Trivedi, 1998) as cited by
> https://doi.org/10.1080/03610926.2018.1563164
> ##
> `x ~ nbinom_type1(mu, theta)`, where `E(x) = mu, Var(x) = (theta + 1) mu`
> This should not be confused with the mu and shape parameterization of nbinom in R or the 'alternative' NB (`neg_binomial_2_...`) in stan
> Note using disp instead of theta because using theta gives the error
> > Error: Currently 'dirichlet' is the only valid prior for simplex parameters. See help(set_prior) for more details.
> when trying to fit the model.


## Recap

- Code initially based on work in 
- `../2022-12-20_rstan_two.piece.qpoisson.with.real.data/rstan_two.piece_fit.various.models.to.real.data.Rmd`
- 
- That model behaved well but, because I was writing stan code directly, I hadn't been able to group data by `male`
- Since then, I've created a `brms` `custom_family()` which uses the Type 1 formulation of the NB. 

## Insights

- When the `disp` (dispersal or `theta`) gets unrealistically large, we get the emergence of a bimodal distribution at both ends of `x0` values.  
Even though we included this value, it is very unlikely to be 25C values.
I interpret this to mean that when things are really noisy (high `theta`), one way to interpret the data is that one set of males has a very long (presumably slow) decline.
It would be good to look at the correlations via `pairs()`.
- To me this is consistent with the infomal knowledge that the 

### Reconciling use of `s0` with `x0`

In an effort to udnerstand hte difference between earlier work with `rstan` and this current work with `brms`, it seems that the key difference is both approaches use an implicit flat prior and, thus, result in inconsistency.

To understand this, I want to transform the flat prior on `s0` into the corresponding prior on `x0`.
In order to do so, I am following`Proposition B` on p. 60 of Rice (1995) stats book.

Specifically, if `s0 = 1/(xmax - x0)` such that `y = y0 * s0 * (xmax - max(x, x0))`

Then,
$f_{X0}(x0) = f_{s0}( g^{-1}(x0)) |\frac{d}{d x0} g^{-1}(x0)$ where

- $g^{-1}(x0) = 1/(x_{\max} - x0) = s0$
- $|\frac{d}{d x0} g^{-1}(x0)| = | \frac{d}{d x0}(1/(x_{\max} - x0)$
  $ = |\frac{1}{(x_{\max} - x0)^2}| = \frac{1}{(x_{\max} - x0)^2}$
- Assuming a flat prior means $f_{S0}(s0) = C$
- Thus, given a flat prior on $s0$, the corresponding prior on $x0$ is $f(x0) = C \frac{1}{(x_{\max} - x0)^2}$ which is not what I expected since it favors higher values of $x0$
- Structure of this prior is similar to a Cauchy distribution with $\gamma = 1$.

# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(brms)
library(loo)


## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Source family

```{r}

source("../../../custom-brms-families/families/nbinom_type1.R")
```
## Load Data

```{r}
sapply(file.path("input", dir("input")),
       load, verbose = TRUE, envir = .GlobalEnv)


head(stats_ind)
names(stats_ind)

head(data_ind)
names(data_ind)


```


# Prior Work

### Fit with Stan: Separate: `y0`, and `x0`, Pooled  $\theta$

- Based on results in `../2022-12-20_rstan_two.piece.qpoisson.with.real.data/rstan_two.piece_fit.various.models.to.real.data.Rmd`
- Histograms of `x0[]` suggest following groupings
1. `c(2, 4, 6, 8, 11)`
2. `c(1, 3, 5, 7, 10)`
3. `9` is ambiguous.


- NOTE: I'm not 100% sure these index values are the same as the ones we are using here!
```{r}

x0_groups_2_base <- list(low = c(2, 4, 6, 8, 11), high = c(1, 3, 5, 7, 10)) ## Does not include bird with index 9

x0_groups_2a <- x0_groups_2b <- x0_groups_2_base

x0_groups_2a$low <- c(x0_groups_2_base$low, 9) %>% sort()
x0_groups_2b$high <- c(x0_groups_2_base$high, 9) %>% sort()


```
# Fit Models

- Code derived from `../2023-02-23_brms_nb_focus.on.x0.fittings/nb_focus.on.estimating.x0.fittings.Rmd`



## Set up functions, parameters, and results tibble 

```{r}

data_stan <- data_ind %>% rename(y = motif_count, x = temp) %>%
  mutate(male = factor(male))

xmax <- 46
stan_two_piece_x0_func <- paste0(" real  two_piece(real x, real x0, real y0) {
 real xmax = ", xmax, "; ## paste in value for xmax\n
 return(y0 * (xmax - fmax(x0, x))/(xmax - x0));
 } ")

stan_two_piece_s0_func <- paste0(" real  two_piece(real x, real s0, real y0) {
 real xmax = ", xmax, "; ## paste in value for xmax\n
 real x0 = xmax - 1/s0;
 real y = y0 * s0 * (xmax - fmax(x, x0));
 return(y);
 } ")


models <- c("x0_piecewise") #, "s0_piecewise") #, "asymptotic")
sampling_dists <- c("nbinom_type1") ##, "com_poisson") ## lognormal doesn't work since the counts can be 0.
flags_x0 <- c("uniform_1",
              "groups_1", ## this doesn't work with x0_Intercept prior, suggests error in priors
              #"groups_2a",
              #"groups_2b",
              "individual")

flags_y0 <- switch(1,
                   c("individual"),
                   c("uniform_1", "groups_1", "individual")
                   )

disp_prior_list <- switch(2,
                          c(0.125), # This works well and is sane
                          c(0.0675, 0.125, 0.25), # used in exploring model behavior
                          c("flat") # not used
                          )

N <- length(data)

fit_tbl <- crossing(model = models,
                    sampling_dist = sampling_dists,
                    x0_flag = flags_x0, y0_flag = flags_y0,
                    disp_prior = disp_prior_list,
                    desc = "NA_character",
                    y0_group_list = list(NA), #tbl_tmp, #list(NA),
                    x0_group_list = list(NA),
                    fit = list(NA),
                    llik = list(NA),
                    r_eff = list(NA),
                    loo = list(NA)
                    )


```

### Run fits

```{r, message = FALSE }

run_fits = TRUE
force_load = FALSE  #reload fit_tbl even if it already exits
save_fits = TRUE

infile_tbl <- file.path(output_dir, paste0("fit_tbl_", Sys.Date(), ".Rda"))
outfile_tbl <- file.path(output_dir, paste0("fit_tbl_", Sys.Date(), ".Rda"))


sampling = "nbinom_type1"

flags_x0_used <- c("uniform_1", "individual") # "groups_1"
flags_y0_used <- c("individual")
disp_prior_used <- disp_prior_list
models_used <- c("x0_piecewise") #, "x0_piecewise")

fit_index <- 0;

for(model in models_used) {  
    switch(model,
#           "piecewise" = {stan_func <- stan_two_piece_x0_func},
           "x0_piecewise" = {
             stan_func <- stan_two_piece_x0_func
             warmup <- 3000 #floor(3/4 * iter)
             iter <- warmup + 3000
             adapt_delta <- 0.8
             thin <- 4
           },
           "s0_piecewise" = {
             stan_func <- stan_two_piece_s0_func
             warmup <- 15000
             iter <- warmup + 10000
             thin <- 4
             adapt_delta <- 0.95             
           }
           )

    print(model)
    print(stan_func)
    
    for(x0_flag in flags_x0_used) {
      s0_flag <- x0_flag ## define synonym
      
      for(y0_flag in flags_y0_used) {

        print_get_prior <- TRUE ## reset value
        print_prior_summary <- TRUE
        
      for(disp_prior in disp_prior_used) {

          ## used when loading fits
          fit_index <- fit_index + 1;
          ## Set up variables for saving model and fit

          desc_short <- paste0("x0 " , x0_flag, "; y0 ", y0_flag, "; disp prior ", disp_prior) 
          desc <- paste0(sampling, "; ", model, "; ", desc_short)

          filename_desc <- gsub("_", "-", desc) %>%
              gsub("; ", "_", .) %>%
              gsub(" ", "-", .)
          
          curr_row <- which(fit_tbl$model == model &
                            fit_tbl$sampling_dist == sampling &
                            fit_tbl$x0_flag == x0_flag &
                            fit_tbl$y0_flag == y0_flag &
                            fit_tbl$disp_prior == disp_prior)
        fit_tbl[ curr_row, ]$desc <- desc

        print(desc)

        if(run_fits) {

            
            print("Fitting Models")
              switch(sampling,
                     "nbinom_type1"= {
                         family <- nbinom_type1(link = "identity", link_disp = "identity")
                         adapt_delta <- adapt_delta #0.95 ## will decreasing value increase ESS? Seems like it
                         iter <- iter
                         warmup <- warmup
                         thin <- thin
                         n_cores <- 4 ## set to 1 if getting errors from stan in order to see relevant message.
                         n_chains <- n_cores
                         nbinom_type1_vars <-
                             stanvar(scode = paste(
                                         stan_func,
                                         stan_nbinom_type1, sep = "\n"),
                                     block = "functions")
                     }
         )

          ## Refresh data in case x0_group or y0_group are all set to 1
          data <- data_stan

          print("Set flags based on fitted model structure")
          if(x0_flag %in% c("uniform_1", "groups_1")) data <- mutate(data, x0_group = 1)
          if(y0_flag %in% c("uniform_1", "groups_1")) data <-mutate(data, y0_group = 1)
          if(x0_flag %in% c("individual")) data <- mutate(data, x0_group = male)
          if(y0_flag %in% c("individual")) data <- mutate(data, y0_group = male)



          ## Note we need to put a tibble into a list because row updates, even if doing
          ## just one cell, require a list format.
          ## Haven't defined "y0_group" or "x0_group" variables
          if(FALSE) {
            fit_tbl[[curr_row, "x0_group_list"]] <- list(unique(data[, c("male", "x0_group")]))
            fit_tbl[[curr_row, "y0_group_list"]] <- list(unique(data[, c("male", "y0_group")]))
          }
          
          print("Define parameter formulas")
          x0_form <- switch(x0_flag,
                            uniform_1 = formula(x0 ~ 1),
                            uniform_2 = formula(x0 ~ x0_group),
                            # Don't include x0_group info which is determined by the data set
                            groups_1 = formula(x0 ~ 0 + Intercept + (1|male)), 
                            groups_2 = formula(x0 ~ (1|male) + x0_group),
                            individual = formula(x0 ~ 0 + male) ## Do not use 1 + male!
                            )
          s0_form <- formula(deparse(x0_form) %>% gsub("x0 ", "s0 ", .))

          
          y0_form <- switch(y0_flag,
                            uniform_1 = formula(y0 ~ 1),
                            uniform_2 = formula(y0 ~ 1 + y0_group),
                            groups_1 = formula(y0 ~ 1 + (1| male)), 
                            groups_2 = formula(y0 ~ 1 + (1| male) + y0_group),
                            individual = formula(y0 ~ 0 + male)
                            ## `-1 + ` gives me the error:
                            ## Warning in parallel::mclapply(1:chains, FUN = callFun, mc.preschedule = FALSE,  :
                            ##  4 function calls resulted in an error
                            ## Error in FUN(X[[i]], ...) : 
                            ##  trying to get slot "mode" from an object (class "try-error") that is not an S4 object
                            ## should I use `0 +` or `-1 +` ?
                            )


          threshold_form <- switch(model,
                                     x0_piecewise = x0_form,
                                     s0_piecewise = s0_form
                                     )



          nlform <- switch(model,
                           x0_piecewise = bf(y ~ two_piece(x, x0, y0), nl = TRUE),
                           s0_piecewise = bf(y ~ two_piece(x, s0, y0), nl = TRUE)
                           ) +
            threshold_form + 
            y0_form


          if(print_get_prior) {
            print(get_prior(nlform,
                            data = data,
                            family = family
                            ))
            print_get_prior <- FALSE # will get reset
          }

          print("Define priors")
          ## Strangely reducing the parameter of the exponential results in a higher estimate of 'disp' and increasing the parameter decreases the estimates of 'disp'
          disp_priors <- set_prior(switch(1,
                               paste0("exponential(", disp_prior, ")"), ## Similar to one used in earlier work with rstan
                               normal(15, 2.8), ## Based on analysis of non-noisy 40C song data
                               normal(6, 2.8), ## Based on analysis of non-noisy 40C song data
                               exponential(0.667), ## 0.625 = 1/16  "
                               paste0("constant(", disp_prior, ")") ## Making disp a constant oesn't really help.
                              ), class = "disp", lb = 0, ub = 20)

          x0_prior <- switch(x0_flag,
                              uniform_1 = NULL,
                              uniform_2 = NULL,
                              # Don't include x0_group info which is determined by the data set
                              groups_1 = prior(student_t(3, 0, 66.7), lb = 0, ub = 3, class = "sd", nlpar = "x0"),
                              groups_2 = NULL,
                              individual = NULL
                              )
          
          x0_priors <- prior(uniform(25, 45), lb = 25, ub = 45, nlpar = "x0") + x0_prior

          s0_prior <- switch(s0_flag,
                              uniform_1 = NULL,
                              uniform_2 = NULL,
                              # Don't include x0_group info which is determined by the data set
                              groups_1 = prior(student_t(3, 0, 66.7), lb = 0, ub = 0.3333, class = "sd", nlpar = "s0"),
                              groups_2 = NULL,
                              individual = NULL
                              )

          s0_priors <- s0_prior +
            prior(uniform(0.02, 1), lb = 0.02, ub = 1, nlpar = "s0")
          
          y0_priors <- switch(y0_flag,
                              uniform_1 = prior(normal(150, 100), nlpar = "y0"),
                              uniform_2 = NULL,
                              groups_1 = NULL,
                              groups_2 = NULL,
                              individual = switch(1,
                                                  prior(normal(150, 1000), nlpar = "y0", lb = 10, ub = 1000), ## save value as with origina rstan fits
                                                  prior(uniform(10, 1000), nlpar = "y0", lb = 10, ub = 1000)
                                                  )
                              )

          threshold_priors <- switch(model,
                                     x0_piecewise = x0_priors,
                                     s0_piecewise = s0_priors
                                     )
                 
          prior <- switch(sampling,
                          "nbinom_type1" = {
                            disp_priors +
                              threshold_priors + y0_priors
                          })
          
            
          save_model <-paste0(paste0(output_dir, "/stan/brms", model, sampling, x0_flag, y0_flag, disp_prior, sep = "_"), ".stan")
          #make_stancode( .... save_model = save.model)
          fit <- brm(nlform,
                     data = data,
                     ## `link` refers to the mapping of the expectation of the distribution: log, sqrt, identity, softplus
                     ## link_shape corresponds to `phi` of `stan`'s
                     ## Negbinomial_2
                     ## Defining `phi = mu/theta` creates a quasipoisson
                     ## distribution with overdispersion parameter (1 +theta)
                     family = family, #negbinomial(link = "identity", link_shape = "identity"),
                     prior = prior,
                     stanvars = nbinom_type1_vars,
                     iter = iter,
                     warmup = warmup,
                     thin = thin,
                     silent = ifelse(interactive(), 1, 2), # 0, 1, or 2. 1 is default
                     control = list(adapt_delta = adapt_delta,
                                    max_treedepth = 12
                                    ##model_name = desc ## Incorrect way to set this.
                                    ),
                     ## Only print out sampling progress if in interactive mode
                     refresh = ifelse(interactive(),max(iter/5, 1), 0),
                     chains = n_chains,
                     cores = n_cores,
                     save_model = save_model
                     )

          # print(desc)
          # print("Prior Information")
          # print(prior_summary(fit))
          # print("Fit Information")
          # print(fit)
          
          #fit_exp <- expose_functions(fit) , vectorize = TRUE)
          #fit_cr <- add_criterion(fit_exp, c("loo", "waic"))
          fit_tbl[[curr_row, "fit"]] <- list(fit)
          ## Print current warnings
          warnings(summary)
          ## Clear warnings()

          ## End if(run_fit)
        } else {
          print("Working with Pre-existing Fits")

                    
          ## Try to assign from local memory.
          fit <- fit_tbl[[curr_row, "fit"]][[1]]

          ## Load fit if undefined or desired
          if(is.na(list(fit)) | force_load) {
            print("Loading Models")            
            #if(fit_index ==1)
            load(file = infile_tbl)
            fit <- fit_tbl[[curr_row, "fit"]][[1]]
          }
        }

        if(is.na(list(fit))) {
          warning(paste0("model fit ", desc, "does not exist.\n Skipping."))
        } else {
          ## Print and plot results, regardless of which fits one uses           
          print(desc)
          if(print_prior_summary) {
            print("Fit Prior Information")
            print(prior_summary(fit)) # %>% filter(nlpar!="y0"))
            print_prior_summary <- TRUE
          }
          print("Fit Information")
          print(summary(fit)) #, pars = "x0*"))          %>% filter(nlpar!="y0"))
          fit_stan <- fit$fit
          #clean up variable names
          fit_stan_rename <-
            fit_stan %>%
            setNames(gsub("b_", "", names(.)) %>%
                       gsub("(x0|s0|y0)_male(T[0-9]{3})", "\\2_\\1", .) %>%
                       gsub("__", "_", .) %>%
                       gsub("r_male_(x0|s0|y0)\\[(T[0-9]{3}),Intercept\\]", "\\2_\\1_r", .) %>%
                       gsub("\\.", " ", .))

          ##
          vars_fit <- names(fit_stan_rename) %>% na.omit(.) %>% sort(., decreasing = TRUE)
          ncol <- 4 
          hist <-  stan_hist(fit_stan_rename,
                             pars = vars_fit,
                             bins = 30,
                             ncol = ncol) +
            ggtitle(desc_short)
          print(hist)
          filename <- paste0("histogram_", filename_desc, ".pdf")
          ggsave(filename = filename, path = file.path(output_dir, "figures"), dpi=300)
        }
         #end else for fitting
      }
    }
  }
}

print(outfile_tbl)

if(save_fits) save(fit_tbl, file = outfile_tbl)
```


## Exit rendering
```{r}

knitr::knit_exit()

```
