---
title: "Fifth Fitting of Thermal Models"
author: "Michael Gilchrist"
date: "date: 2022-10-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
  )


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=1E5, output = TRUE)

```
# Goal

- Simple, but reasonable fits using GLM qpoisson with either
  - `motif_count` and `1|male`
  - `motif_count` and `offset(total_count)`
  - `motif_prop` with weights
- Use `temp_ref = 45` per Liz's request

## Notes from 27 Oct 2022 meeting

- TODO Me
  - Document simplest fits of 
     - quad qpoisson models to `motif_count`
     - general quad with weights to `motif_prop`
     - KEY QUESTION: Does `motif_prop` work as well as `motif_count`? 
       - If so, use all `motif_prop` data with `rTPC`
       - Fit lactin 2 and other models in fixed effects framework
       - Do model selection amongst
     - What's the diff between `motif_count` vs `motif_prop`
       - using `male` wiht `motif_count` is more precise accommodation for `male`, but costs us 1 df/male
       - not using `male` with `motif_prop` is a less precise accommodation with no cost in terms of df.
         It also allows us to use rTPC models (we think)
         
## Resources

### Poisson Regression and its extensions

- Provides some clear examples of R code
  - https://stats.oarc.ucla.edu/r/dae/negative-binomial-regression/
  
- Gives important details on variance terms
  - file://Rodriguez_2022_models.for.count.data.with.over.dispersion_Lecture.Notes.pdf
  


## Recap

- Using a Gamma model for the error converges, but I don't think it's the appropriate model, it assumes var ~ x^2
  See: 
## First and Second Fitting

- Temp is hard to control in chambers, so we should work with `temp_mean` (mean value during trial), not `temp_target`.
- `humidity_mean` and `temp_mean` strongly co-vary so consider using aggregate index as explanatory variable
  - Vapor Pressure Deficit (provided by `humidity` package)
  - Heat Index (formulated for humans, and provided by `weathermetrics`)
- Round 3 data only for curve fitting
  - Only round where `temp_mean` and `humidity_mean` exist.
- `count_total_round` are consistent between Rounds 1 and 3, so could use that info to classify birds, estimate variance function in response, etc.
- One bird in `round = 3` collapsed during the trial so it was terminated.
  We've set the `motif_count` from NA to 0 and should consider making at `temp_working` column using `temp_target` in this instance and `temp_mean` in all other instances.
  Would need to do something similar for relative humidity, i.e. use `mean(humidity_mean)` for the `temp_target` value.
  
## Third Fitting

- Created `temp` and `humidity` variables from `temp_mean` and `humidity_mean` for males who didn't collapse, used `mean(temp_mean)` and `mean(humidity_mean)` values for one male that did.
- Copied `data_full` and set  `male = "combined"` so we could look at all of the data at once.
- Learned about using `model` argument for glm models
- Can't fit ME models
  - I don't understand why the ME model with `motif_count` where we use a RE for the intercept doesn't have a similar effect as using motif_prop.
    This intercept value is essentially multiplying by a constant, so we could try and specify this value using `offset(log(count_total_round))`.
    So why doesn't this work?


## Fourth Fitting

- Tried fitting lots of variations of `glm` and `glmer`
- Optimizer matters in terms of convergence.
- Adding fixed `male` term rather than treating it as RE via `(1|male)` was very important since the variation in overall male singing varies so greatly.
- poisson fits indicate data is greatly overdispersed.
- Got one particular form of GLMM (`glmer`) working well (i.e. when there's no covariance matrix).
  I think this is a valid approach, but am unsure and a bit concerned since the `|` fitting shows the RE for $x1$ and $x1^2$ being close to perfectly correlated.
- Overall results indicate taht this approach is problematic
  - Data is extremely overdispersed in these fits.
  - As a result, no signal beyond a `male` effect on overall activity is reliably detected.
  
## Future Steps


# Set up

## Install libraries

```{r}

## install packages user might not have by replacing FALSE with TRUE
if(FALSE) {
    BiocManager::install("mixOmics") ## needed by RVAideMemoire
    install.packages(c("RSQLite", "nls.multstart", "lme4", "RVAideMemoire"))
    ##  Install the thermal curve package from git_hub, not cran
    remotes::install_github("padpadpadpad/rTPC")

}

## load libraries
library(stats)
require(MASS) # provides negative binomial fitting:  glm.nb
library(RSQLite)  # Don't think we need this.
library(rTPC)  ## 
library(nls.multstart)
library(broom)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(gridExtra)
library(viridisLite)

#options(ggplot2.continuous.colour="viridis",
#        ggplot2.discrete.colour="viridis",
#        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
#        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(GGally)
library(reshape2)
library(lme4)
library(rsample) ## provides bootstraps()

library(RVAideMemoire) # provides overdisp.glmer()
library(humidity) ## provides VPD
library(weathermetrics)
library(latex2exp)

```

## Load Data

```{r}

## Read in ZEFI Data sets
## Treat 'repeatability' as round = 0
## Add round info

## Repeatability was done between round 1 and 2, female was present, but only one temp. so treating as `round = 2` and redefining `round = 2` as `round = 3`

git_root <- system("git rev-parse --show-toplevel", intern = TRUE)

data_raw = list()

data_raw[[1]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Round-1-Heat-Trials.csv")) %>% mutate(round = 1) %>%
    ## Note T237 and T230 are missing numbers in the motif_count column
    ## so we are filtering these observations out until they are found
    filter(!is.na(motif_count))

data_raw[[2]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Repeatability-Song-Count.csv")) %>%
    mutate(round = 2) %>%
    group_by(male) %>%
    mutate(test_order = rank(date)) %>%
    ungroup()

data_raw[[3]] <-read.csv(file.path(git_root, "data", "collated", "HSPi-Round-2-Heat-Trials.csv")) %>%
    mutate(round = 3) %>%
    ## Deal with missing temp_mean and humidity_mean values
    ## in round == 3
    ## 2022/10/19 - code no longer needed
    ## group_by(temp_target) %>% 
    ##mutate(temp = if_else((round == 3 & is.na(temp_mean)),
    ##                      mean(temp_mean, na.rm = TRUE),
    ##                      temp_mean)) %>%
    ##mutate(humidity = if_else((round == 3 & is.na(humidity_mean)),
    ##                          mean(humidity_mean, na.rm = TRUE),
    ##                          humidity_mean)) %>%
    ungroup() 


## Join data and discard empty columns
data_full <- full_join(data_raw[[1]], data_raw[[2]]) %>%
    full_join(data_raw[[3]]) %>%
    discard(~all(is.na(.) | . =="")) %>% ## get rid of columns of only NA
    mutate(trial_completed = !(is.na(motif_count)) ) %>%
    mutate(motif_count = ifelse(is.na(motif_count), 0, motif_count)) %>%
    mutate(motif_count = motif_count*1.0) %>% ## convert to a double so it's not treated as an integer
    mutate(chamber = as.factor(chamber), male = as.factor(male)) %>%
    ## create a global variable trial_order based on individual rounds
    mutate(trial_index = as.integer(round*10+test_order)) %>%
    mutate(motif_count_plus_1 = (motif_count + 1)) %>%
    mutate(log_motif_count_plus_1 = log(motif_count + 1)) %>%
    mutate(temp_target = as.numeric(temp_target)) %>%
    ## Create generic 'temp' column which is either
    ## temp_mean, if it exists, or temp_target, if it doesn't
    mutate(temp = if_else(is.na(temp_mean),
                          temp_target,
                          temp_mean)) %>%
    ## Add column with total motif_count for a given round
    group_by(male, round) %>%
    mutate(count_total_round = sum(motif_count)) %>%
    ungroup() %>%
    mutate(motif_prop = motif_count/count_total_round) %>%
    ## assuming poisson error
    ## From glm man page
    ## > Non-‘NULL’ ‘weights’ can be used to indicate that different
    ## >  observations have different dispersions (with the values in
    ## >  ‘weights’ being inversely proportional to the dispersions);
    ## add +1 to deal with single 0
    ## Interpret dispersion as ~sd() or se() not var()
    mutate(count_wt = sqrt(1/(motif_count + 1))) %>%
    ## need to rescale wts for motif_prop data
    mutate(prop_wt = count_wt * count_total_round) %>% 
    ## Add vpd 
    mutate(svp = SVP(t = temp_mean + 273.15, isK = TRUE), vpd = svp*(1-humidity_mean/100) ) %>%
    group_by(round) %>%
    mutate(vpd_offset = vpd - mean(vpd)) %>%
    ungroup() %>%
    relocate(motif_count, motif_prop, vpd, temp_mean, humidity_mean, .after = male) %>% 
    mutate() ## Dummy function so we can comment out lines above it w/o any issues


```


# Examine Data  

## Create Working Dataset

```{r}

data_ind <- data_full %>%
    filter(round %in% c(2,3)) %>%
    filter(count_total_round >= 1) %>%
    mutate()

## copy data frame and assign `male =  "combined")
data_comb <- data_ind %>% mutate(male = "combined")

data <- bind_rows(data_ind, data_comb)

```

## Examine How Var(Mean) using `round = 2`

```{r}

motif_count_round_2  <- data_ind %>%
    filter(round == 2) %>%
    select(c(male, motif_count)) %>%
    group_by(male) %>% 
    mutate(mean = mean(motif_count), sd = sd(motif_count), var = var(motif_count)) %>%
    ungroup() %>%
    unique()

## Bootstrap data to get SE of var and mean
bs_data <- motif_count_round_2 %>%
select(male, motif_count) %>%
    bootstraps(times=100, strata = male)

bs_moments <- map_dfr(
  bs_data$splits,
  function(x) {
    dat <- as.data.frame(x) %>% group_by(male) %>%
    summarize(mean = mean(motif_count), var = var(motif_count), sd = sd(motif_count))
  }, .id = "sample")

bs_ci <- bs_moments %>%
    group_by(male) %>%
    summarize(mean_low = quantile(mean, 0.025),
              mean_high = quantile(mean, 0.975),
              sd_low = quantile(sd, 0.025),
              sd_high = quantile(sd, 0.975),
              var_low = quantile(var, 0.025),
              var_high = quantile(var, 0.975)
              )

##Clearly we have very little confidence in our moment estimators

motif_count_stats <- full_join(motif_count_round_2 %>% select(-motif_count) %>% unique(), bs_ci, by = "male")


g1 <- ggplot(motif_count_round_2) +
    aes(x=motif_count,
        color = male,
        fill = male) + 
    geom_histogram() +
    labs(title = "")

g2 <- ggplot(motif_count_stats) +
    aes(x=mean, y = sd) + 
    geom_point(aes(color = male)) +
    geom_smooth(method='lm', formula= y~x) +
    geom_errorbar(aes(ymin = sd_low, ymax = sd_high)) +
    geom_errorbarh(aes(xmin = mean_low, xmax = mean_high)) +
    labs(title = "Mean vs. sd for Round 2: temp_target = 40")

g3 <- ggplot(motif_count_stats) +
    aes(x=mean, y = var) +
    geom_point(aes(color = male)) +
    geom_smooth(method='lm', formula= y~x) +
##    geom_errorbar(aes(ymin = var_low, ymax = var_high)) +
##    geom_errorbarh(aes(xmin = mean_low, xmax = mean_high)) +
    labs(title = "Mean vs. Var for Round 2: temp_target = 40",
         subtitle = "log(var) ~ log(mean)") + 
    scale_x_log10() +
    scale_y_log10()
    
#ggcoefstats(
    
g4 <- ggplot(motif_count_stats) +
    aes(x=mean, y = var) +
    geom_point(aes(color = male)) +
    geom_smooth(method='glm', formula= y~-1 + x+I(x^2)) +
    geom_errorbar(aes(ymin = var_low, ymax = var_high)) +
    geom_errorbarh(aes(xmin = mean_low, xmax = mean_high)) +
    labs(title = "Mean vs. Var for Round 2: temp_target = 40",
         subtitle = "var ~ mean + mean^2")
#    scale_x_log10() +
#    scale_y_log10()

grid.arrange(g1, g2, g3, g4, ncol=2)


```

## Compare Rounds 2 and 3

```{r}

count_total_round_2_and_3  <- data_ind %>%
    select(c(male, count_total_round, round)) %>%
    unique() %>%
    pivot_wider(names_from = "round", values_from = "count_total_round")%>%
    select(male, `2`, `3`) %>%
    print(n=100)

```

## Rounds 2 and 3 vs Temp

```{r}


xlab <- "Temperature"
ylab <- "motif_count"

plot_temp_data <-
    ggplot(data) + 
    aes(x = temp,
        y = motif_count) +
    facet_wrap("male", scales = "free_y") +
    geom_point() +
labs( title = paste( ylab, " vs ", xlab))
last_plot()


pivot_wider(data_ind, names_from = "round", values_from = "count_total_round") %>% select(male, `2`, `3`) %>% filter(!is.na(`3`)) %>% unique()

```


## Formal Model Fits to `motif_count`


Using

- `x = temp -45C`
- use `offset(log(count_total_round))` instead of `male` as a factor

```{r}

temp_ref <- 45
verbose <- 0
trace <- FALSE

## Try filtering the data a bit more
## Goal is to get good starting values

data <- data_ind %>%
    mutate(x1 = (temp - temp_ref)) %>%
    filter(
    (round == 3 & count_total_round >= 400) |
    (round == 2 & count_total_round > 30)
    ) %>%
    mutate() %>%
#    filter( !(male %in% c("T231", "T260"))) %>% 
    mutate()

make_plot = TRUE;

glm_poisson_1 <- glm(motif_count ~
                       (1 + male  +  (x1) + I(x1^2)),
                   data = data,
                   family = poisson(link = "log")
                   )

summary(glm_poisson_1)

## Add round effect
glm_poisson_2 <- glm(motif_count ~
                       (1 + male + round +  (x1) + I(x1^2)),
                   data = data,
                   family = poisson(link = "log")
                   )
## Results support round effect
summary(glm_poisson_2)

male_coef <-  coef(glm_poisson_2) %>% keep(str_detect(names(.), 'male'))

hist(male_coef, breaks = 30)

## Remove male effect
glm_poisson_3 <- glm(motif_count ~
                       (1 + round +  (x1) + I(x1^2)),
                   data = data,
                   family = poisson(link = "log")
                   )
## Note that using qpoisson doesn't affect MLE

summary(glm_poisson_3)

if(make_plot){
    ifelse(length(dev.list()) < 3, {dev.new(); dev.next()}, dev.next())
    par(mfrow = c(2, 2))
    plot(glm_poisson_2, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
}

glm_qpoisson_1 <- update(glm_poisson_1,
                       family = quasipoisson(link = "log")
                       )
summary(glm_qpoisson_1)

if(make_plot){
    ifelse(length(dev.list()) < 3, dev.new(), dev.next())
    par(mfrow = c(2, 2))
    plot(glm_qpoisson_1, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
##    dev.new()
##    ggplot(data, aes(temp, motif_count))
}

glm_qpoisson_2 <- update(glm_poisson_2,
                       family = quasipoisson(link = "log")
                       )
summary(glm_qpoisson_2)

if(make_plot){
    ifelse(length(dev.list()) < 3, dev.new(), dev.next())
    par(mfrow = c(2, 2))
    plot(glm_qpoisson_2, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
##    dev.new()
##    ggplot(data, aes(temp, motif_count))
}


glm_qpoisson_3 <- update(glm_poisson_3,
                       family = quasipoisson(link = "log")
                       )
summary(glm_qpoisson_3)

if(make_plot){
    ifelse(length(dev.list()) < 3, dev.new(), dev.next())
    par(mfrow = c(2, 2))
    plot(glm_qpoisson_3, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
##    dev.new()
##    ggplot(data, aes(temp, motif_count))
}

```

### Result

- Overdispersion of data makes parameters non-significant
- Don't how a `motif_prop` approach will solve this issue.
  

