---
title: "Piecewise Regression with NB Error on 2022 and Earlier Data using BRMS"
author: "Michael Gilchrist"
date: "Created: 2023-02-07; Code run on `r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE, }

## Global options
options(digits = 3)

### Knitr specific
knitr::opts_chunk$set(
                      warning = TRUE, # show warnings
                      message = FALSE, # show messages
                      error = TRUE, # do not interrupt generation in case of errors,
                      echo = TRUE,  # show R code
                      fig.width = 8,
                      fig.path = "Figures/"
                  )

options("warn" = 1) ## print warnings when they occur

if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=5E1, output = TRUE)

```
# Goal

- Fit two piece nb to real data using `brms` where the goal is to extract as much info on `x0` as possible.
- Code largely derived from `../2023-01-12_brms_generate.and.fit.realistic.simulated.data/brms_two.piece_create.and.fit.simulated.data.Rmd`


## Recap

- Code runs.
- Have tried nb model as well as more exotic error models.
- Having set `x0max` to 45.5, we get 95% PI that are right up to this point.
Have changed ub to 45.9, but haven't tested it yet.
- Considered modeling `y0 ~lognormal()`, but that doesn't appear to be a good approximation.

# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = 4) #(parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(brms)
library(loo)
library(shinystan)


## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Load Data

```{r}
resave_data <- FALSE # flag for saving data_stan and possibly others in future

input_dir <- file.path("input")
output_dir <- file.path("output")

load(file.path(input_dir, "data.processing_2022-12-15.Rda"),
     verbose = TRUE)


motif_data
```

# Process Data

## Create Working Dataset

```{r}

filter_data <- TRUE

if(filter_data) {
    males_filtered_disp <- motif_stats_40C %>%
        filter(dispersion < 50) %>%
        pull(male)

    males_filtered_mean <- motif_stats %>%
        filter(mean > 10) %>% # changing from 10 to 40 removes previous male 7 (T258)
        pull(male)

    male_vector <- intersect(males_filtered_mean, males_filtered_disp)
} else {
    male_vector <- motif_data %>% select(male) %>% distinct()
}

data_ind <- motif_data %>%
  filter(male %in% male_vector) %>%
  mutate(male = droplevels(male)) %>%
  mutate(index = as.integer(male)) %>%
  mutate(male = as.character(male)) %>% 
  arrange(index) %>%
  select(male, index, motif_count, temp_target, temp, round, trial_round, date, counter) %>% 
  ##    left_join(index_shape, by = "index") %>%
  group_by(male) %>% mutate(y0_simple_est = mean(motif_count), phi_ind = var(motif_count)/y0_simple_est) %>% ## phi is overdispersion parameter
  ungroup() %>% 
  mutate()

stats_ind <- motif_stats %>%
    filter(male %in% male_vector) 


summary(data_ind)

y0_hist <- data_ind %>%
  select(y0_simple_est) %>%
  unique() %>%
  ggplot(aes(y0_simple_est)) + geom_histogram(bins = 8) 
print(y0_hist)

log_y0_hist <- y0_hist +
  scale_x_log10()
print(log_y0_hist)


temp_hist <- ggplot(data_ind, aes(temp)) + geom_histogram()
print(temp_hist)

n_male <- length(unique(data_ind$male))

## for calculation of 'theta_bar' (size in `rnbinom`) see note in `Set Up Simulate Data`
summary_stats <- data_ind %>%
    ungroup() %>%
    summarize(y0_bar = mean(y0_simple_est),
              y0_sd = sd(y0_simple_est),
              log_y0_bar = mean(log(y0_simple_est)),
              log_y0_sd = sd(log(y0_simple_est)),
              phi_bar = median(phi_ind), # qpoisson over dispersion
              theta_bar = y0_bar^2/(y0_sd^2 - y0_bar)) # nb shape parameter
comment(summary_stats) <- "summary stats for observed bird motifs"

save(summary_stats, file = file.path(output_dir, "obs_summary_stats.Rda"))

```
# Fit Models

## Common Parameters

```{r}

xmax <- 46
x0max <- xmax - 0.1;
x0min <- 20;

y_xmax <- 0
y0_min <- 10
sd_y0_prior <- 200
alpha_theta_prior <- 10 ## exponential dist scale parameter for overdispersion theta in quasipoisson
alpha_phi_prior <- 0.01 ## gamma dist shape parameter for nb. brms default is 0.01
beta_phi_prior <- 0.01 ## gamma dist rate parameter for nb. brms default is 0.01


## values to use for model predictions
x_for_predictions = seq(25, xmax, length.out = 100)
```

### Analyze `x ~ uniform_1 | groups_1, y ~ individual`

Set up Data

```{r}

data_stan <- data_ind %>%
    mutate(x0_group = NA, y0_group = NA, male = factor(male)) %>%
    rename(y = motif_count, x = temp)

if(resave_data) save(file = file.path(output_dir, paste0("data_stan_", format(Sys.time(), "%Y-%m-%d_%H-%M"), ".Rda")), data_stan)

```

```{r}

models <- c("piecewise")
sampling_dists <- c("nb") #, "com_poisson") ## lognormal doesn't work since the counts can be 0.
flags_x <- c("uniform_1",
             "groups_1" ## this doesn't work with x0_Intercept prior, suggests error in priors
             ##"individual"
             )
flags_y <- c("individual")
N <- length(data)

fit_tbl <- crossing(model = models,
                    sampling_dist = sampling_dists,
                    x0_flag = flags_x, y0_flag = flags_y,
                    desc = "NA_character",
                    y0_group_list = list(NA), #tbl_tmp, #list(NA),
                    x0_group_list = list(NA),
                    fit = list(NA),
                    llik = list(NA),
                    r_eff = list(NA),
                    loo = list(NA)
                    )
```


## Run Analysis

```{r}

model = models[[1]]

for(sampling in sampling_dists) {

    switch(sampling,
           "nb"= {
             family <- negbinomial(link = "identity", link_shape = "identity")
             adapt_delta <- 0.9 #0.95
             iter <- 4000
             warmup <- floor(1/2 * iter)
             thin <- 1
             n_cores <- 4 ## set to 1 if getting errors from stan in order to see relevant message.
             n_chains <- n_cores

           },
           "lognormal" = { ## doesn't work since we have 0 counts and 0 is undefined in lognormal
             family <- lognormal(link = "identity", link_sigma = "log")
             adapt_delta <- 0.9 #0.95
             iter <- 2000
             warmup <- floor(1/2 * iter)
             thin <- 1
             n_cores <- 4 ## set to 1 if getting errors from stan in order to see relevant message.
             n_chains <- n_cores

           },
           "com_poisson"={
             ## generalized poisson
             ## quasipoisson
             ## COM-Poisson: See https://discourse.mc-stan.org/t/brms-and-conway-maxwell-poisson-distribution/7368/17
             ##   Also: str(brmsfamily("com_poisson"))
             family <- brmsfamily("com_poisson", link="identity", link_shape="identity") #com_poisson(link="identity", link_shape="log")
             adapt_delta <- 0.99 #0.95
             iter <- 6000
             warmup <- floor(4/5 * iter)
             thin <- 1
             n_cores <- 4 ## set to 1 if getting errors from stan in order to see relevant message.
             n_chains <- n_cores
             }
           )
  
        print(family)
        

    
    for(x_flag in flags_x[[2]]) {
    for(y_flag in flags_y) {

        ## Set up variables for saving model and fit
        desc <- paste0(model, ": ", sampling, ", " , x_flag, ", ", y_flag)
        curr_row <- which(fit_tbl$sampling_dist == sampling &
                          fit_tbl$x0_flag == x_flag &
                          fit_tbl$y0_flag == y_flag)
        fit_tbl[ curr_row, ]$desc <- desc

      print(desc)
      ## Refresh data in case x0_group or y0_group are all set to 1
      data <- data_stan

      ## Set flags based on fitted model structure
      if(x_flag %in% c("uniform_1", "groups_1")) data <- mutate(data, x0_group = 1)
      if(y_flag %in% c("uniform_1", "groups_1")) data <-mutate(data, y0_group = 1)
      if(x_flag %in% c("individual")) data <- mutate(data, x0_group = male)
      if(y_flag %in% c("individual")) data <- mutate(data, y0_group = male)



      ## Note we need to put a tibble into a list because row updates, even if doing
      ## just one cell, require a list format.
      fit_tbl[[curr_row, "x0_group_list"]] <- list(unique(data[, c("male", "x0_group")]))
      fit_tbl[[curr_row, "y0_group_list"]] <- list(unique(data[, c("male", "y0_group")]))

        ## Parameter Structure
        x_form <- switch(x_flag,
                         uniform_1 = formula(x0 ~ 1),
                         uniform_2 = formula(x0 ~ x0_group),
                         # Don't include x0_group info which is determined by the data set
                         groups_1 = formula(x0 ~ (1|male)), ## adding 0 + causes prior error
                         groups_2 = formula(x0 ~ 0 + (1||male) + x0_group),
                         individual = formula(x0 ~ 0 + male)
                         )
        
        y_form <- switch(y_flag,
                         uniform_1 = formula(y0 ~ 1),
                         uniform_2 = formula(y0 ~ 0 + y0_group),
                         groups_1 = formula(y0 ~ 0 + (1|| male)), 
                         groups_2 = formula(y0 ~ 0 + (1|| male) + y0_group),
                         individual = formula(y0 ~ 0 + male)
                         ## `-1 + ` gives me the error:
                         ## Warning in parallel::mclapply(1:chains, FUN = callFun, mc.preschedule = FALSE,  :
                         ##  4 function calls resulted in an error
                         ## Error in FUN(X[[i]], ...) : 
                         ##  trying to get slot "mode" from an object (class "try-error") that is not an S4 object
                         ## should I use `0 +` or `-1 +` ?
                         )

      ## Priors 
      prior_core <- prior(
        uniform(10, 1000), nlpar = "y0", lb = 10, ub = 1000
        ##normal(150, 200), nlpar = "y0", lb = 10, ub = 1000 ## Previously used          
        ##lognormal(log_y0_mean, 4 * log_y0_sd ## Never used due to data not really looking ~ logN
        ##gamma(0.1, 100), nlpar = "y0", lb = 10, ub = 1000
      ) + 
        prior(uniform(30, 45.5), lb = 30, ub = 45.5, nlpar = "x0")
      
      prior <- switch(sampling,
                      "nb" = prior_core, #  + prior(exponential(1), dpar = "shape", class = "sd"),
                      "lognormal" = prior_core,
                      "com_poisson" = prior_core  + prior(uniform(0.01, 100), class = "shape", lb = 1, ub = 20) ## adding 1 <= lb and ub greatly speeds up model and improves fit
                      )
      
      # Crazy parameter rich model where shape parameters vary by male
      hyper_form <- switch(sampling,
                      "nb" = formula(shape ~ 1|male),
                      "lognormal" = formula(sigma ~ 1|male),
                      "com_poisson" = formula(shape ~ 1|male) )
      hyper_form <- NA
      
      nlform <- bf(
            y ~  0 - (46 - fmax(x0, x)) * (0 - y0)/(46 - x0),
            x_form,
            y_form,
            hyper_form,
            nl = TRUE)

        if(TRUE) print(get_prior(nlform,
                                 data = data,
                                 family = family
                                 ))

        save_model <-paste0(paste("brms", model, sampling, x_flag, y_flag, sep="_"), ".stan")
        
        fit <- brm(nlform,
                   data = data,
                   ## `link` refers to the mapping of the expectation of the distribution: log, sqrt, identity, softplus
                   ## link_shape corresponds to `phi` of `stan`'s
                   ## Negbinomial_2
                   ## Defining `phi = mu/theta` creates a quasipoisson
                   ## distribution with overdispersion parameter (1 +theta)
                   family = family, #negbinomial(link = "identity", link_shape = "identity"),
                   prior = prior,
                   iter = iter,
                   warmup = warmup,
                   thin = thin,
                   silent = ifelse(interactive(), 1, 2), # 0, 1, or 2. 1 is default
                   control = list(adapt_delta = adapt_delta,
                                  max_treedepth = 12
                                  ##model_name = desc ## Incorrect way to set this.
                                  ),
                       ## Only print out sampling progress if in interactive mode
                   refresh = ifelse(interactive(),max(iter/5, 1), 0),
        chains = n_chains,
        cores = n_cores,
        save_model = save_model
        )
    ## Avoid having brms recompile model by defining
    ## model in global environment
        fit_tmp <- fit

                                        #fit_exp <- expose_functions(fit) , vectorize = TRUE)
                                        #fit_cr <- add_criterion(fit_exp, c("loo", "waic"))
        print(fit)
        save(file = file.path(output_dir, paste0(save_model, ".Rdata-tmp")), fit)
        fit_tbl[[curr_row, "fit"]] <- list(fit)

      ## Print current warnings
      warnings(summary)
      ## Clear warnings()
      ##warning(immediate. = FALSE)

    }
    }
}

save(file = file.path(output_dir, paste0("fit_tbl_", format(Sys.time(), "%Y-%m-%d_%H-%M"), ".Rdata")), fit_tbl)

```

##### Output from model fits

