---
title: "Evaluate How `motif_count` changes with trial at 40C"
author: "Michael Gilchrist"
date: "date: 2022-10-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE#,  # show R code
  ##results="asis" # Hopefully print summary() in for loops
  )


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=1E5, output = TRUE)

```
# Goal

- Evaluate trends in motif_count under a (near) constant temperature.

# Set up

## Load libraries

```{r}


## load libraries
library(stats)
require(MASS) # provides negative binomial fitting:  glm.nb
library(RSQLite)  # Don't think we need this.
library(rTPC)  ## 
library(nls.multstart)
library(broom)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(grid) ## provides textGrob
library(gridExtra)
library(viridisLite)

#options(ggplot2.continuous.colour="viridis",
#        ggplot2.discrete.colour="viridis",
#        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
#        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(GGally)
library(reshape2)
library(lme4)
library(nlme)
library(gnm)
library(rsample) ## provides bootstraps()

library(RVAideMemoire) # provides overdisp.glmer()
library(humidity) ## provides VPD
library(weathermetrics)
library(latex2exp)

```
## Local Functions

```{r}

kprint <- function(input, ...) {
    print(knitr::kable(input, ...))
cat('\n\n<!-- -->\n\n')
}
```

## Plotting settings

```{r}

## From: https://data-se.netlify.app/2018/12/12/changing-the-default-color-scheme-in-ggplot2/

theme_set(theme_minimal(base_size = 9))
theme_update(
    plot.title = element_text(size = rel(1.1)),
    plot.subtitle = element_text(size = rel(1)))

if(!exists("old_opts")) old_opts <- options()  # save old options

options(ggplot2.continuous.colour="viridis")
options(ggplot2.continuous.fill = "viridis")
options(ggplot2.discrete.colour="viridis")
options(ggplot2.discrete.fill = "viridis")


```

## Load Data

```{r}

## Read in ZEFI Data sets
## Treat 'repeatability' as round = 0
## Add round info

## Repeatability was done between round 1 and 2, female was present, but only one temp. so treating as `round = 2` and redefining `round = 2` as `round = 3`

output_dir <- "output"

if(!dir.exists(output_dir)) dir.create(output_dir)
git_root <- system("git rev-parse --show-toplevel", intern = TRUE)

data_raw = list()

data_raw[[1]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Round-1-Heat-Trials.csv")) %>% mutate(round = 1) %>%
    ## Note T237 and T230 are missing numbers in the motif_count column
    ## so we are filtering these observations out until they are found
    filter(!is.na(motif_count))

data_raw[[2]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Repeatability-Song-Count.csv")) %>%
    mutate(round = 2) %>%
    ungroup()

data_raw[[3]] <-read.csv(file.path(git_root, "data", "collated", "HSPi-Round-2-Heat-Trials.csv")) %>%
    mutate(round = 3) %>%
    ## Deal with missing temp_mean and humidity_mean values
    ## in round == 3
    ## 2022/10/19 - code no longer needed
    ## group_by(temp_target) %>% 
    ##mutate(temp = if_else((round == 3 & is.na(temp_mean)),
    ##                      mean(temp_mean, na.rm = TRUE),
    ##                      temp_mean)) %>%
    ##mutate(humidity = if_else((round == 3 & is.na(humidity_mean)),
    ##                          mean(humidity_mean, na.rm = TRUE),
    ##                          humidity_mean)) %>%
    ungroup() 


## Join data and discard empty columns
data_full <- full_join(data_raw[[1]], data_raw[[2]]) %>%
    full_join(data_raw[[3]]) %>%
    discard(~all(is.na(.) | . =="")) %>% ## get rid of columns of only NA
    mutate(trial_completed = !(is.na(motif_count)) ) %>%
    mutate(motif_count = ifelse(is.na(motif_count), 0, motif_count)) %>%
    mutate(motif_count = motif_count*1.0) %>% ## convert to a double so it's not treated as an integer
    mutate(chamber = as.factor(chamber), male = as.factor(male)) %>%
    ## create a male specific round and global trial index `trial`
    group_by(male, round) %>%
    mutate(trial_round = rank(date)) %>%
    ungroup(round) %>% 
    mutate(trial = rank(date)) %>%
    ungroup() %>%
    mutate(motif_count_plus_1 = (motif_count + 1)) %>%
    mutate(log_motif_count_plus_1 = log(motif_count + 1)) %>%
    mutate(temp_target = as.numeric(temp_target)) %>%
    ## Create generic 'temp' column which is either
    ## temp_mean, if it exists, or temp_target, if it doesn't
    mutate(temp = if_else(is.na(temp_mean),
                          temp_target,
                          temp_mean)) %>%
    ## Add column with total motif_count for a given round
    group_by(male, round) %>%
    mutate(count_total_round = sum(motif_count)) %>%
    ungroup() %>%
    mutate(motif_prop = motif_count/count_total_round) %>%
    ## assuming poisson error
    ## From glm man page
    ## > Non-‘NULL’ ‘weights’ can be used to indicate that different
    ## >  observations have different dispersions (with the values in
    ## >  ‘weights’ being inversely proportional to the dispersions);
    ## add +1 to deal with single 0
    ## Interpret dispersion as ~sd() or se() not var()
    mutate(count_wt = sqrt(1/(motif_count + 1))) %>%
    ## need to rescale wts for motif_prop data
    mutate(prop_wt = count_wt * count_total_round) %>% 
    ## Add vpd 
    mutate(svp = SVP(t = temp_mean + 273.15, isK = TRUE), vpd = svp*(1-humidity_mean/100) ) %>%
    group_by(round) %>%
    mutate(vpd_offset = vpd - mean(vpd)) %>%
    ungroup() %>%
    relocate(motif_count, motif_prop, vpd, temp_mean, humidity_mean, .after = male) %>% 
    mutate() ## Dummy function so we can comment out lines above it w/o any issues


```


# Examine Data

## Create Working Dataset

```{r}

data_ind <- data_full %>%
#    filter(round %in% c(2,3)) %>%
#    filter(count_total_round >= 1) %>%
    mutate()

## copy data frame and assign `male =  "combined")
data_comb <- data_ind %>% mutate(male = "combined")

data <- bind_rows(data_ind, data_comb)

```

## Examine How Var varies with Mean `temp_target = 40`

**WARNING:** `warning()` are off

```{r, warning = FALSE}

data_40  <- data_ind %>%
    filter(temp_target == 40) %>%
    unique()

dim(data_40)

stats_40 <- data_40 %>%
    group_by(male) %>%
    summarize(count = length(motif_count),
              mean = mean(motif_count),
              var = var(motif_count, na.rm = TRUE),
              cv = sqrt(var)/mean,
              dispersion = var/mean) %>% mutate()

print("We have 15 males, 5 of which we only have 1 observation at 40C")
stats_40

motif_stats <- stats_40

print("Flag birds birds with a coefficient of variation `cv` > 0.5 & `mean` < 50") 
cv_threshold <- 0.5
mean_threshold <- 50

male_high_cv <- stats_40 %>% filter(cv > cv_threshold) %>% pull(male)
data_male_high_cv <- data_40 %>% filter(male %in% male_high_cv)

male_low_mean <- stats_40 %>% filter(mean < mean_threshold) %>% pull(male)
data_male_low_mean <- data_40 %>% filter(male %in% male_low_mean)


g0 <- ggplot(data_male_high_cv) +
    aes(x=motif_count,
        color = male,
        fill = male) + 
    geom_histogram() +
    labs(title = paste0("Data for males with exceptionally high cviances", paste(male_high_cv, collapse = " & "))
         )


for(filter_high_cv in c(FALSE, TRUE)){
    for(filter_low_mean in c(FALSE, TRUE)){

        data_tmp <- data_40
        motif_stats_tmp <- motif_stats
        
        if(filter_high_cv){
            data_tmp <- filter(data_tmp, !(male %in% male_high_cv) )
            motif_stats_tmp <- filter(motif_stats_tmp, !(male %in% male_high_cv) )
        }

        if(filter_low_mean){
            data_tmp <- filter(data_tmp, !(male %in% male_low_mean) )
            motif_stats_tmp <- filter(motif_stats_tmp, !(male %in% male_low_mean) )
        }

        g0 <- ggplot(data_tmp) +
            aes(x=male,
                y = motif_count,
                color = male,
                fill = male) + 
            geom_point() +
            geom_rug(size=0.75) +
            ## theme(axis.ticks.x=element_line(size=30))+
            geom_point(
                mapping = aes(x = male,
                              y = mean,
                              color = male),
                size = 2.5,
                shape = 24,
                data = motif_stats_tmp) +
            theme(legend.position = "none", axis.text.x = element_text(angle = -45)) +
            labs(title = "Raw Count Data")

        g1 <- ggplot(data_tmp) +
            aes(x=motif_count,
                color = male,
                fill = male) + 
            geom_histogram() +
            xlim(0, NA) +
            labs(title = "Raw Count Data")
        
        formula = y ~ -1 + x
        g2 <- ggplot(motif_stats_tmp) +
            aes(x=mean, y = var) +
            geom_point(aes(color = male)) +
            geom_smooth(method='lm', formula = formula, fullrange = TRUE) +
            ##stat_cor(label.y.npc = 0.95) +
            stat_regline_equation(label.y.npc = 0.9, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~")), formula = formula, size = 2.5) +
            theme(legend.position = "none") +
            xlim(0, NA) +
            labs(title = "Mean vs. var")

        fit_g2 <- lm(var ~ -1 + mean, data = motif_stats_tmp)
        
        g3 <- ggplot(motif_stats_tmp) +
            aes(x=mean, y = var) +
            geom_point(aes(color = male)) +
            geom_smooth(method='lm', formula= formula, fullrange = TRUE) +
            stat_regline_equation(label.y.npc = 0.9, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~")), formula = formula, size = 2.5) +
            theme(legend.position = "none") +
            labs(title = "Mean vs. Var",
                 subtitle = "log(var) ~ log(mean)") + 
            scale_x_log10() +
            scale_y_log10()


        formula <- y ~ -1 + x + I(x^2)
        g4 <- ggplot(motif_stats_tmp) +
            aes(x=mean, y = var) +
            geom_point(aes(color = male)) +
            geom_smooth(method='lm', formula = formula, fullrange = TRUE) +
            stat_regline_equation(label.y.npc = 0.9, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~")), formula = formula, size = 2.5) +
            theme(legend.position = "none") +
            xlim(0, NA) +
            labs(title = "Mean vs. Var",
                 subtitle = "var ~ mean + mean^2")
                                        #    scale_x_log10() +
                                        #    scale_y_log10()
        fit_g4 <- lm(var ~ -1 + mean +  I(mean^2), data = motif_stats_tmp)
        
        formula <- y ~ -1 + I(x^2)
        g5 <- ggplot(motif_stats_tmp) +
            aes(x=mean, y = var) +
            geom_point(aes(color = male)) +
            geom_smooth(method='lm', formula = formula, fullrange = TRUE) +
            stat_regline_equation(label.y.npc = 0.9,
                                  aes(label = paste(
                                          ..eq.label..,
                                          ..adj.rr.label.., sep = "~~")),
                                  formula = formula, size = 2.5) +
            theme(legend.position = "none") +
            labs(title = "Mean vs. Var",
                 subtitle = "var ~ mean^2")

        fit_g5 <- lm(var ~ -1 + I(mean^2), data = motif_stats_tmp)
        #ifelse(length(dev.list()) < 4, dev.new(), dev.next())
        
        ga <- grid.arrange(g0, g2, g4, g5,
                     ncol=2,
                     top=textGrob(
                         paste0("Mean vs. Var: Filter High CV = ", filter_high_cv,
                                ", Low Mean = ", filter_low_mean, "\n temp_target = 40C"),
                         gp=gpar(fontsize = 11))
                     )
        ##cat("\n\n\\pagebreak\n")
        ##print("<P style='page-break-before: always'>")    #forced new-page happens here.

        print(paste0("Mean vs Var; filter high cv: ", filter_high_cv, "; filter low mean: ", filter_low_mean))
        print(summary(fit_g2))
        print(summary(fit_g4))
        print(summary(fit_g5))

ga
        
        dev.print(device = pdf, file = file.path(output_dir, paste0("mean.vs.var_filter.high-", filter_high_cv, ".filter.low.mean-", filter_low_mean, ".pdf")))
        
    }
}

```


### Conclusion

- Excluding males with very high variances indicates that for the remaining males `var ~ mean`, but overdispersed relative to the poisson.
- Using multiple trials at one temp could be an effective way to identify 'reliable' males.


## Compare `motif_count` vs `trial` at 40C

**WARNING:** `warning()` are off

```{r, warning = FALSE}

for(filter_high_cv in c(FALSE, TRUE)){
    for(filter_low_mean in c(FALSE, TRUE)){
        

        data_tmp <- data_40 %>% group_by(male) %>%
            arrange(trial, .by_group = TRUE) %>%
            mutate(cummean = cummean(motif_count)) %>%
            select(male, trial, motif_count, cummean)
        
        if(filter_high_cv){
            data_tmp <- filter(data_tmp, !(male %in% male_high_cv) )
        }

        if(filter_low_mean){
            data_tmp <- filter(data_tmp, !(male %in% male_low_mean) )
        }
        
        


        g1 <- ggplot(data_tmp) +
            aes(x = trial, y = motif_count) +
            geom_point(aes(color = male), position = "jitter") +
            geom_line(aes(x = trial, y = cummean, color = male)) +
            theme(legend.position="bottom")
        
        legend <- get_legend(g1)


        ## Apply a linear regression

        formula <- y ~ 1 + x
        g0 <- ggplot(data_tmp, aes(x = trial, y = motif_count)) +
            geom_point(aes(color = male))

        
        g2 <- g0 +
            geom_smooth(method = 'glm', formula = formula, fullrange = TRUE) +
            stat_regline_equation(label.y.npc = 0.9, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~")), formula = formula, size = 2.5) +
            theme(legend.position = "none") +
            xlim(1, NA) +
            labs(title = "Linear")
        ##g2

        fit_g2 <- glm( motif_count ~ 1 + trial, data = data_tmp)

        g3 <- g0 + geom_smooth(method = 'glm',
                               formula= formula,
                               ## Can't use link = identity because it leads to negative expectations
                               method.args = list(family = quasipoisson(link = "log"))) +
            stat_regline_equation(label.y.npc = 0.9, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~")), formula = formula, size = 2.5) +
            theme(legend.position = "none") +
            xlim(1, NA) +
            labs(title = "log-linear: family = qpoisson(link = log)")
        ##g3
        
        fit_g3 <- glm( motif_count ~ 1 + trial, data = data_tmp,
                      family = quasipoisson(link = "log"))
        formula <- y ~ 1 + x + I(x^2)
        g4 <- g0 + geom_smooth(method='glm',
                               formula= formula,
                               ## Can't use link = identity because it leads to negative expectations
                               method.args = list(family = quasipoisson(link = "log"))) +
            stat_regline_equation(label.y.npc = 0.9, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~")), formula = formula, size = 2.5) +
            theme(legend.position = "none") +
            xlim(1, NA) +
            labs(title = "log-quad: family = qpoisson(link = log)")
        ##g4
        
        fit_g4 <- glm(motif_count ~ 1 + trial + I(trial^2), data = data_tmp,
                      family = quasipoisson(link = "log"))

        ga <- grid.arrange(g2, g3, g4, as_ggplot(legend),
                           ncol=2,
                           top=textGrob(
                               paste0("motif_count vs. trial: Filter High CV = ", filter_high_cv, ", Low Mean = ", filter_low_mean, "\n temp_target = 40C"),
                               gp=gpar(fontsize = 11))
                           )
        ##cat("\n\n\\pagebreak\n")
        ##print("<P style='page-break-before: always'>")    #forced new-page happens here.

        print(paste0("motif_count vs. trial; filter.high: ", filter_high_cv, "; filter.low.mean: ", filter_low_mean))

        print(summary(fit_g2))
        print(summary(fit_g3))
        print(summary(fit_g4))

        ga
        
        dev.print(device = pdf, file = file.path(output_dir, paste0("motif_count.vs.trial_filter.high-", filter_high_cv, ".filter.low.mean-", filter_low_mean, ".pdf")))
    }
}


```

### Conclusion

When filtering out low mean (which have few trials and, thus, bias the analysis towards a small intercept) results in no evidence that `motif_count` increases with `trial`

# End

```{r, error = FALSE}

knitr::knit_exit()

```


## Try asymptotic regression

```{r}

## See: https://www.statforbiology.com/nonlinearregression/usefulequations

    if(FALSE) {
        ## Try using `nlme`
        ## I get errors
        ## Also not specifying error model
        getInitial(motif_count ~ SSasymp(trial, Asym, resp0, lrc), data = data_tmp)  
        ar <- nls(motif_count ~ SSasymp(trial, Asym, resp0, lrc), data = data_tmp)

        ## Try using motif_count_plus_1
        getInitial(motif_count_plus_1 ~ SSasymp(trial, Asym, resp0, lrc), data = data_tmp)  
        ar <- nls(motif_count_plus_1 ~ SSasymp(trial, Asym, resp0, lrc), data = data_tmp)
    }


    
    ## Try using gnm
    ## a - (a - b) exp(-c x) = a - b' exp(-c x) = a - exp(b'' - c x)
    ## According to TurnerAndFirth2007 R News, this model is written Exp(1 + trial)
    ## However the model doesn't converge unless we use -1
    ## I clearly don't understand the model 
    formula <- motif_count ~ 1 + Mult(Const(-1), Exp(1 + trial))
    gnm_poisson <- gnm(formula = formula,
                       data = data_tmp,
                       ## start = c(50, -1, -0.1),
                       family = poisson(link = "identity"),
                       ##weights = 1/motif_count_plus_1,
                       trace = TRUE,
                       iterMax = 5000
                   )


    
    model <- drm(Y ~ X, fct = DRC.asymReg())


count_total_round_2_and_3  <- data_ind %>%
    select(c(male, count_total_round, round)) %>%
    unique() %>%
    pivot_wider(names_from = "round", values_from = "count_total_round")%>%
    select(male, `2`, `3`) %>%
    print(n=100)

```

## Rounds 2 and 3 vs Temp

```{r}


xlab <- "Temperature"
ylab <- "motif_count"

plot_temp_data <-
    ggplot(data) + 
    aes(x = temp,
        y = motif_count) +
    facet_wrap("male", scales = "free_y") +
    geom_point() +
labs( title = paste( ylab, " vs ", xlab))
last_plot()


pivot_wider(data_ind, names_from = "round", values_from = "count_total_round") %>% select(male, `2`, `3`) %>% filter(!is.na(`3`)) %>% unique()

```


## Formal Model Fits to `motif_count`


Using

- `x = temp -45C`
- use `offset(log(count_total_round))` instead of `male` as a factor

```{r}

temp_ref <- 45
verbose <- 0
trace <- FALSE

## Try filtering the data a bit more
## Goal is to get good starting values

data <- data_ind %>%
    mutate(x1 = (temp - temp_ref)) %>%
    filter(
    (round == 3 & count_total_round >= 400) |
    (round == 2 & count_total_round > 30)
    ) %>%
    mutate() %>%
#    filter( !(male %in% c("T231", "T260"))) %>% 
    mutate()

make_plot = TRUE;

glm_poisson_1 <- glm(motif_count ~
                       (1 + male  +  (x1) + I(x1^2)),
                   data = data,
                   family = poisson(link = "log")
                   )

print(summary(glm_poisson_1))

## Add round effect
glm_poisson_2 <- glm(motif_count ~
                       (1 + male + round +  (x1) + I(x1^2)),
                   data = data,
                   family = poisson(link = "log")
                   )
## Results support round effect
print(summary(glm_poisson_2))

male_coef <-  coef(glm_poisson_2) %>% keep(str_detect(names(.), 'male'))

hist(male_coef, breaks = 30)

## Remove male effect
glm_poisson_3 <- glm(motif_count ~
                       (1 + round +  (x1) + I(x1^2)),
                   data = data,
                   family = poisson(link = "log")
                   )
## Note that using qpoisson doesn't affect MLE

print(summary(glm_poisson_3))

if(make_plot){
ifelse(length(dev.list()) < 3, {dev.new(); dev.next()}, dev.next())
    par(mfrow = c(2, 2))
    plot(glm_poisson_2, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
}

glm_qpoisson_1 <- update(glm_poisson_1,
                       family = quasipoisson(link = "log")
                       )
print(summary(glm_qpoisson_1))

if(make_plot){
    ifelse(length(dev.list()) < 3, dev.new(), dev.next())
    par(mfrow = c(2, 2))
    plot(glm_qpoisson_1, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
##    dev.new()
##    ggplot(data, aes(temp, motif_count))
}

glm_qpoisson_2 <- update(glm_poisson_2,
                       family = quasipoisson(link = "log")
                       )
print(summary(glm_qpoisson_2))

if(make_plot){
    ifelse(length(dev.list()) < 3, dev.new(), dev.next())
    par(mfrow = c(2, 2))
    plot(glm_qpoisson_2, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
##    dev.new()
##    ggplot(data, aes(temp, motif_count))
}


glm_qpoisson_3 <- update(glm_poisson_3,
                       family = quasipoisson(link = "log")
                       )
print(summary(glm_qpoisson_3))

if(make_plot){
    ifelse(length(dev.list()) < 3, dev.new(), dev.next())
    par(mfrow = c(2, 2))
    plot(glm_qpoisson_3, ask = FALSE)
    page_label <- "glm with poisson error"
    mtext(page_label, outer=TRUE,  cex=1, line=-1.1)
##    dev.new()
##    ggplot(data, aes(temp, motif_count))
}

```

### Result

- Overdispersion of data makes parameters non-significant
- Don't how a `motif_prop` approach will solve this issue.
  

