---
title: "Piecewise Regression with NB Error on Real and simulated Data using rstan and BRMS"
author: "Michael Gilchrist"
date: "date: 2023-01-09"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = FALSE, # show warnings
  message = FALSE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=2E2, output = TRUE)

```
# Goal

- Fit two piece nb to data using brms

## Recap

While `brms` can build and run the models, there are serious issues

- We get negative `mu` parameters for the NB.
- We get lots of `NaN` in the output
- I am unsure if I'm actually fitting the model as I intend.
  For example, when using two groups for "x0", I don't get two estimates of "x0"
- I have failed to use male as a fundamental grouping beyond the `separate` setting.

# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(brms)
library(loo)
library(shinystan)


## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Load Data

```{r}
load(file.path("input", "data.processing_2022-12-15.Rda"),
     verbose = TRUE)


motif_data
```
# Process Data
## Create Working Dataset

```{r}

filter_data <- TRUE

if(filter_data) {
  males_filtered_disp <- motif_stats_40C %>%
    filter(dispersion < 50) %>%
    pull(male)

  males_filtered_mean <- motif_stats %>%
    filter(mean > 10) %>% # changing from 10 to 40 removes previous male 7 (T258)
    pull(male)

  male_vector <- intersect(males_filtered_mean, males_filtered_disp)
} else {
  male_vector <- motif_data %>% select(male) %>% distinct()
}

data_ind <- motif_data %>%
    filter(male %in% male_vector) %>%
    mutate(male = droplevels(male)) %>%
    mutate(index = as.integer(male)) %>%
    mutate(male = as.character(male)) %>% 
    arrange(index) %>%
    select(male, index, motif_count, temp_target, temp, round, trial_round, date, counter) %>% 
  ##    left_join(index_shape, by = "index") %>%
  mutate()

stats_ind <- motif_stats %>%
    filter(male %in% male_vector) 


data_ind <- data_ind %>% filter(temp < 38) %>%
    group_by(male) %>% mutate(y0_simple.est = mean(motif_count), phi_ind = var(motif_count)/y0_simple.est) %>% ## phi is overdispersion parameter
    ungroup()

summary(data_ind)

## for calculation of 'theta_bar' (size in `rnbinom`) see note in `Set Up Simulate Data`
summary_stats <- data_ind %>% ungroup() %>%  summarize(y0_bar = mean(y0_simple.est), y0_sd = sd(y0_simple.est), phi_bar = median(phi_ind), theta_bar = y0_bar^2/(y0_sd^2 - y0_bar))

n_male <- length(unique(data_ind$male))

```



## Create Simulated Dataset

- Simulation based on code from "code/2022-12-15_rstan_two.piece.qpoisson.with.simulated.data/rstan_twopiece_fit.various.models.to.simulated.data.Rmd"

From `rnbinom` help
  >  An alternative parametrization (often used in ecology) is by the
  >   _mean_ ‘mu’ (see above), and ‘size’, the _dispersion parameter_,
  >   where ‘prob’ = ‘size/(size+mu)’.  The variance is ‘mu + mu^2/size’
  >   in this parametrization.

So above we can estimate `size = mu^2/(var - mu)`.
We do this above in summary_stats `theta_bar = y0_bar^2/(y0_sd^2 - y0_bar)`.

### Functions Used

```{r}

sim_parms <- function(flag = "uniform_1", n_male = 10, mean_global = 100, sd = 10, n_sd = 2, min = 10, max = 600) {

    mean_low <- mean_global - n_sd * sd
    mean_high <- mean_global + n_sd * sd
    n_low <- floor(n_male/2)
    n_high <- n_male - n_low
    
    switch(flag,
           ## Parameter values are uniform across males within a category (high vs. low)
           uniform_1 = {
               tibble(parm = rep(mean_global, times = n_male), grp = 1L)},
           
           uniform_2 = {
               tibble(parm = c(rep(mean_low, times = n_low),
                              rep(mean_high, times = n_high)),
                      
                      grp =  c(rep(1L, times = n_low), 
                               rep(2L, times = n_high))
                      )},
           ## Parameter values vary between males, but pulled from distribution common to each category (high vs. low)
           ## Note syntax differs from model usage (this would 'be pooled')
           ## Should update model usage
           groups_1 = {
               tibble(parm = rnorm(n_male, mean_global, sd),
                      grp = 1L
                      )},
           groups_2 = {
               tibble(parm = c(rnorm(n_low, mean_low, sd),
                               rnorm(n_high, mean_high, sd)),
                      grp =  c(rep(1L, times = n_low),
                               rep(2L, times = n_high))
                      )}
           ) %>%
        mutate(parm = round(parm, 1)) %>%
        mutate(parm = pmax(min, parm)) %>% ## ensure parm is not below `min`
        mutate(parm = pmin(max, parm)) %>% ## ensure parm is not below `max`
        #mutate(parm = as.integer(parm)) %>%
        slice_sample(n = n_male, replace = FALSE) %>%
        mutate(index = row_number())
}

var_temp_mean <- function(temp_target) {
    ## Returns expected variance in temp_mean given temp_target
    ## numbers from analysis in `code/2023-01-13_temp_target.vs.var/compare.temp_target.vs.var.Rmd`
    var = max(0.891, -5.16 + 0.167 * temp_target)

    return(var)
}

sd_temp_mean <- function(temp_target) {
    sqrt(var_temp_mean(temp_target))
    }


sim_nb_counts <- function(temp, y0, x0, tmax, theta) {
    ## Calculate expected value given parametes
    mu <- ifelse(temp < x0, y0, y0 * (1-(temp-x0)/(tmax-x0))) 
                                        #print(paste("size: ", size, "count: ", count))
    rnbinom(1, size = theta, mu = mu)
}


sim_qpoisson_counts <- function(temp, y0, x0, tmax, phi) {

    ## Calculate expected value given parametes
    mu <- ifelse(temp < x0, y0, y0 * (1-(temp-x0)/(tmax-x0))) 
    
    ## calculate theta parameter based on mu and phi
    theta = mu/(phi - 1)
    #print(paste("size: ", size, "count: ", count))
    rnbinom(1, size = theta, mu = mu)
    }


```

### Generate Parameter Tibble

- Parameter values in this tibble will be used to create simulated data.

```{r}
phi <- summary_stats$phi_bar ## overdispersion parameter
theta <- summary_stats$theta_bar ## size parameter
y0_bar <- summary_stats$y0_bar
y0_sd <- summary_stats$y0_sd
tmax <- 46
n_male <- 12


## Generate 'true' parameters
## If TRUE, replace estimated y_0 with simulated values,
## else use estimates from observed data 
y_flags <- c("uniform_1", "uniform_2", "groups_1", "groups_2")
x_flags <- y_flags

data_sim <- tibble()

replicates <- c(2, 8, 32)
temp_target <- c(30, 38, 40, 42, 44)

set.seed(2022)

for(sampling_dist in c("qpoisson", "nb")) {
    for(x_flag in x_flags) {
        for(y_flag in y_flags) {
            for(n_reps in replicates) {

                y0_sim <- sim_parms(flag = y_flag, n_male = n_male, mean = y0_bar, sd = y0_sd) %>%
                    rename(y0 = parm, yy = grp)
                x0_sim <- sim_parms(flag = x_flag, n_male = n_male, mean = 39, sd = 1, min = 35, max = 45.9)  %>%
                    rename(x0 = parm, xx = grp)

                x0_y0_sim <- full_join(x0_sim, y0_sim, by = "index") %>%
                    relocate(index) ## move index to first column


                ## Create relevant subset data matrix
                data_parms <- crossing(sampling_dist = sampling_dist,
                                       x0_y0_sim,
                                       temp_target = temp_target,
                                       n_reps = n_reps,
                                       rep = 1:n_reps,
                                       x_flag = x_flag,
                                       y_flag = y_flag)

                data_parms <- data_parms %>%
                    rowwise() %>%
                    mutate(temp_sd = sd_temp_mean(temp_target),
                           temp_mean = min(rnorm(1, mean = temp_target, sd = temp_sd), 45.9)) %>% 
                    select(-temp_sd)

                summary(data_parms)
                
                data_tmp <- switch(sampling_dist,
                                   qpoisson = {
                                       data_parms %>%
                                           mutate(motif_count= sim_qpoisson_counts(temp_mean, y0, x0, tmax, phi))},
                                   nb = {
                                       data_parms %>%
                                           mutate(motif_count = sim_nb_counts(temp_mean, y0, x0, tmax, theta))
                                   }
                                   )
                
                if(nrow(data_sim) == 0) {
                    data_sim <- data_tmp
                } else {
                    data_sim <- bind_rows(data_sim, data_tmp)
                }
            }
        }
    }
}



```


# Fit Models
## Common Parameters
 
```{r}

iter <- 4000
tmax <- 46
t0max <- tmax - 0.5;
t0min <- 20;

y_xmax <- 0
y0_min <- 10
sd_y0_prior <- 200
alpha_theta_prior <- 10 ## exponential dist scale parameter for overdispersion theta in quasipoisson
alpha_phi_prior <- 0.01 ## gamma dist shape parameter for nb. brms default is 0.01
beta_phi_prior <- 0.01 ## gamma dist rate parameter for nb. brms default is 0.01


## values to use for model predictions
tp = seq(25, tmax, length.out = 100)
n_cores <- 4
n_chains <- n_cores
```
## Indicate Dataset to Use: Real vs Simulated

```{r}

data_type <- "sim"
x_f <- "uniform_1"
y_f <- "uniform_1"
sampling_d <- "nb"
n_r <- 32

data <- switch(data_type,
               real = data_ind,
               sim = {
                   data_sim %>% filter(x_flag == x_f,
                                       y_flag == y_f,
                                       sampling_dist == sampling_d,
                                       n_reps == n_r) %>%
                       rename(temp = temp_mean)
               }
               )
                                       

motif_count <- data %>% pull(motif_count)
temp <- data %>% pull(temp)
N <- length(temp)
index <- data %>% pull(index) ## male identifier
## parameters to be printed

```


## Fitting Routines
 
```{r}

##y0_grouping <- map_int(data$male, ~ if_else(. %in% y0_group[[1]], 1, 2))
models <- c("rstan", "brms")

## Define groups

flags <- c("separate", "grouping_1", "pooled")

#flags <- c("pooled")
#flags <- c("grouping_1", "pooled")
models <- c("brms")
flags_x <- flags
flags_y <- flags
fit_tbl <- crossing(model = models,
                    x0 = flags_x, y0 = flags_y,
                    desc = "NA_character",
                    y0_group_list = list(NA),
                    x0_group_list = list(NA),
                    fit = list(NA),
                    llik = list(NA),
                    r_eff = list(NA),
                    loo = list(NA)
                    )


stan_file <- "two.piece_nb_1.1.stan"
model_name <- "two.piece.rstan"
## For debugging
##cmodel <- cmdstan_model(stan_file = stan_file)

## To avoid it being continually recompiled
if(! exists("rstan_model")) {
    rstan_model <-stan_model(file = stan_file,
                             verbose = FALSE,
                             model_name = model_name
                             )
}

for(model in models) {
    for(x_flag in flags_x) {
        for(y_flag in flags_y) {
            
            desc <- paste0(model, ": ", x_flag, ", ", y_flag)
            curr_row <- which(fit_tbl$model == model &
                              fit_tbl$x0 == x_flag &
                              fit_tbl$y0 == y_flag)
            
            fit_tbl[ curr_row, ]$desc <- desc
            print(desc)
            
            x0_group_list <- list()
            y0_group_list <- list()
            
            switch(x_flag,
                   separate = {
                       x0_group_list <- data$male %>% unique() %>% as.list()
                   },
                   grouping_1 = {
                       ## set up groupings based on 2022-12-20 analysis
                       ## Using male ID's instead index to make code more robust
                       x0_group_list[[1]] <- c("T235", "T237", "T244", "T247", "T257", "T260")
                       x0_group_list[[2]] <- c("T234", "T236", "T243", "T246", "T258")
                   },
                   pooled = {
                       x0_group_list[[1]] <- data$male
                   }
                   )


            switch(y_flag,
                   separate = {
                       y0_group_list <- data$male %>% unique() %>% as.list()
                   },
                   grouping_1 = {
                       ## set up groupings based on 2022-12-20 analysis
                       ## Using male ID's instead index to make code more robust
                       y0_group_list[[1]] <- c("T234", "T243", "T244", "T246", "T258", "T260") 
                       y0_group_list[[2]] <- c("T235", "T236", "T237", "T247", "T257")
                   },
                   pooled = {
                       y0_group_list[[1]] <- data$male
                   }
                   )

            fit_tbl[ curr_row, ]$x0_group_list[[1]] <- x0_group_list

            fit_tbl[ curr_row, ]$y0_group_list[[1]] <- y0_group_list    

            ## Convert lists to a vector of concatenated strings
            ## This will simplify mapping male to an x0/y0 index 
            x0_group <-lapply(x0_group_list, paste, collapse = " ") %>% unlist()
            y0_group <-lapply(y0_group_list, paste, collapse = " ") %>% unlist()

            x0_index <- sapply(as.character(data$male), function(x) str_which(x0_group, x))
            y0_index <- sapply(as.character(data$male), function(x) str_which(y0_group, x))

            switch(model,
                   rstan = {
                     
                       data_list <-list(x = temp,
                                        y = motif_count,
                                        N = N,
                                        X = length(x0_group),
                                        Y = length(y0_group),
                                        NB = 1,
                                        xx = x0_index,
                                        yy = y0_index,
                                        ## male_index = index,
                                        nbb = rep(1,N),
                                        xmax = tmax,
                                        x0_min = t0min,
                                        x0_max = t0max,
                                        y_xmax = y_xmax,
                                        y0_min = y0_min,
                                        sd_y0_prior = sd_y0_prior,
                                        alpha_theta_prior = alpha_theta_prior,
                                        alpha_phi_prior = alpha_phi_prior,
                                                beta_phi_prior = beta_phi_prior
                                        ##tp = tp,
                                        ## max threshold value.
                                        ## having it too close to xmax *sometimes* leads to sampling
                                        ## near xmax, but with lower lp and very high E13) b0 values
                                        )
                       
                       fit <- stan(file = stan_file,
                                   model_name = model_name,
                                   data = data_list,
                                   cores = n_cores,
                                   chains = n_chains,
                                   iter = iter,
                                   warmup = floor(iter/2),
                                   verbose = FALSE)
                       
                   },
                   brms = {

                       ## Can't use normal R function within stan
                       ## Need to define functions in stan language
                       ## in a character string that's passed to brms
                       
                       df <- data.frame(
                           x = temp,
                           y = motif_count,                   
                           xx = x0_index,
                           yy = y0_index,
                           id = index
                       )

                       my_priors <- prior(normal(150, 200), nlpar = "y0", lb = 0.1) +
                         ## same as default, but defining here to document syntax
                         prior(gamma(0.01, 0.01), class = "Intercept", dpar = "shape", lb = 0) +
                         prior(uniform(25, 45.5), lb = 25, ub = 45.5, nlpar = "x0") + 
                         prior(uniform(0.1, 10), lb = 0.1, ub = 4, class = "sd", group = "xx", nlpar = "x0") + 
                         prior(uniform(0.1, 200), lb = 0.1, ub = 200, class = "sd", group = "yy", nlpar = "y0") #+
        #                 prior(constant(46), nlpar = "xmax")
                       ## Have to hard code xmax = 46 and y0_xmax = 0
                       nlform <- bf(y ~  0 - (46 - fmax(x0, x)) * (0 - y0)/(46 - x0),
                                    shape ~ 1,
                                    x0 ~ 1 + (1|xx),#|index),
                                    y0 ~ 1 + (1|yy),#|index),
                                    nl = TRUE)
                       
                       fit <- brm(nlform,
                                  data = df,
                                  ## `link` refers to the mapping of the expectation of the distribution: log, sqrt, identity, softplus
                                  ## link_shape corresponds to `phi` of `stan`'s
                                  ## Negbinomial_2
                                  ## Defining `phi = mu/theta` creates a quasipoisson
                                  ## distribution with overdispersion parameter (1 +theta)
                                  family = negbinomial(link = "identity", link_shape = "identity"),
                                  prior = my_priors,
                                  iter = iter,
                                  control = list(adapt_delta = 0.975),
                                  save_model = "brms_two.piece.stan")
                       ## Avoid having brms recompile model by defining
                       ## model in global environment
                       #if(!exists("stanfit"))
                       stanfit <- fit$fit

                   }
                 )
            print(fit)
            fit_tbl[ curr_row, ]$fit <- list(fit)
            
        }
    }
}

nb_fit_tbl <- fit_tbl

## save(file = "fit_tbl.Rda", fit_tbl)

```  

# Model Comparison

Below doesn't work for a number of reasons

- Need to use log_lik() with `brmsfit` objects
- Mostly `NaN` in the `brmsfit` fits `log_lik()` output.

## LOO Analysis

```{r}

## define fmax() as used in stan
## This is necessary for using loo()
fmax <- function(x, y) max(x,y)

fit_tbl <- nb_fit_tbl

for(curr_row in 1:length(fit_tbl$fit)) {

  desc <- fit_tbl[[curr_row, "desc"]]
  fit <- fit_tbl[[curr_row, "fit"]][[1]]
  
  if(class(fit) == "brmsfit") fit <- fit$fit
  
  print(paste0("Model ", curr_row, ": ", desc))

  # loo analysis based on: http://mc-stan.org/loo/articles/loo2-with-rstan.html
  #
  # Extract pointwise log-likelihood
  # using merge_chains=FALSE returns an array, which is easier to 
  # use with relative_eff()
  llik <- extract_log_lik(fit, parameter_name = "target", merge_chains = FALSE)
  fit_tbl[[curr_row, "llik"]] <- list(llik)

  # as of loo v2.0.0 we can optionally provide relative effective sample sizes
  # when calling loo, which allows for better estimates of the PSIS effective
  # sample sizes and Monte Carlo error
  r_eff <- relative_eff(exp(llik), cores = n_cores)
  fit_tbl[[curr_row, "r_eff"]] <- list(r_eff)
  
  # preferably use more than 2 cores (as many cores as possible)
  # will use value of 'mc.cores' option if cores is not specified
  loo <- loo(llik, r_eff = r_eff,
             cores = n_cores,
             save_psis = TRUE,
             moment_match = TRUE) 
  fit_tbl[[curr_row, "loo"]] <- list(loo)
  print(loo)

}


comp <- loo_compare(fit_tbl$loo)
index <- comp %>% rownames() %>% sub(pattern = "model", x= ., "") %>% as.integer() 
desc <-fit_tbl$desc[index]
rownames(comp) <- desc
    
#loo_tbl<- bind_cols( desc = desc, index = index, comp) %>% tibble() %>% arrange(index)

print(comp)

```
## Plot Results

- Incomplete; doesn't work.

```{r}

pars <- c("t0", "y0")
pars_full <- c(pars, "lp__")


for(curr_row in 1:length(fit_tbl$fit)) {

    desc <- fit_tbl[[curr_row, "desc"]]
    fit <- fit_tbl[[curr_row, "fit"]]

    x0_group_list <- fit_tbl[[curr_row, "x0_group_list"]][[1]]
    y0_group_list <- fit_tbl[[curr_row, "y0_group_list"]][[1]]

    
}

    
    pars <- c("x0", "y0", "theta")
    pars_full <- c(pars, "lp__")

    print(desc)
    print(fit, pars = pars)
                                        #traceplot(model, pars = pars, inc_warmup = FALSE)
                                        #plot(model, pars = pars) #, ggtitle(title))

        pairs(model, pars = pars_full)

        ## Plot parameter estimate summaries
        tmp_plot <- list()
        for(par in pars) {
            tmp_plot[[par]] <- stan_plot(model, pars = par)
        }
        gt <- arrangeGrob(grobs = tmp_plot)
        as_ggplot(gt)

    }
}
```


