---
title: "Fit Thermal Models"
author: "Michael Gilchrist"
date: date: "`r Sys.time()`"
output: pdf_format
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Goal

- Fit series of thermal models found in `rTPC` to data collected in the Derryberry lab.
- Based on initial `.R` file created by Kayci Messerly `<messerlykayci1@gmail.com>`
- Intial analysis should be with GLM and mixture models.



# Set up

## Install libraries

```{r}
## install packages user might not have by replacing FALSE with TRUE
if(FALSE) {
    BiocManager::install("mixOmics") ## needed by RVAideMemoire
    install.packages(c("RSQLite", "nls.multstart", "lme4", "RVAideMemoire"))
    ##  Install the thermal curve package from git_hub, not cran
    remotes::install_github("padpadpadpad/rTPC")

}

## load libraries
library(stats)
require(MASS) # provides negative binomial fitting:  glm.nb
library(RSQLite)  # Don't think we need this.
library(rTPC)  ## 
library(nls.multstart)
library(broom)
library(tidyverse)
require(ggplot2)
require(GGally)
require(reshape2)
require(lme4)
require(RVAideMemoire) # provides overdisp.glmer()

```

## Load Data

```{r}

## Read in ZEFI Data sets
## Treat 'repeatability' as round = 0
## Add round info

## Repeatability was done between round 1 and 2, female was present, but only one temp. so treating as `round = 2` and redefining `round = 2` as `round = 3`

git_root <- system("git rev-parse --show-toplevel", intern = TRUE)

data_raw = list()

data_raw[[1]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Round-1-Heat-Trials.csv")) %>% mutate(round = 1) %>%
    ## Note T237 and T230 are missing numbers in the motif_count column
    ## so we are filtering these observations out until they are found
    filter(!is.na(motif_count))

data_raw[[2]] <- read.csv(file.path(git_root, "data", "collated", "HSPi-Repeatability-Song-Count.csv")) %>%
    mutate(round = 2) %>%
    group_by(male) %>%
    mutate(test_order = rank(date)) %>%
    ungroup()

data_raw[[3]] <-read.csv(file.path(git_root, "data", "collated", "HSPi-Round-2-Heat-Trials.csv")) %>% mutate(round = 3)

## Join data and discard empty columns
data_full <- full_join(data_raw[[1]], data_raw[[2]]) %>%
    full_join(data_raw[[3]]) %>%
    discard(~all(is.na(.) | . =="")) %>% ## get rid of columns of only NA
    mutate(trial_completed = !(is.na(motif_count)) ) %>%
    mutate(motif_count = ifelse(is.na(motif_count), 0, motif_count)) %>%
    mutate(motif_count = motif_count*1.0) %>% ## convert to a double so it's not treated as an integer
    mutate(chamber = as.factor(chamber), male = as.factor(male)) %>%
    ## create a global variable trial_order based on individual rounds
    mutate(trial_index = as.integer(round*10+test_order)) %>%
    mutate(motif_count_plus_1 = (motif_count + 1)) %>%
    mutate(log_motif_count_plus_1 = log(motif_count + 1)) %>%
    mutate(temp_target = as.numeric(temp_target)) %>%
    mutate() ## Dummy function so we can comment out lines above it w/o any issues


```

# Analysis


## Examine Data

```{r}

data <- data_full #%>% filter(round ==3)

data %>% group_by(temp_target) %>%
    filter(motif_count < 400) %>%
    summarize(ave = mean(motif_count), sd = sd(motif_count), n = length(motif_count), n_completed = sum(trial_completed))

## Look at occurrance of males across rounds
tmp1 <- data_full %>%
    filter(round==1) %>%
    pull(male) %>%
    unique(); tmp1

tmp2 <- data_full %>%
    filter(round %in% c(2,3)) %>%
    pull(male) %>%
    unique() ; tmp2

lapply(list(tmp1, tmp2, intersect(tmp1, tmp2)), length)

```

# Plot Data


## Raw Plot

```{r}


xlab <- "Temperature"
ylab <- "motif_count"

plot_data_only <-
    ggplot(data_full) + 
    aes(x = temp_target,
        y = motif_count,
        color = as.factor(round)) +
    facet_wrap("male", scales = "free_y") +
    geom_jitter() +
    labs(x = xlab, y = ylab,
         title = "Song Count vs Temp: Data Only") 

pdf(file = "motif_count_vs_temp_data_only.pdf")

plot_data_only

dev.off()
```

## Model fits via `stat_smooth`

- This is what I shared with Liz and Kayci

### `log(motif_count_plus_1)`

```{r}


## Plotting based on: https://www.tjmahr.com/plotting-partial-pooling-in-mixed-effects-models/

xlab <- "Temperature"
ylab <- "log(motif_count + 1)"

#pdf(file = "log-plus-1_vs_temp_all.pdf")

plot_log_plus_1 <- ggplot(data_full) + 
    aes(x = temp_target,
        y = log_motif_count_plus_1,
        color = as.factor(round))  +
    stat_smooth(method = "lm",
                se = TRUE,
                formula = y ~ x + I(x^2), size = 1,
                ) +
  # Put the points on top of lines
  #  geom_point() +
    geom_jitter() +
  facet_wrap("male", scales = "free_y") +
    labs(x = xlab, y = ylab,
         subtitle = "lm:y ~ 1 + x + x^2",
         title = "log(count + 1) vs Temp")

#dev.off()



```

### Song Count 
```{r}


## Code below used for posting to stackexchange
data_full %>% filter(male %in% c("T258", "T259"))%>% dplyr::select(male, temp_target, motif_count) %>% rename(ID = male, x = temp_target, y = motif_count)  %>% dput()

## Choose some males to use for starting values based on log fit
reasonable.males = c("T229", "T237", "T258")

glm_combined <- glm(motif_count ~ 1 + temp_target + I(temp_target^2),
                    family = quasipoisson(),
    data = data_full %>% filter(male %in% reasonable.males))

xlab <- "Temperature"
ylab <- "motif_count"

plot_glm <- plot_data_only +
    aes(x = temp_target,
        y = motif_count,
        color = (round == 1)) + ## Combine round 2 and 3
    stat_smooth(method = "glm",
                method.args = list(
                    family = quasipoisson(link = "log"),
                    start = coefficients(glm_combined),
                    maxit = 100),
                se = FALSE,
                formula = y ~ 1 + x + I(x^2), size = 1,
                ) +
    labs(x = xlab, y = ylab,
         subtitle = "glm:y ~quasipoisson(exp[1 + x + x^2])",
         title = "Song Count vs Temp: Quasi-Poisson Regression") 

## Rescale axes to deal with poor GLM fit that messes up plotting of T259
p1 <- ggplot_build(plot_data_only)
p2 <- ggplot_build(plot_glm)


## original post suggested using
## p2$layout$panel_scales_y <- p1$layout$panel_scales_y

## Only replace problematic scale
p2$layout$panel_params[[14]] <- p1$layout$panel_params[[14]]

pdf(file = "motif_count_vs_temp_multiple.pdf")

plot_data_only

plot_log_plus_1

grid::grid.newpage()
grid::grid.draw(ggplot_gtable(p2))

dev.off()

```



# Fit models outside of plotting -- not shared with Liz

- In `round = 3`, 
  - The only `trial_completed = FALSE` is in highest temp.
  - Only one of 6 chambers was used for a given male, thus chamber and male are conflated.
  - Actually makes life easier!

```{r}

data_r3 <- data_full %>% filter(round ==3)

```
## LM Fits on `log(motif_counts_plus_one)`

- None of these models fit very well



### Fit each individual separately

- Learning how to use `lmList()` function

```{r}

## Individual fits to each male
## This isn't that useful, but is my firstuse of lmList()

lmList <- list()

lmList$by_male <- tmp <-
    lme4::lmList(log_motif_count_plus_1 ~ 1 + temp_target + I(temp_target^2) | male,
           data = data %>% filter(round %in% c(2,3)),
           na.action = na.omit)
           

summary(tmp)

plot.lmList4(tmp, log_motif_count_plus1 ~ fitted(.)| male)


```
### Fit Models

```{r}


lm = list()

formula <- as.formula("log_motif_count_plus_1 ~ 1 + temp_target + I(temp_target^2)")
## Note syntax below model `formula` doesn't appear in the model summary().
## lm$basic <- lm(formula = formula , data = data_r3)

## Here string of formula is shown in summary()
## The .() forces the evaluation in bquote()
## See: https://win-vector.com/2018/09/01/r-tip-how-to-pass-a-formula-to-lm/
## for detailed explanation of how this works
lm$basic <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp)

## Add trial_index effect
formula <- as.formula("log_motif_count_plus_1 ~ 1 + trial_index+ temp_target + I(temp_target^2)")
lm$trial_index <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3) ))
summary(tmp)

## Add male effect 
formula <- as.formula("log_motif_count_plus_1 ~ 1 + male + temp_target + I(temp_target^2)")
lm$male <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp) 

## Add chamber effect (fewer chambers than males)
formula <- as.formula("log_motif_count_plus_1 ~ 1+ chamber + temp_target + I(temp_target^2)")
lm$chamber <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3)))
summary(tmp)

## Add trial_index and male effects
formula <- as.formula("log_motif_count_plus_1 ~ 1 + trial_index + male + temp_target + I(temp_target^2)")
lm$trial_index_and_male <- tmp <- eval(bquote(lm(formula = .(formula) , data = data_r3) ))
summary(tmp)


## Add male x trial order
## This doesn't work since we have as many coefficients as data points.
## Oops
formula <- as.formula("log_motif_count_plus_1 ~ 1 + male*trial_index + temp_target + I(temp_target^2)")
lm$trial_index_by_male <- tmp <- lm(formula = eval(formula) , data = data_r3)
summary(tmp)


## add slope x male interaction term

lm$male_by_temp <- tmp <- lm(formula = log_motif_count_plus_1 ~ 1 + temp_target*male + I(temp_target^2), data = data_r3)
summary(tmp)


## slope x male - male intercept
## This suggests that the slopes are similar between males (though the SE are large)
lm$male_by_temp <- tmp <- lm(formula = log_motif_count_plus_1 ~ 1 + temp_target:male + I(temp_target^2), data = data_r3)
summary(tmp)

## curvature x male - male intercept
## This suggests that the slopes are similar between males (though the SE are large)
lm$male_by_temp_sq <- tmp <- lm(formula = log_motif_count_plus_1 ~ 1 + temp_target + I(temp_target^2):male, data = data_r3)
summary(tmp)


```

### Compare Models

```{r}


## Results: Male is much better predictor than trial_index
anova(lm$male, lm$trial_index)

## Results: Male is better predictor than chamber.
## This indicates males vary more than chambers
anova(lm$male, lm$chamber)

## Results: Adding trial_index to male provides no real gain in fit
anova(lm$male, lm$trial_index_and_male)
```

### Notes

- We do see evidence of male effects.
  The important males have lower intercepts than the rest
- There is little evidence of consistent trial_index effects.
- Adjusted $R^2$ don't seem to vary much.


## GLM and GLMM Poisson, Quasipoisson and Negative-Binomial Fits



## GLM on Song Rate

- See [webpage](https://stats.oarc.ucla.edu/r/faq/random-coefficient-poisson-models/)
  for mixed-effects poisson models
  
```{r}

glm <- list()

## chamber and male are conflated
## Data is very overdispersed given poisson assumptions
glm$poisson <- glm(motif_count ~ 1 + male + trial_index + temp_target + I(temp_target^2), data = data_r3, family = poisson(link = "log"))

## This allows for over dispersion which greatly elevates the SE
glm$qpoisson <- glm(model$glm_poisson, family = quasipoisson(link = "log"))

## Gives similar results as with the quasipoisson
glm$nb <- glm.nb(model$glm_poisson)

anova.list(glm)
```

## GLMM (glmer) Model fits

```{r}

glmer = list()

glmer$poisson <- tmp <-
    glmer(
        motif_count ~
            1+ 1|male + temp_target + I(temp_target^2),
        data = data_r3,
        family = poisson(link = "log"),
        control = glmerControl(optCtrl = list(maxiter = 300)),
        verbose = TRUE)

## Still a lot of overdispersion
glmer$od <- overdisp.glmer(glmer$poisson)

## I think this is a correct fit.  If I don't include the 1 + temp... w/o the |male, I don't get any effects.
glmer$nb_1 <- glmer.nb(
    motif_count ~ 1 + temp_target + I(temp_target^2) + trial_index + ((1 + temp_target + I(temp_target^2))|male),
    data = data_r3,
    family = poisson(link = "log"),
    control = glmerControl(optCtrl = list(maxiter = 300)),
    verbose = TRUE)


glmer$nb_2 <- update(glmer$nb_1, formula =  motif_count ~ 1 + temp_target + I(temp_target^2) + ((1 + temp_target + I(temp_target^2))|male) )

mode
```

