---
title: "Fit rTPC models to `motif_prop` after filtering"
author: "Michael Gilchrist"
date: "date: 2022-10-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
                      warning = TRUE, # show warnings
                      message = TRUE, # show messages
                      error = TRUE, # do not interrupt generation in case of errors,
                      echo = TRUE#,  # show R code
                      ##results="asis" # Hopefully print summary() in for loops
                  )


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=1E5, output = TRUE)

```
# Goal

- Evaluate trends in motif_count under a (near) constant temperature.

# Set up

## Load libraries

```{r}


## load libraries
library(stats)
require(MASS) # provides negative binomial fitting:  glm.nb
library(RSQLite)  # Don't think we need this.
library(rTPC)  ## 
library(nls.multstart)
library(broom)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(grid) ## provides textGrob
library(gridExtra)
library(viridisLite)

                                        #options(ggplot2.continuous.colour="viridis",
                                        #        ggplot2.discrete.colour="viridis",
                                        #        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
                                        #        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(GGally)
library(reshape2)
library(lme4)
library(nlme)
library(gnm)
library(rsample) ## provides bootstraps()

library(RVAideMemoire) # provides overdisp.glmer()
library(humidity) ## provides VPD
library(weathermetrics)
library(latex2exp)

```
## Local Functions

- Copied from brms-first.fitting.Rmd


```{r}
## Taken from: https://stackoverflow.com/a/51330864/5322644
## Use to get model equations for models in rTPC

help_text <- function(...) {
    file <- help(...)
    path <- dirname(file)
    dirpath <- dirname(path)
    pkgname <- basename(dirpath)
    RdDB <- file.path(path, pkgname)
    rd <- tools:::fetchRdDB(RdDB, basename(file))
    capture.output(tools::Rd2txt(rd, out="", options=list(underline_titles=FALSE)))
}

get_model_eq <- function(model) {
    txt <- help_text(model)
    eqn_line <- grep("^ +rate = .*$", txt, value = TRUE)
    #print(paste(model, eqn_line))
    eqn <- gsub("(^ +rate = | *$)", "", eqn_line) %>%
        gsub("([^0-9])\\.(\\w+)", "\\1 * \\2", .) %>%
        gsub("\\.([^0-9])", " * \\1", .) %>%
        gsub("\\_", "", .)  %>%
        gsub(" / ", "/", .)  %>%
        sub("(^ +| +$)", "", .) %>%
        paste(., collapse = "; ")
#    print(eqn)
    df <- tibble(model = model, eq = eqn)
    return(df)
}


```
- Create `nls_multstart_fit_db()` to get all model fits.

```{r}

nls_multstart_db <- function (formula, data = parent.frame(), iter, start_lower, 
    start_upper, supp_errors = c("Y", "N"), convergence_count = NULL, 
    control, modelweights, verbose = FALSE, ...) {
    if (missing(supp_errors)) {
        supp_errors <- "N"
    }
    formula <- stats::as.formula(formula)
    if (any(class(data) %in% c("data.frame", "list", "environment"))) {
        params_ind <- all.vars(formula[[3]])[all.vars(formula[[3]]) %in% 
            names(data)]
        params_est <- all.vars(formula[[3]])[!all.vars(formula[[3]]) %in% 
            names(data)]
        params_dep <- all.vars(formula[[2]])
    }
    else {
        stop("data should be a data.frame, list or an environment")
    }
    if (missing(start_lower) || missing(start_upper)) {
        cat("No boundaries specified for the starting values of sought parameters. \n", 
            "Default values of +/- 1e+10 will be used. This is likely \n", 
            "to slow the process of finding the best model. \n")
        r <- readline("Continue with default values [y/n]? ")
        if (tolower(r) == "n") {
            stop("Please enter upper and lower parameter boundaries as start_lower and start_upper in function argument.")
        }
    }
    if (missing(start_lower)) 
        start_lower <- rep(-10^10, length(params_est))
    if (missing(start_upper)) 
        start_upper <- rep(10^10, length(params_est))
    if (length(start_lower) != length(params_est) || length(start_upper) != 
        length(params_est)) {
        stop("There must be as many parameter starting bounds as there are parameters")
    }
    params_bds <- data.frame(param = params_est, low.bds = unlist(start_lower), 
        high.bds = unlist(start_upper), stringsAsFactors = FALSE)
    if (missing(control)) {
        control <- minpack.lm::nls.lm.control(maxiter = 1000, 
            ftol = .Machine$double.eps, ptol = .Machine$double.eps)
    }
    silent <- ifelse(supp_errors == "Y", TRUE, FALSE)
    if (silent == TRUE) {
        oo <- options(warn = -1)
        on.exit(options(oo))
    }
    if ("modelweights" %in% all.vars(formula)) {
        stop(paste0("The variable name 'modelweights' is reserved for model weights. Please change the name\n", 
            "of this variable"))
    }
    if (missing(modelweights)) {
        data$modelweights <- rep(1, length(data[[params_dep]]))
    }
    else {
        data$modelweights <- eval(substitute(modelweights), data)
    }
    if (length(iter) == 1) {
        multistart_type <- "shotgun"
    }
    else if (length(iter) == nrow(params_bds)) {
        multistart_type <- "gridstart"
    }
    else {
        stop(paste0("iter should be of length 1 for shotgun approach and of the same length as the\n", 
            "number of parameters for the gridstart approach."))
    }

    if (multistart_type == "shotgun") {
        if(verbose) print("Setting up shotgun strt...")
        strt <- purrr::map2(params_bds$low.bds, params_bds$high.bds, 
            ~runif(iter, .x, .y))
        names(strt) <- params_bds$param
        strt <- dplyr::bind_rows(strt)
    }
    if (multistart_type == "gridstart") {
        params_bds$iter <- iter
        strt <- purrr::pmap(as.list(params_bds[, -1]), function(low.bds, 
            high.bds, iter) seq(from = low.bds, to = high.bds, 
            length.out = iter))
        names(strt) <- params_bds$param
        strt <- tibble::as_tibble(expand.grid(strt))
    }

    if(verbose) {
        print(paste0("strt values for ", multistart_type))
        print(strt)
    }

    strt$iteration <- 1:nrow(strt)
    strt <- tidyr::nest(strt, startpars = -iteration)

    fit <- NULL
    fit_best <- NULL
    get_fit <- function(start.vals) {
        try(fit <- minpack.lm::nlsLM(formula, start = start.vals, 
                                         control = control, data = data, weights = modelweights, 
                                         ...), silent = silent)
        return(fit)
    }

    browser()
    
    fits <- purrr::map(strt$startpars, get_fit) %>% tibble("fit" = .)
    fit_db <- bind_cols(strt, fits) %>% mutate( aic = ifelse(!is.null(fit), stats::AIC(fit), 
                                 Inf) )
                                        #dplyr::mutate(allfits,
                            ## purrr works, but fit is embedded in a list.
                            ## fit = purrr::map(startpars, get_fit)) ##, AICval = get_aic(fit))
                                        #fit = get_fit(unlist(startpars)) ) ##, AICval = get_aic(fit))
    
    return(fit_db)
    
}

```

- Copied from: (https://rmazing.wordpress.com/2012/07/19/a-weighting-function-for-nls-nlslm/)
- Key usage: `wfct(1/fitted^2)` which uses the fitted value in nls

```{r}

wfct <- function(expr)
{
    expr <- deparse(substitute(expr))

    ## create new environment
    newEnv <- new.env()

    ## get call
    mc <- sys.calls()[[1]]
    mcL <- as.list(mc)

    ## get data and write to newEnv
    DATA <- mcL[["data"]]
    DATA <- eval(DATA)
    DATA <- as.list(DATA)
    NAMES <- names(DATA)
    for (i in 1:length(DATA)) assign(NAMES[i], DATA[[i]], envir = newEnv)

    ## get parameter, response and predictor names
    formula <- as.formula(mcL[[2]])
    VARS <- all.vars(formula)
    RESP <- VARS[1]
    RHS <- VARS[-1]
    PRED <- match(RHS, names(DATA))
    PRED <- names(DATA)[na.omit(PRED)]

    ## calculate variances for response values if "error" is in expression
    ## and write to newEnv
    if (length(grep("error", expr)) > 0) {
        y <- DATA[[RESP]]
        x <- DATA[[PRED]]
        ## test for replication
        if (!any(duplicated(x))) stop("No replicates available to calculate error from!")
        ## calculate error
        error <- tapply(y, x, function(e) var(e, na.rm = TRUE))
        error <- as.numeric(sqrt(error))
        ## convert to original repititions
        error <- rep(error, as.numeric(table(x)))
        assign("error", error, envir = newEnv)
    }

    ## calculate fitted or residual values if "fitted"/"resid" is in expression
    ## and write to newEnv
    if (length(grep("fitted", expr)) > 0 || length(grep("resid", expr)) > 0) {
        mc2 <- mc
        mc2$weights <- NULL
        MODEL <- eval(mc2)
        fitted <- fitted(MODEL)
        resid <- residuals(MODEL)
        assign("fitted", fitted, newEnv)
        assign("resid", resid, newEnv)
    }

    ## return evaluation in newEnv: vector of weights
    OUT <- eval(parse(text = expr), envir = newEnv)
    return(OUT)
}

## Plotting settings

```{r}

## From: https://data-se.netlify.app/2018/12/12/changing-the-default-color-scheme-in-ggplot2/

theme_set(theme_minimal(base_size = 9))
theme_update(
    plot.title = element_text(size = rel(1.1)),
    plot.subtitle = element_text(size = rel(1)))

if(!exists("old_opts")) old_opts <- options()  # save old options

options(ggplot2.continuous.colour="viridis")
options(ggplot2.continuous.fill = "viridis")
options(ggplot2.discrete.colour="viridis")
options(ggplot2.discrete.fill = "viridis")


```
## Create Model Tibble

```{r}

model_def_tbl <- lapply(get_model_names(), get_model_eq) %>% bind_rows(, .id = NULL) %>% tibble()
                                        #print(model_def_tbl, n = 200, width = 200)

str_rm <- c("exp", "[0-9.]+", "log(2|10|)", "sin", "abs", "pi", "temp")
pattern <- paste0("\\b", paste0(str_rm, collapse = "\\b|\\b"), "\\b")

## more robust than param_old
param <- lapply(model_def_tbl$model, function(x) {
    args <- names(formals(x))
    args <- args[args !="temp"]
    str <- paste(args, collapse = ", ")
}) %>% unlist()

names(param) <- model_def_tbl$model

param_old  <- stringi::stri_extract_all_words(model_def_tbl$eq) %>%
    lapply(., unique) %>%
    lapply(., paste, collapse= " ") %>%
    str_replace_all(., pattern, "") %>%
    str_replace_all(., "( +$|^ +)", "") %>%
    str_split(., " +") %>%
    lapply(., sort) %>%
    lapply(., paste, collapse = ", ") %>% unlist()


n_param <- param %>% str_count(., boundary("word"))

model_tbl <- bind_cols(model_def_tbl, n_param= n_param) %>%
    bind_cols(. , param = param) %>%
    arrange(n_param, model) %>%
    relocate(eq, .after = param) %>%
    mutate(tmax = as.integer(grepl(pattern = "tmax", x = param, fixed = TRUE))) %>% 
    mutate(tmin = as.integer(grepl(pattern = "tmin", x = param, fixed = TRUE)))

grid.arrange(tableGrob(model_tbl))

model_tbl_tmax <- model_tbl %>% filter(tmax == 1)

model_tbl_tmax %>% arrange("model")

model_tbl_tmin <- model_tbl %>% filter(tmax == 1)

model_tbl_tmax_no_tmin <- model_tbl %>% filter(tmax == 1 & tmin == 0)

## model_tbl
```

## Load Data

```{r}
load(file.path("input", "data.processing_2022-11-09.Rda"),
     verbose = TRUE)

```


# Examine Data

## Create Working Dataset

```{r}

males_filtered_disp <- motif_stats_40C %>%
    filter(dispersion < 50) %>%
    pull(male)

males_filtered_mean <- motif_stats %>%
    filter(mean > 10) %>%
    pull(male)

males_filtered <- intersect(males_filtered_mean, males_filtered_disp)

##males_selected <-

data_ind <- motif_data %>%
    filter(male %in% males_filtered) %>%
    arrange(male) %>%
    ##    left_join(male_shape, by = "male") %>%
    mutate()

data_aug <- tibble(
    male = rep.int(
    (data_ind$male %>% unique()), 5),
    temp = 46,
    motif_count = 0,
    motif_prop = 0,
    round = 0,
    weights = 10) %>%
    bind_rows(data_ind, .)
## copy data frame and assign `male =  "combined")

data_comb <- data_ind %>% mutate(male = "combined")

stats_ind <- motif_stats %>%
    filter(male %in% males_filtered)

```


## Plot motif_count


```{r}

g1 <- ggplot(data = data_ind) +
    aes(x = temp, y = motif_count, color = male, shape = male) +
    ## Redefine shapes. Note need to set 'shape = male' above to prevent there from
    ## begin two legends: 1 for shape and 1 for color.
    scale_shape_manual(values = rep(c(16:18), length.out = length(males_filtered))) +
    geom_point() +
    scale_color_viridis_d() +
    labs(title = "motif_count") +
    theme(legend.position="none")

```



```{r}

g1 <- ggplot(data = data_ind) +
    aes(x = temp, y = motif_count, color = male, shape = male) +
    ## Redefine shapes. Note need to set 'shape = male' above to prevent there from
    ## begin two legends: 1 for shape and 1 for color.
    scale_shape_manual(values = rep(c(16:18), length.out = length(males_filtered))) +
    geom_point() +
    scale_color_viridis_d() +
    labs(title = "motif_count") +
    theme(legend.position="none")

g2 <- ggplot(data = data_ind) +
    aes(x = temp, y = motif_prop, color = male, shape = male) +
    scale_shape_manual(values = rep(c(16:18), length.out = length(males_filtered))) +
    geom_point() +
    scale_color_viridis_d() +
    labs(title = "motif_prop") +
    theme(legend.position="bottom")


legend <- get_legend(g2)

g2 <- g2 + theme(legend.position="none")

g3 <- tableGrob(format(data.frame(stats_ind %>% select(male, n_obs, total, mean) %>% unique() ),
                       digits = 1),
                theme = ttheme_default(base_size = 8))

grid.arrange(g1, g2, g3, legend, ncol = 2,
             top=textGrob("Males filtered for dispersion < 50 at 40C & count_mean < 10",
                          gp=gpar(fontsize = 11))
             )


```

# Analyze Data: 


## Models with 3 or 4 parameters

```{r}

model_set <- model_tbl %>% filter(n_param < 5 & n_param > 2)
model_grob <- tableGrob(model_set)

grid.arrange(model_grob)
```
### Mulitple Fits

```{r}

maxiter <- 4000
minFactor <- min(1/(2*maxiter), 1/1024)

weights_index <- 1:1
data_indexes <- 1:2
data_descriptions <- c("Unaugmented", "motif_count at 46C Augmented")

for(data_index in data_indexes ){
    data <- switch(data_index,
                   data_ind,
                   data_aug
                   )
    data_desc <- data_descriptions[[data_index]]
    data <- data %>% rename(rate = motif_prop) %>%
        select(temp, rate, weights, motif_count_plus_1) %>%
        data.frame()


    ## Define x_range for predictions
    x_range <- data.frame(temp = seq(min(data$temp), max(data$temp), 0.5))

    pred_tbl <- tibble(x_range)

    fit_list <- list()
    graph_list <- list()
    tgrob_list <- list()
    pred_list <- list()
    for(wts in weights_index){
        pred_list[[wts]] <- x_range
    }


    g_base <- ggplot(data, aes(temp, rate)) +
        geom_point()

    g_preds <- g_base
    
    for(model_str in model_set$model) {

        n_param <- formals(model_str) %>% length() - 1
        iter <- 10 ##rep(5, n_param)

        model_fits <- list()
        model_graphs <- list()
        model_tgrobs <- list()
        ##    model_preds <- list()


        start_vals <- get_start_vals(data$temp, data$rate, model_name = model_str)
        lower <- get_lower_lims(data$temp, data$rate, model_name = model_str)
        if(sum(!is.finite(lower))) {
            print(paste0("Warning() in ", model_str, " lower = ", lower, " replacing problematic values with -1000"))
            lower[!is.finite(lower)] <- -1000
        }
        upper <- get_upper_lims(data$temp, data$rate, model_name = model_str)
        if(sum(!is.finite(upper))) {
            print(paste0("Warning() in ", model_str, " upper = ", upper, " replacing problematic values with 1000", collapse = " "))
            upper[!is.finite(upper)] <- 1000
        }
        
        my.formals <- names(formals(model_str)) %>% paste(., collapse = ", ") %>% sub("temp,", "temp = temp,", .)
        formula <- paste0("rate ~ ", model_str, "(", my.formals, ")") 
        for(wts in weights_index) {

            ## WARNING: if you use 'weights' as a variable, it uses the column in data$weights even if weights is locally defined.
            
            local.weights <- switch(wts,
                                    rep(1, nrow(data)),
                                    data$weights
                                    )
            
            
            fit_db <- nls_multstart_db(formula = formula,
                                 data = data,
                                 iter = iter, 
                                 start_lower = lower*1.01,
                                 start_upper = upper*0.99,
                                 ## lower = lower, ## These are passed to nls
                                 ## upper = upper, ## These are passed to nls
                                 supp_errors = 'Y',
                                 modelweights =  local.weights,
                                 convergence_count = FALSE, ## misleading terminology, set to FALSE
                                 control = c(maxiter = maxiter,
                                             maxfev = 100000,
                                             minFactor = minFactor)
                                 )
            
            get_aic <- function(fit) {
                AICval <- ifelse(!is.null(fit), stats::AIC(fit), 
                                 Inf)
                return(AICval)
            }

            AIC_all <- lapply(fits_all, get_aic)
            fit <- first(fits_all)
            

            summary(fit) %>% print()
            ##fit_list[[model_str]] <- fit

            ## calculate additional traits
            if(FALSE) calc_params(fit) %>% mutate_all(round, 2)
            
            ## Get predictions of our model using broom::augment(newdata = x_range)
            ## `newdata` argument runs the predict() function and returns
            ## predictor values (temp) and fitted values (.fitted).
            preds <- augment(fit, newdata = x_range)
            ## plot data and model fit
            g <- g_base +
                geom_line(aes(temp, .fitted), preds, col = 'blue') +
                theme_bw(base_size = 12) +
                labs(x = 'Temperature (ºC)',
                     y = 'Song Count',
                     title = paste0("Fitting ", model_str, ", using weight_index: ", wts))
            g_preds <- g_preds +
                geom_line(aes(temp, .fitted), preds, col = 'blue')
            model_graphs[[wts]] <- g

            model_fits[[wts]] <- fit
            pred_list[[wts]] <- left_join(pred_list[[wts]], preds, by = "temp") %>%
                rename("{model_str}" := .fitted)
            
            ##        model_tgrobs[[wts]] <- text_grob(label = summary(fit))

        }

        if(FALSE) {
            ##grid.arrange(grobs = c(model_graphs, model_tgrobs), top = model_str)
            grid.arrange(grobs = model_graphs, top = model_str)
        }
        
        fit_list[[model_str]] <- model_fits;
        graph_list[[model_str]] <- model_graphs;
        tgrob_list[[model_str]] <- model_tgrobs;

    }

}



for(data_index in data_indexes ){
    data_desc <- data_descriptions[[data_index]]
    for(wts in weights_index){
        pred_data <- pred_list[[wts]]
        data_long <- melt(pred_data, id = "temp", variable_name= "model")
        
        g <- g_base + geom_line(data = data_long, aes(x = temp, y = value, color = variable)) +
            labs(x = 'Temperature (ºC)',
                 y = 'Song Prop',
                 title = paste0("Using weight_index: ", wts, " and data: ", data_desc ))
    }
    print(g)
}

```

