---
title: "Gilchrist Chamber Temp Compare"
author: "Tara Empson"
date: "`r Sys.Date()`"
output: html_document
---
```{r, include=FALSE}
library(tidyverse)
library(patchwork)
setwd("/Users/taraempson/Documents/GitHub/ZEFI-Thermal-Curves/data/song/collated/")
round3<- read.csv("HSPi-Repeatability-Song-Count.csv")
round2<- read.csv("HSPi-Round-2-Heat-Trials.csv")
```

# Looking at the Round 3. 
## Repeated trials at the same target temperature.
```{r echo=FALSE, message=FALSE, warning=FALSE}
ch2<- round3 %>% 
    filter(chamber=="2") %>% 
    ggplot(aes(x=date, y=temp_mean))+
    geom_point()+
    ylim(36,41)+
    geom_hline(aes(yintercept=40, color="red"))+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
    
    labs(x="Date", y="Head Height Logger Mean", title="Chamber 2")

ch4 <- round3 %>% 
    filter(chamber=="4") %>% 
    ggplot(aes(x=date, y=temp_mean))+
    geom_point()+
    ylim(36,41)+
    geom_hline(aes(yintercept=40, color="red"))+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
    labs(x="Date", y="Head Height Logger Mean", title="Chamber 4")

ch6 <- round3 %>% 
    filter(chamber=="6") %>% 
    ggplot(aes(x=date, y=temp_mean))+
    geom_point()+
    ylim(36,41)+
    geom_hline(aes(yintercept=40, color="red"))+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
    labs(x="Date", y="Head Height Logger Mean", title="Chamber 6")

ch2+ch4+ch6
```

# Looking at the Round 2 in a similar way. 
## But instead of separating by chamber it would be better to separate by target temperatures.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
t27 <- round2 %>% 
    filter(temp_target=="27") %>% 
    ggplot(aes(x=date, y=temp_mean, color=as.factor(chamber)))+
    geom_point()+
    ylim(25,46)+
    geom_hline(aes(yintercept=27, fill="black"))+
    geom_smooth(method = "lm", formula = y ~ x, color="red")+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
    labs(x="Date", y=str_wrap("Head Height Logger Mean", width = 15), title="Target 27")

t35 <- round2 %>% 
    filter(temp_target=="35") %>% 
    ggplot(aes(x=date, y=temp_mean, color=as.factor(chamber)))+
    geom_point()+
    ylim(25,46)+
    geom_hline(aes(yintercept=35, fill="black"))+
    geom_smooth(method = "lm", formula = y ~ x, color="red")+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
    labs(x="Date", y=str_wrap("Head Height Logger Mean", width = 15), title="Target 35")

t40 <- round2 %>% 
    filter(temp_target=="40") %>% 
    ggplot(aes(x=date, y=temp_mean, color=as.factor(chamber)))+
    geom_point()+
    ylim(25,46)+
    geom_hline(aes(yintercept=40, fill="black"))+
    theme(axis.text.x = element_text(angle = 90))+
    labs(x="Date", y=str_wrap("Head Height Logger Mean", width = 15), title="Target 40", color="Chamber")

t42 <- round2 %>% 
    filter(temp_target=="42") %>% 
    ggplot(aes(x=date, y=temp_mean, color=as.factor(chamber)))+
    geom_point()+
    ylim(25,46)+
    geom_hline(aes(yintercept=42, fill="black"))+
    geom_smooth(method = "lm", formula = y ~ x, color="red")+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
    labs(x="Date", y=str_wrap("Head Height Logger Mean", width = 15), title="Target 42")

t44<- round2 %>% 
    filter(temp_target=="44") %>% 
    ggplot(aes(x=date, y=temp_mean, color=as.factor(chamber)))+
    geom_point()+
    ylim(25,46)+
    geom_hline(aes(yintercept=44, fill="black"))+
    geom_smooth(method = "lm", formula = y ~ x, color="red")+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
    labs(x="Date", y=str_wrap("Head Height Logger Mean", width = 15), title="Target 44")

t27+t35+t40+t42+t44
```

# Checking the relationship between target temp and mean head height temp for Round 2. 
### Looks really linear, less of a spread than I expected.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
round2 %>% 
    ggplot(aes(x=temp_target, y=temp_mean, color=as.factor(chamber)))+
    geom_point()+
    ylim(25,46)+
    geom_smooth(method = "lm", formula = y ~ x, color="red")+
    labs(x="Target Temp", y="Head Height Logger Mean", color="Chamber")
```

### I'm not certain on how to interpret AIC scores, I think this is saying that chamber's/temp logger do vary a bit in how close they get to target temp but not much.
```{r}
mod <- lm(temp_mean~temp_target+date+chamber, data=round2)
step(mod)
```