---
title: "Compare NegBinomial with Constrained Shape to Poisson"
author: "Michael Gilchrist (mikeg@utk.edu)"
date: "Created: 2023-02-24; Code run on `r Sys.Date()`"
output: html_notebook
---

```{r}
library(tidyr)
library(dplyr)
library(purrr)
library(ggplot2)

```


```{r}

mu =c(1, 4, 64, 256)
x <- 0:400
theta <- c(1.000001,2,4,8)
```

## Shape (or `size` in R's `pnbinom()`) is function of mean and theta

This approach ensures that the variance scales linearly with the expected value.

```{r}
tbl_1 <- crossing(
  x,
  theta, 
  mu) %>%
 mutate(size = mu/(theta-1), y = dnbinom(x = x, mu = mu, size = size), theta = factor(round(theta))) %>%
  group_by(theta, mu) %>%
  filter( (x < mu + 5*sqrt(mu)))

ggplot(tbl_1, aes(x, y, color = theta, fill = theta)) + 
   geom_line(alpha = 1) +#
#  scale_y_log10() + 
  facet_wrap(vars(mu), 
             scales = "free",
  #labeller = label_both
  labeller = label_bquote(mu:.(mu))
  ) +
  labs(title = "shape = mu/(theta-1)",
       subtitle = "var(y) = theta mu")
```


## Shape is constant

This is the standard approach for negative binomial regression
```{r}
tbl_2 <- crossing(
  x,
  theta, 
  mu) %>%
  mutate(size = theta, y = dnbinom(x = x, mu = mu, size = size), size = factor(round(size))) %>%
  group_by(size, mu) %>%
  filter( (x < mu + 5*sqrt(mu)))

ggplot(tbl_2, aes(x, y, color = size, fill = size)) + 
   geom_line(alpha = 1) +#
#  scale_y_log10() + 
  facet_wrap(vars(mu), scales = "free",
  labeller = label_bquote(mu:.(mu))
  ) +
  labs(title = "shape = Constant",
       subtitle = expression("var(y) = "~mu+mu^2/shape))
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
