---
title: "Piecewise Regression"
author: "Michael Gilchrist"
date: "date: 2022-11-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
  )


if(interactive()) default::default(.ess.eval) <- list(max.deparse.length=1E5, output = TRUE)

```
# Goal

- Fit piecewise GLM 

## Recap

# Set up

## Install libraries

```{r}

## install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(GGally)
library(segmented)
library(broom)
library(tidyverse)
library(viridisLite)
library(rstan)
#options(ggplot2.continuous.colour="viridis",
#        ggplot2.discrete.colour="viridis",
#        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
#        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(lme4)
library(latex2exp)

```

## Load Data

```{r}
load(file.path("input", "data.processing_2022-11-09.Rda"),
     verbose = TRUE)

```


# Examine Data

## Create Working Dataset

```{r}

filter_data <- TRUE

if(filter_data) {
    males_filtered_disp <- song_stats_40C %>%
        filter(dispersion < 50) %>%
        pull(male)
    
    males_filtered_mean <- song_stats %>%
        filter(mean > 10) %>%
        pull(male)

    male_vector <- intersect(males_filtered_mean, males_filtered_disp)
} else {
    male_vector <- song_data %>% select(male) %>% distinct()
}


##males_selected <-

data_ind <- song_data %>%
    filter(male %in% male_vector) %>%
    arrange(male) %>%
    ##    left_join(male_shape, by = "male") %>%
    mutate()

stats_ind <- song_stats %>%
    filter(male %in% male_vector)

```


## Plot song_count


```{r}

g1 <- ggplot(data = data_ind) +
    aes(x = temp, y = song_count, color = male, shape = male) +
    ## Redefine shapes. Note need to set 'shape = male' above to prevent there from
    ## begin two legends: 1 for shape and 1 for color.
    scale_shape_manual(values = rep(c(16:18), length.out = length(male_vector))) +
    geom_point() +
    scale_color_viridis_d() +
    labs(title = "song_count") +
    theme(legend.position="none")

g2 <- ggplot(data = data_ind) +
    aes(x = temp, y = song_prop, color = male, shape = male) +
    scale_shape_manual(values = rep(c(16:18), length.out = length(male_vector))) +
    geom_point() +
    scale_color_viridis_d() +
    labs(title = "song_prop") +
    theme(legend.position="bottom")


legend <- get_legend(g2)

g2 <- g2 + theme(legend.position="none")

g3 <- tableGrob(format(
    data.frame(stats_ind %>%
               select(male, n_obs, total, mean) %>% unique()),
                       digits = 1),
    theme = ttheme_default(base_size = 8))

grid.arrange(g1, g2, g3, legend, ncol = 2,
             top=textGrob("Males filtered for dispersion < 50 at 40C & count_mean < 10",
                          gp=gpar(fontsize = 11))
             )


```

# Analyze Data:


## Simple GLM


```{r}

glm_qpoisson_1 <- glm(song_count ~
                       (1 + male),
                   data = data_ind,
                   family = quasipoisson(link = "identity")
                   )


summary(glm_qpoisson_1)

```

## Stan


### Normal error

```{r}

y <- data_ind %>% pull(song_count)
x <- data_ind %>% pull(temp)
N <- length(y)
iter <- 10000
fit_1 <- stan(file = "one.step_piecewise_normal.stan",
              model_name = "one step gaussian",
              data=list(x = x, y = y,
                        N = N,
                        tmax = 46,
                        y_tmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = iter/2)
```

- Even with a (poorly chosen) prior on sigma, this fails to converge.
  Here's some output.
  
  > Chain 1: Iteration: 6000 / 6000 [100%]  (Sampling)
  > Chain 1: 
  > Chain 1:  Elapsed Time: 15.038 seconds (Warm-up)
  > Chain 1:                14.647 seconds (Sampling)
  > Chain 1:                29.685 seconds (Total)
  > Chain 1: 
  > Warning messages:
  > 1: There were 1 divergent transitions after warmup. See
  > https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
  > to find out why this is a problem and how to eliminate them. 
  > 2: There were 3819 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. See
  > https://mc-stan.org/misc/warnings.html#maximum-treedepth-exceeded 
  > 3: Examine the pairs() plot to diagnose sampling problems
  >  
  > 4: The largest R-hat is 1.55, indicating chains have not mixed.
  > Running the chains for more iterations may help. See
  > https://mc-stan.org/misc/warnings.html#r-hat 
  > 5: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
  > Running the chains for more iterations may help. See
  > https://mc-stan.org/misc/warnings.html#bulk-ess 
  > 6: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
  > Running the chains for more iterations may help. See
  > https://mc-stan.org/misc/warnings.html#tail-ess 
  
  

### Poisson

    
```{r}

y <- data_ind %>% pull(song_count)
x <- data_ind %>% pull(temp)
N <- length(y)
iter <- 6000
fit_2 <- stan(file = "one.step_piecewise_poisson.stan",
              model_name = "one step poisson",
              data=list(x = x, y = y,
                        N = N,
                        tmax = 46,
                        y_tmax = 0),
              chains = 4,
              cores = 4,
              iter = iter,
              warmup = iter/2)

## Examine output

traceplot(fit_2, pars = c("a1, bt1"), inc_warmup = TRUE)

```

### Stan data example
```{r}

y_1 <- c(3, 5)
y_2 <- rep(c(2,3,4,5), c(10,20,30,40))
y <- c(y_1, y_2)
N <- length(y)
movie <- rep(c(1,2), c(length(y_1), length(y_2)))
movie_data <- list(y=y, N=N, movie=movie)
#fit_1 <- stan("ratings_1.stan", data=movie_data)

```
